### 2019-01-02 docs/supplement.Rnw
start with Oslo pipeline for GDSC and CCLE

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\bsigma}{\bm{\sigma}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bepsilon}{\bm{\epsilon}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bPsi}{\bm{\Psi}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bgamma}{\bm{\gamma}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\Em}{\mathbf{E}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\Lm}{\mathbf{L}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Cov}{\mathbb{C}\text{ov}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

% R packages
<<echo=FALSE>>=
library(cambridge)
@

% settings
\graphicspath{{/Users/magnusmunch/Documents/OneDrive/PhD/project_cambridge/graphs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Supplementary material to: Drug efficacy prediction in cell lines}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

\begin{document}
	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge, United Kingdom \\

	\section{Introduction}
	This document contains the Supplementary Material (SM) to the document `Drug efficacy prediction in cell lines'. In the following, this document is referred to as Main Document (MD).

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	In the following all expectations are with respect to the variational posterior $Q$.
	\begin{align*}
	\log q_{\B} (\B) & \propto \E_{\mathbf{U}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\bgamma^2, \bsigma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\mathbf{u}_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E_{\gamma^2_d, \sigma^2_d} \( \frac{\beta_{jd}^2}{\gamma_d^2 \sigma_d^2} \), \\
	q_{\B} (\B) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	& \text{ with } \bSigma_d = \E(\sigma_d^{-2})^{-1} [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1}, \\
	& \text{ and } \bmu_d = [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1} \X \tr [\y_d - \Z \cdot \E(\mathbf{u}_d)].
	\end{align*}

	\begin{align*}
	\log q_{\U}(\U) & \propto \E_{\mathbf{B}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2} [\log \pi (\U | \mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\bbeta_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \E_{\bm{\Xi}_d, \sigma_d^2}\( \frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \), \\
	q_{\U}(\U) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_T (\mathbf{m}_d, \Sm_d), \\
	& \text{ with } \Sm_d = \E(\sigma_d^{-2})^{-1} [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1}, \\
	& \text{ and } \mathbf{m}_d = [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1} \Z \tr [\y_d - \X \cdot \E(\bbeta_d)].
	\end{align*}

	\begin{align*}
	\log q_{\bgamma^{2}}(\bgamma^{2}) & \propto \E_{\B, \bsigma^2} [\log \pi (\B | \bgamma^{2}, \sigma^2)] + \log \pi (\bgamma^{2}) \\
	& \propto -\frac{p}{2} \sum_{d=1}^D \log \gamma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} - \frac{3}{2} \sum_{d=1}^D \log \gamma_d^{2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{\lambda}{2 \theta^2} \sum_{d=1}^D (\gamma_d^2 - \theta)^2 \gamma_d^{-2} \\
	& \propto \sum_{d=1}^D \left\{ \( -\frac{p + 1}{2} - 1\) \log \gamma_d^{2} - \frac{\lambda}{2\theta^2} \gamma_d^2 - \frac{1}{2}\[ \lambda + \E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^{2}}\] \gamma_d^{-2} \right\}, \\
	q_{\bgamma^{2}}(\bgamma^{2}) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG} \(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_{d} \), \\
	& \text{with } \delta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} + \lambda.
	\end{align*}

	\begin{align*}
	\log q_{\bsigma^{2}}(\bsigma^{2}) & \propto \E_{\B,\mathbf{U}} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\B, \bgamma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \,\,\,\,\,\,\,\,\,\, + \E_{\mathbf{U}, \bm{\Xi}_1, \dots, \bm{\Xi}_D} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bsigma^{2}) \\
	& \propto -\frac{n}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \sigma_d^{2} - \frac{p}{2} \sum_{d=1}^D \log \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \E( \gamma_d^{-2}) \frac{\E ( \beta_{jd}^2)}{\phi_{g(j)}^{2}} \sigma_d^{-2} - \frac{T}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \E(\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}) \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{3}{2} \sum_{d=1}^D \log \sigma^2_d \\
	& = -\sum_{d=1}^{D} \(\frac{n + p + T + 3}{2} \) \log \sigma_d^{2} - \sum_{d=1}^{D} \frac{1}{2}\Bigg\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \\
	& \,\,\,\,\,\,\,\,\,\, + \E(\gamma_d^{-2})\sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^2} + \E(\mathbf{u}_{d} \tr \bm{\Xi}_d^{-1} \mathbf{u}_d) \Bigg\} \sigma_d^{-2}, \\
	q_{\bsigma^{2}}(\bsigma^{2}) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + T + 1}{2}, \zeta_{d} \), \\
	& \text{ with } \zeta_{d} = \frac{\y_d \tr \y_d}{2} - \y_d \tr [\X \E(\bbeta_d) + \Z \E(\mathbf{u}_d)] + \E(\bbeta_d \tr) \X \tr \Z \E (\mathbf{u}_d) \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\trace [\X \tr \X \V(\bbeta_d)]}{2} + \frac{\E(\bbeta_d \tr) \X \tr \X \E(\bbeta_d)}{2} + \frac{\trace [\Z \tr \Z \V(\mathbf{u}_d)]}{2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\E(\mathbf{u}_d \tr) \Z \tr \Z \E(\mathbf{u}_d)}{2} + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+ \frac{\E(\mathbf{u}_d \tr) \E(\bm{\Xi}_d^{-1}) \E(\mathbf{u}_d)}{2} + \frac{\trace [\E(\bm{\Xi}_d^{-1}) \V(\mathbf{u}_d)]}{2},
	\end{align*}

	\begin{align*}
	\log q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \propto \E_{\mathbf{U}, \bm{\sigma}^2} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D) \\
	& \propto -\frac{1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \E \(\frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \) \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{k + T + 1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace (\bm{\Omega} \bm{\Xi}_d^{-1}) \\
	& = -\frac{\tau + T + 2}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace \left\{ \[\E(\sigma_d^{-2}) \E( \mathbf{u}_d \mathbf{u}_d \tr) + \bm{\Omega}\] \bm{\Xi}_d^{-1} \right\}, \\
	q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \overset{D}{\propto} \prod_{d=1}^D \mathcal{W}_T^{-1} (\bm{\Psi}_d, \nu + 1),  \\ 
	& \text{ with } \bm{\Psi}_d =\E(\sigma^{-2}_d) \[\V(\mathbf{u}_d) + \E(\mathbf{u}_d) \E( \mathbf{u}_d \tr) \] + \bm{\Omega}.
	\end{align*}

  \section{Model without tissue effects}
  \subsection{Introduction}
  In order to investigate the inverse Gaussian prior model and its estimation, we consider a simpler model. The simpler model ignores tissue effects and fixes $\forall g: \phi_g^2=1$.

  \subsection{Model and estimation}
  The model is given by:
  \begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \\
  \gamma_d^2 & \sim \mathcal{IG}(\theta_d, \lambda), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}
  The corresponding variational Bayes posterior is given by:
  \begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_d\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
  We update the variational parameters until convergence:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = \frac{2 \zeta_d^{(h)}}{n+p+1} (\X \tr \X + a_d^{(h)} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + a_d^{(h)} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \frac{n+p+1}{2 \zeta_d^{(h)}} [\trace(\bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}] + \lambda, \\ \zeta_d^{(h+1)} & = \frac{a_d^{(h+1)}}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace [ (\X \tr \X + \I_p) \bm{\Sigma}_d^{(h+1)}] + (\bm{\mu}_d^{(h+1)}) \tr (\X \tr \X + \I_p)\bm{\mu}_d^{(h+1)}\],
  \end{align*}
  with $a_d^{(h)}$ as in the full model in the MD. Equivalently, the empirical Bayes updates for $\bm{\alpha}$ and $\lambda$ are the same as in the full model in the MD.

  \subsection{Evidence lower bound}
  We monitor convergence through the evidence lower bound (ELBO):
  \begin{align*}
  \text{ELBO} & (Q^{(l)}) \propto \frac{1}{2} \sum_{d=1}^D \log |\bm{\Sigma}_d^{(l)}| - \frac{n + p + 1}{2} \sum_{d=1}^D \log \zeta_d^{(l)} + \sum_{d=1}^D \log K_{\frac{p+1}{2}}\(\sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}\) \\
  & - \frac{n+p+1}{4} \sum_{d=1}^D (\zeta_d^{(l)})^{-1} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
  & + \frac{D(p+3)}{4} \log \lambda^{(l)} + \lambda^{(l)} \sum_{d=1}^D (\theta_d^{(l)})^{-1} - \frac{p+1}{2} \sum_{d=1}^D \log \theta_d^{(l)} - \frac{p+1}{4} \sum_{d=1}^D \log \delta_d^{(l)} \\
  & + \frac{1}{2} \sum_{d=1}^D \[ \delta_d^{(l)} - \lambda^{(l)} - \frac{n+p+1}{2\zeta_d^{(l)}} (\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - \frac{n+p+1}{2\zeta_d^{(l)}} \trace (\bm{\Sigma}_d^{(l)}) \] a_d^{(l)}. 
  \end{align*}

  - Different options for convergence. We may monitor parameters, ELBO, or a combination. We may also iterate a fixed number of times, especially for the VB (currently doing this, using 2 VB iterations and 20 EB iterations).

  \subsection{Efficient computation}
  The empirical Bayes updates require the following quantities: $\zeta_d$, $a_d$, $\delta_d$, $\trace (\bm{\Sigma}_d)$, $\trace (\X \tr \X \bm{\Sigma}_d)$, $\bm{\mu}_d \tr \bm{\mu}_d$, and $\bm{\mu}_d \tr \X \tr \X \bm{\mu}_d$. In addition, we need $\log|\bm{\Sigma}|$ and $\mathbf{y}_d \tr \X \bm{\mu}_d$ to monitor the ELBO. The first three quantities are obtained by scalar operations of $\mathcal{O}(n)$. The rest is easily calculated using the SVD $\X = \mathbf{U} \mathbf{D} \mathbf{V} \tr$:
  \begin{align*}
  \trace (\bm{\Sigma}_d^{(h+1)}) & = \frac{2\zeta_d^{(h)}}{n + p + 1} \[\sum_{i=1}^n (v_i^2 + a_d^{(h)})^{-1} + \max (p - n, 0)\cdot (a_d^{(h)})^{-1}\], \\
  \trace (\X \tr \X \bm{\Sigma}_d^{(h+1)}) & = \frac{2\zeta_d^{(h)}}{n + p + 1} \sum_{i=1}^n v_i^2(v_i^2 + a_d^{(h)})^{-1}, \\
  (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + a_d^{(h)})^2, \\
  (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^4 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + a_d^{(h)})^2, \\
  \log|\bm{\Sigma}^{(h+1)}| & \propto p \log \zeta_d^{(h)} - \sum_{i=1} \log (v_i^2 + a_d^{(h)}) - \max(p - n, 0) \cdot \log a_d^{(h)}, \\
  \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + a_d^{(h)}).
  \end{align*}
  where $v_i$ are the singular values of $\X$. The SVD and $\mathbf{U} \tr \mathbf{y}_d$ are $\mathcal{O}(pn^2)$ and $\mathcal{O}(n^2)$ operations, respectively, but have to be calculated only once at the start of the algorithm. The rest of the calculations involve just $n$ scalar multiplications and/or additions.

  \subsection{Independent error variance model parameters}
  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ and $\sigma_d^2$ should be independent \textit{a priori}. However, in our case that is debatable. The $D$ regressions need not be on the same scale, so some form of calibration of the $\bm{\beta}_d$ posteriors to this scale is required. A simple way of doing this is to include the scales $\sigma_d^2$ in the priors for the $\bm{\beta}_d$. For a sense of completeness we still compare to the independent prior model in our simulations. The independent prior model is given by:
  \begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \gamma_d^2 & \sim \mathcal{N} (0, \gamma_d^2), \\
  \gamma_d^2 & \sim \mathcal{IG}((\mathbf{c}_d \tr \bm{\alpha})^{-1}, \lambda), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}
  Its variational posterior is given by:
  \begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+1}{2}, \lambda (\mathbf{c}_d \tr \bm{\alpha})^2, \delta_d\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + 1}{2}, \zeta_d\),
  \end{align*}
  with variational parameter updates:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = \(\frac{n+1}{2 \zeta_d^{(h)}}\X \tr \X + a_d^{(h)} \I_p\)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = \frac{n+1}{2 \zeta_d^{(h)}} \bm{\Sigma}_d^{(h+1)} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \trace(\bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)} + \lambda, \\ \zeta_d^{(h+1)} & = \frac{1}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace (\X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)}\],
  \end{align*}
  with $a_d^{(h)}$ and the empirical Bayes updates $\bm{\alpha}^{(l+1)}$ and $\lambda^{(l+1)}$ as in the full model in the MD. A singular value decomposition $\X = \mathbf{U} \mathbf{D} \mathbf{V} \tr$ leads to the following efficient computations:
  \begin{align*}
  \trace (\bm{\Sigma}^{(h+1)}_d) & = \frac{2\zeta^{(h)}_d}{n + 1} \sum_{i=1}^n \(v_i^2 + \frac{2\zeta^{(h)}_d a^{(h)}_d}{n+1}\)^{-1} + \max (p - n, 0)\cdot (a^{(h)}_d)^{-1}, \\
  \trace (\X \tr \X \bm{\Sigma}^{(h + 1)}_d) & = \frac{2\zeta^{(h)}_d}{n + 1} \sum_{i=1}^n v_i^2 \(v_i^2 + \frac{2\zeta^{(h)}_d a^{(h)}_d}{n+1}\)^{-1}, \\
  (\bm{\mu}_d^{(h + 1)}) \tr \bm{\mu}_d^{(h + 1)} & = \sum_{i=1}^n [(\mathbf{U} \tr \mathbf{y}_d)_i]^2 v_i^2 \(v_i^2 + \frac{2 \zeta_d^{(h)} a_d^{(h)}}{n + 1}\)^{-2}, \\
  (\bm{\mu}_d^{(h + 1)}) \tr \X \tr \X \bm{\mu}_d^{(h + 1)} & = \sum_{i=1}^n [(\mathbf{U} \tr \mathbf{y}_d)_i]^2v_i^4 \(v_i^2 + \frac{2 \zeta_d^{(h)} a_d^{(h)}}{n + 1}\)^{-2}, \\
  \log|\bm{\Sigma}^{(h + 1)}| & \propto p \log \zeta_d^{(h)} - \sum_{i=1} \log \(v_i^2 + \frac{2 \zeta_d^{(h)} a_d^{(h)}}{n + 1}\) \\
  & \,\,\,\,\,\,\,\,\,\,\, - \max(p - n, 0) \cdot (\log a_d^{(h)} + \log \zeta_d^{(h)}), \\
  \mathbf{y}_d \tr \X \bm{\mu}_d^{(h + 1)} & = \sum_{i=1}^n [(\mathbf{U} \tr \mathbf{y}_d)_i]^2 v_i^2 \(v_i^2 + \frac{2 \zeta_d^{(h)} a_d^{(h)}}{n + 1}\),
  \end{align*}
  where $v_i$ are the singular values of $\X$. As before, this renders $\mathcal{O}(pn^2)$ and $\mathcal{O}(n^2)$ calculations once at the start of the algorithm and iterative updates of $\mathcal{O}(n)$.

  \subsection{Simulations}
  We simulate from a correctly specified model. We assume random $y_{id}$, $\x_i$, and $\bm{\beta}_d$, while we fix $\mathbf{C}$, $\gamma_d^2$, $\sigma_d^2$ and $\bm{\alpha}$. We estimate the $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, $\bm{\alpha}$ and $\lambda$, while we fix the $\sigma_d^2$ to their true values.

  We set $n=100$, $p=200$, and $D=10$. We consider 5 evenly sized classes of drugs, such that $\mathbf{c}_{kd} = \mathbbm{1}(k = \text{class}_d)$ and $\begin{bmatrix} \mathbf{c}_1 & \cdots & \mathbf{c}_D \end{bmatrix} \tr$. From these the $\gamma_d^2 = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ are created. Next, we simulate $\bm{\beta}_{jd} \sim \mathcal{N} (0, \sigma^2_d \gamma_d^2)$. We fix the mean signal-to noise ratio $\overline{\text{SNR}}_d$, such that the predictor data variance is $s^2 = \V(\mathbf{x}_i) = \overline{\text{SNR}}_d / (\overline{\gamma^2_d} p)$ (NOT CORRECT). The last step consists of simulating the data by $\mathbf{x}_{ij} \sim \mathcal{N}(0,s^2)$ and $y_{id} = \mathcal{N} (\mathbf{x}_i \tr \bm{\beta}_d, \sigma_d^2)$. The remaining parameters are set as follows: $\bm{\alpha} = \begin{bmatrix} 1 & \dots & 5 \end{bmatrix} \tr$, $\sigma_d^2 = 1$ for all $d$, and $\overline{\text{SNR}}_d = 10$. To account for the random data and parameter generation, we repeat the simulation 100 times.

  We compare the results to a model where the $\gamma_d^2$ are drawn from a class-specific inverse Gamma distribution; a generalisation of the models in \cite{kpogbezan_empirical_2017} and \cite{leday_gene_2017}:
  \begin{align*}
  y_{id} | \bm{\beta}_d ,\sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \gamma_d^2 ,\sigma_d^2 & \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2), \\
  \gamma_d^2 & \sim \Gamma^{-1} (a_{\text{class}(d)}, b_{\text{class}(d)}), \\
  \sigma_d^2 & \sim \Gamma^{-1} (0.001, 0.001),
  \end{align*}
  where $\text{class}(c)$ denotes the class of drug $d$. This allows for more flexible empirical Bayes estimation of the $a_d$ and $b_d$, and thus prior means of the different drug classes. As a consequence however, the risk of overfitting is increased. Additionally, it does not allow to include continuous covariates on the drugs. We present the results in Figures \ref{fig:boxplot_igaussian_res1_sigma2}-\ref{fig:boxplot_igaussian_res1_alpha}.

<<"boxplot_igaussian_res1_sigma2", fig.cap="MSEs for posterior means of $\\sigma_d^2$ in simulation 1", out.width="50%", fig.align="center", echo=FALSE>>=
res1 <- read.table("../results/simulations_igaussian_res1.csv")
set1 <- read.table("../results/simulations_igaussian_set1.csv")
names1 <- c("inv Gauss", "ind inv Gauss", "inv Gamma")
temp1 <- list(res1[, substr(colnames(res1), 1, 14)=="zeta.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 18)=="zeta.ind.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 5)=="cpost"],
              res1[, substr(colnames(res1), 1, 5)=="dpost"])
temp2 <- cbind(colMeans((2*temp1[[1]]/(set1$n + set1$p - 1) - as.numeric(set1[, substr(colnames(set1), 1, 5)=="sigma"])^2)^2),
  colMeans((2*temp1[[2]]/(set1$n - 1) - 
              as.numeric(set1[, substr(colnames(set1), 1, 5)=="sigma"])^2)^2),
               colMeans((temp1[[4]]/(temp1[[3]] - 1) - 
                           as.numeric(set1[, substr(colnames(set1), 1, 5)=="sigma"])^2)^2))
par(cex=1.3)
boxplot(temp2, outline=FALSE, names=names1, ylab="MSE")
legend("topright", title=expression(bold("median MSE")), bty="n", 
       paste(names1, round(apply(temp2, 2, median), 2), sep="="))
par(cex=1)
@

<<"boxplot_igaussian_res1_gamma2", fig.cap="a) MSE and b) correlation for posterior means of $\\gamma_d^2$ in simulation 1", fig.align="center", echo=FALSE>>=
res1 <- read.table("../results/simulations_igaussian_res1.csv")
set1 <- read.table("../results/simulations_igaussian_set1.csv")
names1 <- c("inv Gauss", "ind inv Gauss", "inv Gamma")
temp1 <- list(res1[, substr(colnames(res1), 1, 15)=="delta.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 15)=="theta.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 16)=="lambda.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 19)=="delta.ind.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 19)=="theta.ind.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 20)=="lambda.ind.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 15)=="apost.inv.Gamma"],
              res1[, substr(colnames(res1), 1, 15)=="bpost.inv.Gamma"])
temp2 <- list(as.matrix(sqrt(temp1[[1]]*temp1[[2]]^2/temp1[[3]]))*
                ratio_besselK_cpp(as.matrix(sqrt(
                  temp1[[1]]*temp1[[3]])/temp1[[2]]), set1$p),
              as.matrix(sqrt(temp1[[4]]*temp1[[5]]^2/temp1[[6]]))*
                ratio_besselK_cpp(as.matrix(sqrt(
                  temp1[[4]]*temp1[[6]])/temp1[[5]]), set1$p),
              as.matrix(temp1[[8]]/(temp1[[7]] - 1)))
temp3 <- sapply(temp2, function(m) {
  colMeans((t(m) - as.numeric(set1[, substr(colnames(set1), 1, 5)=="gamma"])^2)^2)})
temp4 <- sapply(temp2, function(m) {apply(t(m), 2, function(r) {
  cor(r, as.numeric(set1[, substr(colnames(set1), 1, 5)=="gamma"])^2)})})
par(cex=1.3)
layout(matrix(c(rep(1, 4), rep(2, 4)), 2, 4), widths=1, heights=1, respect=TRUE)
boxplot(temp3, names=names1, ylab="MSE", main="a)")
legend("right", title=expression(bold("median MSE")), bty="n", 
       paste(names1, round(apply(temp3, 2, median), 2), sep="="))
boxplot(temp4, names=names1, ylab="Correlation", main="b)")
legend("bottomleft", title=expression(bold("median correlation")), bty="n", 
       paste(names1, round(apply(temp4, 2, median), 2), sep="="))
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

<<"boxplot_igaussian_res1_mu", fig.cap="a) MSE and b) correlation for posterior means of $\\beta_d$ in simulation 1", fig.align="center", echo=FALSE>>=
res1 <- read.table("../results/simulations_igaussian_res1.csv")
names1 <- c("inv Gauss", "ind inv Gauss", "inv Gamma")
temp1 <- cbind(res1$mse.mu.inv.Gauss, res1$mse.mu.ind.inv.Gauss, 
               res1$mse.mu.inv.Gamma)
temp2 <- cbind(res1$cor.mu.inv.Gauss, res1$cor.mu.ind.inv.Gauss, 
               res1$cor.mu.inv.Gamma)
par(cex=1.3)
layout(matrix(c(rep(1, 4), rep(2, 4)), 2, 4), widths=1, heights=1, respect=TRUE)
boxplot(temp1, outline=FALSE, names=names1, ylab="MSE", main="a)")
legend("topleft", title=expression(bold("median MSE")), bty="n", 
       paste(names1, round(apply(temp1, 2, median), 2), sep="="))
boxplot(temp2, names=names1, ylab="Correlation", main="b)")
legend("bottomleft", title=expression(bold("median correlation")), bty="n", 
       paste(names1, round(apply(temp2, 2, median), 2), 
             sep="="))
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

<<"boxplot_igaussian_res1_theta", fig.cap="a) MSE and b) correlation for prior means of $\\gamma_d^2$ ($\\theta_d$) in simulation 1", fig.align="center", echo=FALSE>>=
res1 <- read.table("../results/simulations_igaussian_res1.csv")
set1 <- read.table("../results/simulations_igaussian_set1.csv")
names1 <- c("inv Gauss", "ind inv Gauss", "inv Gamma")
temp1 <- list(res1[, substr(colnames(res1), 1, 15)=="theta.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 19)=="theta.ind.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 16)=="aprior.inv.Gamma"],
              res1[, substr(colnames(res1), 1, 16)=="bprior.inv.Gamma"])
temp2 <- list(colMeans((t(temp1[[1]]) - 
                          as.numeric(set1[, substr(colnames(set1), 1, 5)=="theta"]))^2),
              colMeans((t(temp1[[2]]) - 
                          as.numeric(set1[, substr(colnames(set1), 1, 5)=="theta"]))^2),
              colMeans((t(temp1[[4]]/(temp1[[3]] - 1)) - 
                          unique(as.numeric(set1[, substr(colnames(set1), 1, 5)=="theta"])))^2))
temp3 <- list(apply(t(temp1[[1]]), 2, function(r) {
  cor(r, as.numeric(set1[, substr(colnames(set1), 1, 5)=="theta"]))}),
              apply(t(temp1[[2]]), 2, function(r) {
                cor(r, as.numeric(set1[, substr(colnames(set1), 1, 5)=="theta"]))}),
              apply(t(temp1[[4]]/(temp1[[3]] - 1)), 2, function(r) {
                cor(r, unique(as.numeric(set1[, substr(colnames(set1), 1, 5)=="theta"])))}))
par(cex=1.3)
layout(matrix(c(rep(1, 4), rep(2, 4)), 2, 4), widths=1, heights=1, respect=TRUE)
boxplot(temp2, names=names1, ylab="MSE", main="a)")
legend("bottomright", title=expression(bold("median MSE")), bty="n",
       paste(names1, round(sapply(temp2, median), 2), sep="="))
boxplot(temp3, names=names1, ylab="Correlation", main="b)")
legend("bottomleft", title=expression(bold("median correlation")), bty="n",
       paste(names1, round(sapply(temp3, median), 2), sep="="))
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

<<"boxplot_igaussian_res1_alpha", fig.cap="a) MSE and b) correlation for $\\alpha$ in simulation 1", fig.align="center", echo=FALSE>>=
res1 <- read.table("../results/simulations_igaussian_res1.csv")
set1 <- read.table("../results/simulations_igaussian_set1.csv")

names1 <- c("inv Gauss", "ind inv Gauss")
temp1 <- list(res1[, substr(colnames(res1), 1, 15)=="alpha.inv.Gauss"],
              res1[, substr(colnames(res1), 1, 19)=="alpha.ind.inv.Gauss"])
temp2 <- list(colMeans((t(temp1[[1]]) - as.numeric(set1[, substr(colnames(set1), 1, 5)=="alpha"]))^2),
              colMeans((t(temp1[[2]]) - as.numeric(set1[, substr(colnames(set1), 1, 5)=="alpha"]))^2))
temp3 <- list(apply(t(temp1[[1]]), 2, function(r) {cor(r, as.numeric(set1[, substr(colnames(set1), 1, 5)=="alpha"]))}),
              apply(t(temp1[[2]]), 2, function(r) {cor(r, as.numeric(set1[, substr(colnames(set1), 1, 5)=="alpha"]))}))
par(cex=1.3)
layout(matrix(c(rep(1, 4), rep(2, 4)), 2, 4), widths=1, heights=1, respect=TRUE)
boxplot(temp2, names=names1, ylab="MSE", main="a)")
legend("topright", title=expression(bold("median MSE")), bty="n",
       paste(names1, round(sapply(temp2, median), 2), sep="="))
boxplot(temp3, names=names1, ylab="Correlation", main="b)")
legend("bottomright", title=expression(bold("median correlation")), bty="n",
       paste(names1, round(sapply(temp3, median), 2), sep="="))
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

<<"lines_igaussian_res1_elbo", fig.cap="Convergence of ELBO for a) inverse Gaussian, b) independent inverse Gaussian, and c) inverse Gamma models in simulation 1", fig.align="center", echo=FALSE>>=
fit1 <- read.table("../results/simulations_igaussian_fit1.csv")
par(cex=1.3)
layout(matrix(c(rep(c(1, 1, 2, 2), times=2), rep(c(0, 3, 3, 0), times=2)), 
              4, 4, byrow=TRUE), widths=1, heights=1, respect=TRUE)
plot(fit1[, substr(colnames(fit1), 1, 14)=="inv.Gauss.elbo"][, 1], type="l",
     ylim=range(fit1[, substr(colnames(fit1), 1, 14)=="inv.Gauss.elbo"]),
     main="a)", xlab="Iteration", ylab="ELBO")
for(i in 2:100) {
  lines(fit1[, substr(colnames(fit1), 1, 14)=="inv.Gauss.elbo"][, i], col=i)
}
plot(fit1[, substr(colnames(fit1), 1, 18)=="ind.inv.Gauss.elbo"][, 1], type="l",
     ylim=range(fit1[, substr(colnames(fit1), 1, 18)=="ind.inv.Gauss.elbo"]),
     main="b)", xlab="Iteration", ylab="ELBO")
for(i in 2:100) {
  lines(fit1[, substr(colnames(fit1), 1, 18)=="ind.inv.Gauss.elbo"][, i], col=i)
}
plot(fit1[, substr(colnames(fit1), 1, 14)=="inv.Gamma.elbo"][, 1], type="l",
     ylim=range(fit1[, substr(colnames(fit1), 1, 14)=="inv.Gamma.elbo"]),
     main="c)", xlab="Iteration", ylab="ELBO")
for(i in 2:100) {
  lines(fit1[, substr(colnames(fit1), 1, 14)=="inv.Gamma.elbo"][, i], col=i)
}
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

<<"lines_igaussian_res1_theta", fig.cap="Convergence of prior means for a) inverse Gaussian, b) independent inverse Gaussian, and c) inverse Gamma models in simulation 1", fig.align="center", echo=FALSE>>=
fit1 <- read.table("../results/simulations_igaussian_fit1.csv")
set1 <- read.table("../results/simulations_igaussian_set1.csv")

par(cex=1.3)
layout(matrix(c(rep(c(1, 1, 2, 2), times=2), rep(c(0, 3, 3, 0), times=2)), 
              4, 4, byrow=TRUE), widths=1, heights=1, respect=TRUE)
plot(fit1[, substr(colnames(fit1), 1, 15)=="inv.Gauss.theta"][, 1], type="l",
     ylim=range(fit1[, substr(colnames(fit1), 1, 15)=="inv.Gauss.theta"]),
     main="a)", xlab="Iteration", ylab=expression(theta[d]))
for(i in 2:5) {
  lines(fit1[, substr(colnames(fit1), 1, 15)=="inv.Gauss.theta"][
    , (i - 1)*(set1$D/set1$nclass) + 1], col=i)
}
plot(fit1[, substr(colnames(fit1), 1, 19)=="ind.inv.Gauss.theta"][, 1], 
     type="l", 
     ylim=range(fit1[, substr(colnames(fit1), 1, 19)=="ind.inv.Gauss.theta"]),
     main="b)", xlab="Iteration", ylab=expression(theta[d]))
for(i in 2:5) {
  lines(fit1[, substr(colnames(fit1), 1, 19)=="ind.inv.Gauss.theta"][
    , (i - 1)*(set1$D/set1$nclass) + 1], col=i)
}
plot(sort(fit1[, substr(colnames(fit1), 1, 16)=="inv.Gamma.bprior"][, 1]/
            (fit1[, substr(colnames(fit1), 1, 16)=="inv.Gamma.aprior"][
              , 1] - 1)), 
     type="l", 
     ylim=range(fit1[, substr(colnames(fit1), 1, 16)=="inv.Gamma.bprior"]/
                  (fit1[, substr(colnames(fit1), 1, 16)=="inv.Gamma.aprior"] - 
                     1)),
     main="c)", xlab="Iteration", ylab=expression(beta/(alpha - 1)))
for(i in 2:5) {
  lines(sort((fit1[, substr(colnames(fit1), 1, 16)=="inv.Gamma.bprior"][, ]/
               (fit1[, substr(colnames(fit1), 1, 16)=="inv.Gamma.aprior"][
                 , 1] - 1))[, i]), col=i)
}
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

Simulations with no difference between classes:

<<"boxplot_igaussian_res2_theta", fig.cap="a) MSE and b) correlation for prior means of $\\gamma_d^2$ ($\\theta_d$) in simulation 2", out.width="50%", fig.align="center", echo=FALSE>>=
res2 <- read.table("../results/simulations_igaussian_res2.csv")
set2 <- read.table("../results/simulations_igaussian_set2.csv")
names1 <- c("inv Gauss", "ind inv Gauss", "inv Gamma")
temp1 <- list(res2[, substr(colnames(res2), 1, 15)=="theta.inv.Gauss"],
              res2[, substr(colnames(res2), 1, 19)=="theta.ind.inv.Gauss"],
              res2[, substr(colnames(res2), 1, 16)=="aprior.inv.Gamma"],
              res2[, substr(colnames(res2), 1, 16)=="bprior.inv.Gamma"])
temp2 <- list(colMeans((t(temp1[[1]]) - 
                          as.numeric(set2[, substr(colnames(set2), 1, 5)=="theta"]))^2),
              colMeans((t(temp1[[2]]) - 
                          as.numeric(set2[, substr(colnames(set2), 1, 5)=="theta"]))^2),
              colMeans((t(temp1[[4]]/(temp1[[3]] - 1)) - 
                          unique(as.numeric(set2[, substr(colnames(set2), 1, 5)=="theta"])))^2))
par(cex=1.3)
boxplot(temp2, names=names1, ylab="MSE", main="a)")
legend("bottomright", title=expression(bold("median MSE")), bty="n",
       paste(names1, round(sapply(temp2, median), 2), sep="="))
par(cex=1)
@

<<"lines_igaussian_res2_alpha", fig.cap="Convergence of $\\alpha$ for a) inverse Gaussian and b) independent inverse Gaussian in simulation 2", fig.align="center", echo=FALSE>>=
fit2 <- read.table("../results/simulations_igaussian_fit2.csv")
set2 <- read.table("../results/simulations_igaussian_set2.csv")

par(cex=1.3)
layout(matrix(rep(c(1, 1, 2, 2), times=2), 2, 4, byrow=TRUE), widths=1, 
       heights=1, respect=TRUE)
plot(fit2[, substr(colnames(fit2), 1, 15)=="inv.Gauss.alpha"][, 1], type="l",
     ylim=range(fit2[, substr(colnames(fit2), 1, 15)=="inv.Gauss.alpha"]),
     main="a)", xlab="Iteration", ylab=expression(theta[d]))
for(i in 2:5) {
  lines(fit2[, substr(colnames(fit2), 1, 15)=="inv.Gauss.alpha"][, i], col=i)
}
plot(fit2[, substr(colnames(fit2), 1, 19)=="ind.inv.Gauss.alpha"][, 1], 
     type="l", 
     ylim=range(fit2[, substr(colnames(fit2), 1, 19)=="ind.inv.Gauss.alpha"]),
     main="b)", xlab="Iteration", ylab=expression(theta[d]))
for(i in 2:5) {
  lines(fit2[, substr(colnames(fit2), 1, 19)=="ind.inv.Gauss.alpha"][, i], 
        col=i)
}
par(cex=1)
layout(matrix(1, 1, 1), widths=1, heights=1, respect=TRUE)
@

  \section{Starting values}
  \subsection{Method 1}
  Starting value for $\sigma_d^2$ as in \cite{chipman_bart:_2010}. They consider a scaled-Chi squared distribution for the error variance with degrees of freedom 3 and scale chosen such that the $q*100$th percentile matches the error variance in the data $s^2(\mathbf{y}_d)$. As starting value for $\sigma_d^2$ we take the mode of this distribution: $\hat{\sigma}_d^2 = 3a/5$, with $a$ the solution to:
  $$
  \Gamma\(\frac{3}{3}\) \[1 - q - F(3a/(2s^2(\mathbf{y}_d)); 3/2, 1) \] = 0,
  $$
  where $F(x; 3/2, 1)$ is the CDF of a gamma function with shape $3/2$ and scale $1$.  In \cite{moran_variance_2018}, REFERENCE to Ročková and George (2018), and \cite{chipman_bart:_2010}, $q=0.9$ is suggested. We find that $q \in ( 0.9, 0.95 )$ gives good results.

  \subsection{Method 2}
  We generate starting values as follows. Consider the $\sigma_d^2$ and $\gamma_d^2$ as fixed and estimate them by MML of the regular ridge model:
  $$
  \hat{\sigma}_d^2, \hat{\gamma}_d^2= \underset{\sigma_d^2,\gamma_d^2}{\argmax} \int_{\bm{\beta}} p(\mathbf{y}_d | \bm{\beta}, \sigma_d^2) p(\bm{\beta} | \gamma_d^2, \sigma_d^2) d\bm{\beta}.
  $$
  The regular ridge model is conjugate, so MML is relatively simple. Next we estimate a common inverse Gaussian prior for the $\gamma_d^2$ by considering the estimates to be samples from the prior:
  \begin{align*}
  \hat{\theta} & = n^{-1} \sum_{d=1}^D \hat{\gamma}_d^2 / n, \\
  \hat{\lambda} & = n \left\{\sum_{d=1}^D \[ (\hat{\gamma}_d^2)^{-1} - \hat{\theta}^{-1}\]\right\}^{-1}.
  \end{align*}
  We estimate the mode of the $\hat{m}=\hat{\gamma}_d^2$ by kernel density estimation and equate it to the theoretical mode of the generalized inverse Gaussian distribution with the estimated $\hat{\theta}$ and $\hat{\lambda}$. We solve to find a common $\delta$:
  $$
  \hat{\delta} = \frac{\hat{m}^2 \hat{\lambda}}{\hat{\theta}^2} + (p + 3) \hat{m}.
  $$
  Lastly, we consider the $\hat{\sigma}^2_d$ fixed samples from the inverse Gamma distribution to estimate a common scale:
  $$
  \hat{\zeta} = \frac{D(n + p + 1)}{2 \sum_{d=1}^D[(\hat{\sigma}_d^2)^{-1}]}.
  $$

  \section{Ratios of modified Bessel functions}
  Ratios of modified Bessel functions of the second kind $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be large, this causes numerical issues in the calculation of various quantities. We alleviate the numerical issues through the following.

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
  \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
  \begin{cases}
  \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
  \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} + \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1}, & \text{for } p \text{ odd}.
  \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, such that the ratio may be computed without numerical issues.

  \section{General model}
  We have drug covariates $\mathbf{c}_d$ and omics feature covariates $\mathbf{z}_j$. The model is now:
  \begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \tau_j^2, \gamma_d^2, \sigma_d^2 & \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2 \tau_j^2), \\
  \gamma_d^2 & \sim \mathcal{IG}\((\mathbf{c}_d \tr \bm{\alpha})^{-1}, \lambda\), \\
  \tau_j^2 & \sim \mathcal{IG}\((\mathbf{z}_j \tr \bm{\theta})^{-1}, \nu\), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}
  Variational posterior:
  \begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+1}{2}, \lambda (\mathbf{c}_d \tr \bm{\alpha})^2, \delta_d\), \\
  q(\bm{\tau}^2) & \overset{D}{=} \prod_{j=1}^p \mathcal{GIG}\(-\frac{D+1}{2}, \nu (\mathbf{z}_j \tr \bm{\theta})^2, \eta_j\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
  Variational parameters:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = \frac{2 \zeta_d^{(h)}}{n+p+1} \[\X \tr \X + a_d^{(h)} \cdot \diag(b_j^{(h)})\]^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + a_d^{(h)} \cdot \diag(b_j^{(h)})\]^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \frac{n+p+1}{2 \zeta_d^{(h)}} \left\{\trace \[\diag(b_j^{(h)}) \bm{\Sigma}_d^{(h+1)} \] + (\bm{\mu}_d^{(h+1)}) \tr \diag(b_j^{(h)}) \bm{\mu}_d^{(h+1)}\right\} + \lambda, \\ 
  \eta_j^{(h+1)} & = \sum_{d=1}^D \frac{a_d^{(h)}(n+p+1)}{2 \zeta_d^{(h)}} \[ (\bmu_{jd}^{(h + 1)})^2 + (\bSigma_d^{(h + 1)})_{jj} \] + \nu, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \Big( \mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + a_d^{(h+1)} \trace \left\{ \[\X \tr \X + \diag(b_j^{(h + 1)})\] \bm{\Sigma}_d^{(h+1)}\right\} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + a_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \[\X \tr \X + \diag(b_j^{(h + 1)})\]\bm{\mu}_d^{(h+1)}\Big),
  \end{align*}
  with
  \begin{align*}
  a_d^{(h)} & = \sqrt{\frac{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2}{\delta_d^{(h)}}} \frac{K_{\frac{p-1}{2}} \( \sqrt{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2 \delta_d^{(h)}} \)}{K_{\frac{p+1}{2}} \( \sqrt{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2 \delta_d^{(h)}} \)} + \frac{p+1}{\delta_d^{(h)}} \text{ and} \\
  b_j^{(h)} & = \sqrt{\frac{\nu (\mathbf{z}_j \tr \bm{\theta})^2}{\eta_j^{(h)}}} \frac{K_{\frac{D-1}{2}} \( \sqrt{\nu (\mathbf{z}_j \tr \bm{\theta})^2 \eta_j^{(h)}} \)}{K_{\frac{D+1}{2}} \( \sqrt{\nu (\mathbf{z}_j \tr \bm{\theta})^2 \eta_j^{(h)}} \)} + \frac{D+1}{\eta_j^{(h)}}.
  \end{align*}
  Hyper-parameter updates:
  \begin{align*}
  \bm{\alpha}^{(l+1)} & = \left[ \mathbf{C} \tr \diag ( e_d^{(l)} ) \mathbf{C} \right]^{-1} \mathbf{C} \tr \mathbf{1}, \\
  \lambda^{(l+1)} & = \[ \sum_{d=1}^D a_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D, \\
  \bm{\theta}^{(l+1)} & = \left[ \mathbf{Z} \tr \diag ( f_j^{(l)} ) \mathbf{Z} \right]^{-1} \mathbf{Z} \tr \mathbf{1}, \\
  \nu^{(l+1)} & = \[ \sum_{j=1}^p b_j^{(l)} - (\bm{\theta}^{(l+1)}) \tr \mathbf{Z} \tr \mathbf{1} \]^{-1} p,
  \end{align*}
  where 
  \begin{align*}
  e_d^{(l)} & = \sqrt{\frac{\delta_d^{(l)}}{\lambda^{(l)}(\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda^{(l)} (\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2 \delta_d^{(l)}} \)}{K_{\frac{p + 1}{2}} \( \sqrt{\lambda^{(l)} (\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2 \delta_d^{(l)}} \)} \text{ and} \\
  f_j^{(l)} & = \sqrt{\frac{\eta_j^{(l)}}{\nu^{(l)}(\mathbf{z}_j \tr \bm{\theta}^{(l)})^2}} \frac{K_{\frac{D - 1}{2}} \( \sqrt{\nu^{(l)} (\mathbf{z}_j \tr \bm{\theta}^{(l)})^2 \eta_j^{(l)}} \)}{K_{\frac{D + 1}{2}} \( \sqrt{\nu^{(l)} (\mathbf{z}_j \tr \bm{\theta}^{(l)})^2 \eta_j^{(l)}} \)}.
  \end{align*}

  \bibliographystyle{author_short3} 
	\bibliography{refs}

\end{document}