### 2019-01-10 docs/supplement.Rnw
simulation models now same framework, partly implemented in R

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\bsigma}{\bm{\sigma}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bepsilon}{\bm{\epsilon}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bPsi}{\bm{\Psi}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bgamma}{\bm{\gamma}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\Em}{\mathbf{E}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\Lm}{\mathbf{L}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Cov}{\mathbb{C}\text{ov}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

% settings
\pgfplotsset{compat=1.13}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Supplementary material to: Drug efficacy prediction in cell lines}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

\begin{document}

% load packages and set options
<<include=FALSE>>=
library(cambridge)
opts_knit$set(base.dir="../figs")
opts_chunk$set(fig.align='center', echo=FALSE, fig.path="../figs/")
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("../code/figures.R")
@

	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge, United Kingdom \\

	\section{Content overview}
	This document contains the Supplementary Material (SM) to the document `Drug efficacy prediction in cell lines'. In the following, this document is referred to as Main Document (MD).

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	In the following all expectations are with respect to the variational posterior $Q$.
	\begin{align*}
	\log q_{\B} (\B) & \propto \E_{\mathbf{U}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\bgamma^2, \bsigma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\mathbf{u}_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E_{\gamma^2_d, \sigma^2_d} \( \frac{\beta_{jd}^2}{\gamma_d^2 \sigma_d^2} \), \\
	q_{\B} (\B) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	& \text{ with } \bSigma_d = \E(\sigma_d^{-2})^{-1} [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1}, \\
	& \text{ and } \bmu_d = [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1} \X \tr [\y_d - \Z \cdot \E(\mathbf{u}_d)].
	\end{align*}

	\begin{align*}
	\log q_{\U}(\U) & \propto \E_{\mathbf{B}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2} [\log \pi (\U | \mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\bbeta_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \E_{\bm{\Xi}_d, \sigma_d^2}\( \frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \), \\
	q_{\U}(\U) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_T (\mathbf{m}_d, \Sm_d), \\
	& \text{ with } \Sm_d = \E(\sigma_d^{-2})^{-1} [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1}, \\
	& \text{ and } \mathbf{m}_d = [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1} \Z \tr [\y_d - \X \cdot \E(\bbeta_d)].
	\end{align*}

	\begin{align*}
	\log q_{\bgamma^{2}}(\bgamma^{2}) & \propto \E_{\B, \bsigma^2} [\log \pi (\B | \bgamma^{2}, \sigma^2)] + \log \pi (\bgamma^{2}) \\
	& \propto -\frac{p}{2} \sum_{d=1}^D \log \gamma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} - \frac{3}{2} \sum_{d=1}^D \log \gamma_d^{2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{\lambda}{2 \theta^2} \sum_{d=1}^D (\gamma_d^2 - \theta)^2 \gamma_d^{-2} \\
	& \propto \sum_{d=1}^D \left\{ \( -\frac{p + 1}{2} - 1\) \log \gamma_d^{2} - \frac{\lambda}{2\theta^2} \gamma_d^2 - \frac{1}{2}\[ \lambda + \E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^{2}}\] \gamma_d^{-2} \right\}, \\
	q_{\bgamma^{2}}(\bgamma^{2}) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG} \(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_{d} \), \\
	& \text{with } \delta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} + \lambda.
	\end{align*}

	\begin{align*}
	\log q_{\bsigma^{2}}(\bsigma^{2}) & \propto \E_{\B,\mathbf{U}} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\B, \bgamma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \,\,\,\,\,\,\,\,\,\, + \E_{\mathbf{U}, \bm{\Xi}_1, \dots, \bm{\Xi}_D} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bsigma^{2}) \\
	& \propto -\frac{n}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \sigma_d^{2} - \frac{p}{2} \sum_{d=1}^D \log \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \E( \gamma_d^{-2}) \frac{\E ( \beta_{jd}^2)}{\phi_{g(j)}^{2}} \sigma_d^{-2} - \frac{T}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \E(\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}) \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{3}{2} \sum_{d=1}^D \log \sigma^2_d \\
	& = -\sum_{d=1}^{D} \(\frac{n + p + T + 3}{2} \) \log \sigma_d^{2} - \sum_{d=1}^{D} \frac{1}{2}\Bigg\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \\
	& \,\,\,\,\,\,\,\,\,\, + \E(\gamma_d^{-2})\sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^2} + \E(\mathbf{u}_{d} \tr \bm{\Xi}_d^{-1} \mathbf{u}_d) \Bigg\} \sigma_d^{-2}, \\
	q_{\bsigma^{2}}(\bsigma^{2}) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + T + 1}{2}, \zeta_{d} \), \\
	& \text{ with } \zeta_{d} = \frac{\y_d \tr \y_d}{2} - \y_d \tr [\X \E(\bbeta_d) + \Z \E(\mathbf{u}_d)] + \E(\bbeta_d \tr) \X \tr \Z \E (\mathbf{u}_d) \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\trace [\X \tr \X \V(\bbeta_d)]}{2} + \frac{\E(\bbeta_d \tr) \X \tr \X \E(\bbeta_d)}{2} + \frac{\trace [\Z \tr \Z \V(\mathbf{u}_d)]}{2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\E(\mathbf{u}_d \tr) \Z \tr \Z \E(\mathbf{u}_d)}{2} + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+ \frac{\E(\mathbf{u}_d \tr) \E(\bm{\Xi}_d^{-1}) \E(\mathbf{u}_d)}{2} + \frac{\trace [\E(\bm{\Xi}_d^{-1}) \V(\mathbf{u}_d)]}{2},
	\end{align*}

	\begin{align*}
	\log q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \propto \E_{\mathbf{U}, \bm{\sigma}^2} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D) \\
	& \propto -\frac{1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \E \(\frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \) \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{k + T + 1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace (\bm{\Omega} \bm{\Xi}_d^{-1}) \\
	& = -\frac{\tau + T + 2}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace \left\{ \[\E(\sigma_d^{-2}) \E( \mathbf{u}_d \mathbf{u}_d \tr) + \bm{\Omega}\] \bm{\Xi}_d^{-1} \right\}, \\
	q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \overset{D}{\propto} \prod_{d=1}^D \mathcal{W}_T^{-1} (\bm{\Psi}_d, \nu + 1),  \\ 
	& \text{ with } \bm{\Psi}_d =\E(\sigma^{-2}_d) \[\V(\mathbf{u}_d) + \E(\mathbf{u}_d) \E( \mathbf{u}_d \tr) \] + \bm{\Omega}.
	\end{align*}

	\section{Simulations}
	\subsection{Models}
	In order to investigate the inverse Gaussian prior model and its estimation, we consider a simpler model. The simpler model ignores tissue effects and fixes $\forall g: \phi_g^2=1$:
	\begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim p(\beta_{jd}), \\
  \gamma_d^2 & \sim p(\gamma_d^2), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ and $\sigma_d^2$ should be independent \textit{a priori}. However, in our case that is debatable. The $D$ regressions need not be on the same scale, so some form of calibration of the $\bm{\beta}_d$ posteriors to this scale is required. A simple way of doing this is to include the scales $\sigma_d^2$ in the priors for the $\bm{\beta}_d$. We call the two versions of the model conjugate (with $\bm{\beta}_d$ and $\sigma_d^2$ dependent):
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2)
  $$
  and non-conjugate (with $\bm{\beta}_d$ and $\sigma_d^2$ independent):
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | \gamma_d^2 \sim \mathcal{N} (0, \gamma_d^2),
  $$
  and compare them in our simulations.

	In addition to the inverse Gaussian model for the prior variances $\gamma_d^2$ (introduced in the MD):
	$$
  \gamma_d^2 \sim \mathcal{IG}(\theta_d, \lambda_d),
  $$
	we investigate a generalisation of the models in \cite{kpogbezan_empirical_2017} and \cite{leday_gene_2017}, where we draw the $\gamma_d^2$ from an inverse Gamma distribution; :
	$$
	\gamma_d^2 \sim \Gamma^{-1}(\alpha_d/2, \lambda_d/2).
	$$
	For the sake of comparability, we have rescaled the inverse Gamma parameters and reparametrised both distributions as generalized inverse Gaussians:
	\begin{align*}
  \gamma_d^2 & \sim \mathcal{GIG}(-1/2, \lambda_d \theta_d^{-2}, \lambda_d), \\
  \gamma_d^2 & \sim \mathcal{GIG}(-\alpha_d/2, 0, \lambda_d).
  \end{align*}
	The two models coincide if $\theta_d \to \infty$ and $\alpha_d = 1$.

	\subsection{Estimation}
	The variational distributions are as follows:
	\begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+\alpha_d}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{df + 1}{2}, \zeta_d\),
  \end{align*}
	where $\alpha_d=1$ and $\theta_d \to \infty$ in the inverse Gaussian and inverse Gamma models, respectively. Furthermore, $df=n + p$ in the conjugate model and $df=n$ in the non-conjugate model. The variational parameters contain cyclic dependencies and are, in the conjugate setting, iteratively updated by:
	\begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + b_d^{(h)} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = a_d^{(h)} \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + b_d^{(h+1)} \trace ( \bm{\Sigma}_d^{(h+1)}) + b_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg].
  \end{align*}
	while in the non-conjugate setting, we use the following updating equations:
	\begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + b_d^{(h)} (a_d^{(h)})^{-1} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + b_d^{(h)} (a_d^{(h)})^{-1} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \].
  \end{align*}
  Here, we set
  \begin{align}
  a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(df+1)/(2 \zeta_d^{(h)}), \nonumber \\
  b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}{\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p + \alpha_d -2}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}{K_{\frac{p+\alpha_d}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + \frac{p+\alpha_d^{(h)}}{\delta_d^{(h)}}. \label{eq:auxvar1}
  \end{align}
  Note that in the inverse Gamma model, the computation of (\ref{eq:auxvar1}) simplifies to $b_d^{(h)}=(p+\alpha_d)/\delta_d^{(h)}$.

  \subsection{Empirical Bayes}
  To avoid subjectivity in our choice of prior, we estimate the hyper-parameters with empirical Bayes. The empirical Bayes procedure differs for the two prior models. In the inverse Gaussian case, we parametrise $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ to allow for the inclusion of drug covariates in the empirical Bayes estimation. The estimating equations are:
  \begin{align*}
  \bm{\alpha} & = \[ \mathbf{C} \tr \diag (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \diag (\lambda_d) \mathbf{1}_{D \times 1}, \\
  \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = (b_d^{(l)} + \delta_d^{(l)}/p) \cdot \delta_d^{(l)} (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$. If we use one common $\lambda=\lambda_d$ then the solution is in closed form:
  \begin{align*}
	\bm{\alpha}^{(l+1)} & = \[ \mathbf{C} \tr \diag(e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \mathbf{1}_{D \times 1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D b_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D.
	\end{align*}
	This corresponds to the model in the MD.

  In the inverse Gamma model we use the following empirical Bayes procedure: Suppose we have a partitioning of our drugs into drug classes and we let prior variance components $\gamma_d^2$ and $\gamma^2_{d'}$ be drawn from the same distribution if drug $d$ and drug $d'$ are from the same class. Or simply put: $\gamma^2_{d}, \gamma^2_{d'} \sim \Gamma^{-1}(\alpha_c/2, \lambda_c/2)$ if $d, d' \in \text{class}_c$. Empirical bayes estimation of the $\alpha_c$ and $\lambda_c$ by maximisation of the marginal likelihood amounts to iteratively solving the estimating equations:
  \begin{subequations}
    \begin{align}
      \log \(\frac{\lambda_c}{2}\) - \psi\(\frac{\alpha_c}{2}\) - \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} & = 0, \label{eq:estequation1}\\
      \frac{\alpha_c}{\lambda_c} - \frac{\sum_{d \in \text{class}_c} b_d^{(l)}}{|\text{class}_c|} & = 0, \label{eq:estequation2}
    \end{align}
  \end{subequations}
  where $e^{(l)}_d = \log \lambda_d^{(l)} - \psi ( \alpha_d^{(l)}/2) - \log 2$. We solve (\ref{eq:estequation2}) for $\lambda_c$, plug it into (\ref{eq:estequation1}) and bound the result using \cite{alzer_inequalities_1997} to obtain an interval $( a_c^*, 2a_c^* )$ containing the solution to (\ref{eq:estequation1}), where:
  $$
  a_c^* = \left\{ \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} + \log \[ \frac{\sum_{d \in \text{class}_c} b_d^{(l)}}{|\text{class}_c|} \]  \right\}^{-1}.
  $$
  With this interval it is straightforward to find $\alpha^{(l+1)}_c$ by any root-finding algorithm and plugging the solution into (\ref{eq:estequation2}) to find $\lambda^{(l+1)}_c$.

	The inverse Gamma model with independent drug classes allows for more flexible empirical Bayes estimation of the prior means for the different drug classes. As a consequence however, the risk of overfitting is increased. Additionally, it does not allow to include continuous covariates on the drugs.

	\subsection{Evidence lower bound}
	We monitor convergence through the evidence lower bound (ELBO). To compute the evidence lower bound after updating the EB parameters from $\alpha_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}$ to $\alpha_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$, we compute $\text{ELBO}^{(l + 1)} = \sum_{d=1}^D \text{ELBO}_d^{(l + 1)}$, with, in the conjugate case:
	\begin{align*}
	\text{ELBO}_d^{(l + 1)} & = \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	& - \frac{df + 1}{2} \log \zeta_d^{(l)} + \frac{p + \alpha_d^{(l)}}{4} \log \lambda_d^{(l)} - \frac{p + \alpha_d^{(l)}}{2} \log \theta_d^{(l)} - \frac{p+\alpha_d^{(l)}}{4} \log \delta_d^{(l)} \\
	& + \log K_{\frac{p+\alpha_d^{(l)}}{2}}\(\sqrt{\lambda_d^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}\) + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - a_d^{(l)} (\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - a_d^{(l)} \trace (\bm{\Sigma}_d^{(l)}) \] \\
	& + \frac{\lambda_d^{(l)} e_d^{(l)}}{2(\theta_d^{(l)})^2} + \frac{(\alpha_d^{(l)} - \alpha_d^{(l + 1)})h_d^{(l)}}{2} - \log K_{\frac{\alpha_d^{(l + 1)}}{2}}\(\lambda_d^{(l + 1)}/\theta_d^{(l + 1)}\) + \frac{\alpha_d^{(l + 1)}}{2} \log \theta_d^{(l)} \\
	& - \frac{\lambda_d^{(l+1)} e_d^{(l)}}{2 (\theta_d^{(l + 1)})^2} - \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2},
	\end{align*}
	where $h_d^{(l)} = \log \delta_d^{(l)} - \psi [(p + \alpha_d^{(l)})/2] - \log 2$ in the inverse Gamma model and
	$$
	h_d^{(l)} = \frac{1}{2}\log \delta_d^{(l)} - \frac{1}{2}\log \lambda_d^{(l)} + \log \theta_d^{(l)} + \frac{\partial}{\partial \nu} \log K_{\nu} (\sqrt{\lambda_d^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}) \bigg|_{\nu = \frac{p + \alpha_d^{(l)}}{2}}
	$$
	in the inverse Gaussian case. Note that in the inverse Gaussian case, the computationally expensive $h_d^{(l)}$ terms never have to be calculated, since we have $\alpha_d^{(l)} - \alpha_d^{(l + 1)}=0$. The non-conjugate ELBO simply exchanges the $a_d^{(l)} (\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)}$ and $a_d^{(l)} \trace (\bm{\Sigma}_d^{(l)})$ terms with $(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)}$ and $\trace (\bm{\Sigma}_d^{(l)})$, respectively. To compute the ELBO after a VB update we simply let $\alpha_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}=\alpha_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$.

  In practice, there are different options to assess convergence. We may monitor the parameters themselves, the ELBO, or a combination of the two. We may also iterate for an \textit{a priori} fixed number of times. We may also choose to use a different convergence criterium for the EB and VB iterations. Currently we fix the number of EB iterations to 20, while we do 2 VB iterations per EB iteration.

  \subsection{Efficient computation}
  The empirical Bayes updates require the following quantities: $\zeta_d$, $a_d$, $\delta_d$, $\trace (\bm{\Sigma}_d)$, $\trace (\X \tr \X \bm{\Sigma}_d)$, $\bm{\mu}_d \tr \bm{\mu}_d$, and $\bm{\mu}_d \tr \X \tr \X \bm{\mu}_d$. In addition, we need $\log|\bm{\Sigma}|$ and $\mathbf{y}_d \tr \X \bm{\mu}_d$ to monitor the ELBO. The first three quantities are obtained by scalar operations of $\mathcal{O}(n)$. Let $c_d^{(h)}=b_d^{(h)}$ in the conjugate setting and $c_d^{(h)}=b_d^{(h)}/a_d^{(h)}$ in the non-conjugate setting. Then, beforementioned quantities are easily calculated using the SVD $\X = \mathbf{U} \mathbf{D} \mathbf{V} \tr$:
  \begin{align*}
  \trace (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} \[\sum_{i=1}^n (v_i^2 + c_d^{(h)})^{-1} + \max (p - n, 0)\cdot (c_d^{(h)})^{-1}\], \\
  \trace (\X \tr \X \bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} \sum_{i=1}^n v_i^2(v_i^2 + c_d^{(h)})^{-1}, \\
  (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)})^2, \\
  (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^4 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)})^2, \\
  \log|\bm{\Sigma}^{(h+1)}| & \propto -p \log a_d^{(h)} - \sum_{i=1} \log (v_i^2 + c_d^{(h)}) - \max(p - n, 0) \cdot \log c_d^{(h)}, \\
  \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)}).
  \end{align*}
  where $v_i$ are the singular values of $\X$. The SVD and $\mathbf{U} \tr \mathbf{y}_d$ are $\mathcal{O}(pn^2)$ and $\mathcal{O}(n^2)$ operations, respectively, but have to be calculated only once at the start of the algorithm. The rest of the calculations involve just $n$ scalar multiplications and/or additions.

  \section{Simulation setting}
  We simulate from a correctly specified model given in \ref{sec:notissuemodel}. We assume random $y_{id}$, $\x_i$, and $\bm{\beta}_d$, while we fix $\mathbf{C}$, $\gamma_d^2$, $\sigma_d^2$ and $\bm{\alpha}$. We estimate the $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, $\bm{\alpha}$ and $\lambda$, while we fix the $\sigma_d^2$ to their true values.

  We set $n=100$, $p=200$, and $D=10$. We consider 5 evenly sized classes of drugs, such that $\mathbf{c}_{kd} = \mathbbm{1}(k = \text{class}_d)$ and $\begin{bmatrix} \mathbf{c}_1 & \cdots & \mathbf{c}_D \end{bmatrix} \tr$. From these the $\gamma_d^2 = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ are created. Next, we simulate $\bm{\beta}_{jd} \sim \mathcal{N} (0, \sigma^2_d \gamma_d^2)$. We fix the mean signal-to noise ratio $\overline{\text{SNR}}_d$, such that the predictor data variance is $s^2 = \V(\mathbf{x}_i) = \overline{\text{SNR}}_d / (\overline{\gamma^2_d} p)$ (NOT CORRECT). The last step consists of simulating the data by $\mathbf{x}_{ij} \sim \mathcal{N}(0,s^2)$ and $y_{id} = \mathcal{N} (\mathbf{x}_i \tr \bm{\beta}_d, \sigma_d^2)$. The remaining parameters are set as follows: $\bm{\alpha} = \begin{bmatrix} 1 & \dots & 5 \end{bmatrix} \tr$, $\sigma_d^2 = 1$ for all $d$, and $\overline{\text{SNR}}_d = 10$. To account for the random data and parameter generation, we repeat the simulation 100 times. We compare the results to the models in \ref{sec:indvarmodel} and \ref{sec:invgammamodel}, and present the results in Figures \ref{fig:boxplot_igaussian_res1_sigma2}-\ref{fig:boxplot_igaussian_res1_alpha}.

<<boxplot_igaussian_res1_sigma2, cache=FALSE, echo=FALSE, fig.cap="MSEs for posterior means of $\\sigma_d^2$ in simulation 1", out.width="50%", fig.align="center">>=
@

<<boxplot_igaussian_res1_gamma2, cache=FALSE, echo=FALSE, fig.cap="a) MSE and b) correlation for posterior means of $\\gamma_d^2$ in simulation 1", fig.align="center">>=
@

<<boxplot_igaussian_res1_mu, cache=FALSE, echo=FALSE, fig.cap="a) MSE and b) correlation for posterior means of $\\beta_d$ in simulation 1", fig.align="center">>=
@

<<boxplot_igaussian_res1_theta, cache=FALSE, echo=FALSE, fig.cap="a) MSE and b) correlation for prior means of $\\gamma_d^2$ ($\\theta_d$) in simulation 1", fig.align="center">>=
@

<<boxplot_igaussian_res1_alpha, cache=FALSE, echo=FALSE, fig.cap="a) MSE and b) correlation for $\\alpha$ in simulation 1", fig.align="center">>=
@

<<lines_igaussian_res1_elbo, cache=FALSE, echo=FALSE, fig.cap="Convergence of ELBO for a) inverse Gaussian, b) independent inverse Gaussian, and c) inverse Gamma models in simulation 1", fig.align="center">>=
@

<<lines_igaussian_res1_theta, cache=FALSE, echo=FALSE, fig.cap="Convergence of prior means for a) inverse Gaussian, b) independent inverse Gaussian, and c) inverse Gamma models in simulation 1", fig.align="center">>=
@

Simulations with no difference between classes:

<<boxplot_igaussian_res2_theta, cache=FALSE, echo=FALSE, fig.cap="a) MSE and b) correlation for prior means of $\\gamma_d^2$ ($\\theta_d$) in simulation 2", out.width="50%", fig.align="center">>=
@

<<lines_igaussian_res2_alpha, cache=FALSE, echo=FALSE, fig.cap="Convergence of $\\alpha$ for a) inverse Gaussian and b) independent inverse Gaussian in simulation 2", fig.align="center">>=
@

  \section{Starting values}
  \subsection{Method 1}
  Starting value for $\sigma_d^2$ as in \cite{chipman_bart:_2010}. They consider a scaled-Chi squared distribution for the error variance with degrees of freedom 3 and scale chosen such that the $q*100$th percentile matches the error variance in the data $s^2(\mathbf{y}_d)$. As starting value for $\sigma_d^2$ we take the mode of this distribution: $\hat{\sigma}_d^2 = 3a/5$, with $a$ the solution to:
  $$
  \Gamma\(\frac{3}{3}\) \[1 - q - F(3a/(2s^2(\mathbf{y}_d)); 3/2, 1) \] = 0,
  $$
  where $F(x; 3/2, 1)$ is the CDF of a gamma function with shape $3/2$ and scale $1$.  In \cite{moran_variance_2018}, REFERENCE to Ročková and George (2018), and \cite{chipman_bart:_2010}, $q=0.9$ is suggested. We find that $q \in ( 0.9, 0.95 )$ gives good results.

  \subsection{Method 2}
  We generate starting values as follows. Consider the $\sigma_d^2$ and $\gamma_d^2$ as fixed and estimate them by MML of the regular ridge model:
  $$
  \hat{\sigma}_d^2, \hat{\gamma}_d^2= \underset{\sigma_d^2,\gamma_d^2}{\argmax} \int_{\bm{\beta}} p(\mathbf{y}_d | \bm{\beta}, \sigma_d^2) p(\bm{\beta} | \gamma_d^2, \sigma_d^2) d\bm{\beta}.
  $$
  The regular ridge model is conjugate, so MML is relatively simple. Next we estimate a common inverse Gaussian prior for the $\gamma_d^2$ by considering the estimates to be samples from the prior:
  \begin{align*}
  \hat{\theta} & = n^{-1} \sum_{d=1}^D \hat{\gamma}_d^2 / n, \\
  \hat{\lambda} & = n \left\{\sum_{d=1}^D \[ (\hat{\gamma}_d^2)^{-1} - \hat{\theta}^{-1}\]\right\}^{-1}.
  \end{align*}
  We estimate the mode of the $\hat{m}=\hat{\gamma}_d^2$ by kernel density estimation and equate it to the theoretical mode of the generalized inverse Gaussian distribution with the estimated $\hat{\theta}$ and $\hat{\lambda}$. We solve to find a common $\delta$:
  $$
  \hat{\delta} = \frac{\hat{m}^2 \hat{\lambda}}{\hat{\theta}^2} + (p + 3) \hat{m}.
  $$
  Lastly, we consider the $\hat{\sigma}^2_d$ fixed samples from the inverse Gamma distribution to estimate a common scale:
  $$
  \hat{\zeta} = \frac{D(n + p + 1)}{2 \sum_{d=1}^D[(\hat{\sigma}_d^2)^{-1}]}.
  $$

  \section{Ratios of modified Bessel functions}
  Ratios of modified Bessel functions of the second kind $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be large, this causes numerical issues in the calculation of various quantities. We alleviate the numerical issues through the following.

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
  \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
  \begin{cases}
  \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
  \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} + \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1}, & \text{for } p \text{ odd}.
  \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, such that the ratio may be computed without numerical issues.

  \section{General model}
  We have drug covariates $\mathbf{c}_d$ and omics feature covariates $\mathbf{z}_j$. The model is now:
  \begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \tau_j^2, \gamma_d^2, \sigma_d^2 & \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2 \tau_j^2), \\
  \gamma_d^2 & \sim \mathcal{IG}\((\mathbf{c}_d \tr \bm{\alpha})^{-1}, \lambda\), \\
  \tau_j^2 & \sim \mathcal{IG}\((\mathbf{z}_j \tr \bm{\theta})^{-1}, \nu\), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}
  Variational posterior:
  \begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+1}{2}, \lambda (\mathbf{c}_d \tr \bm{\alpha})^2, \delta_d\), \\
  q(\bm{\tau}^2) & \overset{D}{=} \prod_{j=1}^p \mathcal{GIG}\(-\frac{D+1}{2}, \nu (\mathbf{z}_j \tr \bm{\theta})^2, \eta_j\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
  Variational parameters:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = \frac{2 \zeta_d^{(h)}}{n+p+1} \[\X \tr \X + a_d^{(h)} \cdot \diag(b_j^{(h)})\]^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + a_d^{(h)} \cdot \diag(b_j^{(h)})\]^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \frac{n+p+1}{2 \zeta_d^{(h)}} \left\{\trace \[\diag(b_j^{(h)}) \bm{\Sigma}_d^{(h+1)} \] + (\bm{\mu}_d^{(h+1)}) \tr \diag(b_j^{(h)}) \bm{\mu}_d^{(h+1)}\right\} + \lambda, \\ 
  \eta_j^{(h+1)} & = \sum_{d=1}^D \frac{a_d^{(h)}(n+p+1)}{2 \zeta_d^{(h)}} \[ (\bmu_{jd}^{(h + 1)})^2 + (\bSigma_d^{(h + 1)})_{jj} \] + \nu, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \Big( \mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + a_d^{(h+1)} \trace \left\{ \[\X \tr \X + \diag(b_j^{(h + 1)})\] \bm{\Sigma}_d^{(h+1)}\right\} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + a_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \[\X \tr \X + \diag(b_j^{(h + 1)})\]\bm{\mu}_d^{(h+1)}\Big),
  \end{align*}
  with
  \begin{align*}
  a_d^{(h)} & = \sqrt{\frac{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2}{\delta_d^{(h)}}} \frac{K_{\frac{p-1}{2}} \( \sqrt{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2 \delta_d^{(h)}} \)}{K_{\frac{p+1}{2}} \( \sqrt{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2 \delta_d^{(h)}} \)} + \frac{p+1}{\delta_d^{(h)}} \text{ and} \\
  b_j^{(h)} & = \sqrt{\frac{\nu (\mathbf{z}_j \tr \bm{\theta})^2}{\eta_j^{(h)}}} \frac{K_{\frac{D-1}{2}} \( \sqrt{\nu (\mathbf{z}_j \tr \bm{\theta})^2 \eta_j^{(h)}} \)}{K_{\frac{D+1}{2}} \( \sqrt{\nu (\mathbf{z}_j \tr \bm{\theta})^2 \eta_j^{(h)}} \)} + \frac{D+1}{\eta_j^{(h)}}.
  \end{align*}
  Hyper-parameter updates:
  \begin{align*}
  \bm{\alpha}^{(l+1)} & = \left[ \mathbf{C} \tr \diag ( e_d^{(l)} ) \mathbf{C} \right]^{-1} \mathbf{C} \tr \mathbf{1}, \\
  \lambda^{(l+1)} & = \[ \sum_{d=1}^D a_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D, \\
  \bm{\theta}^{(l+1)} & = \left[ \mathbf{Z} \tr \diag ( f_j^{(l)} ) \mathbf{Z} \right]^{-1} \mathbf{Z} \tr \mathbf{1}, \\
  \nu^{(l+1)} & = \[ \sum_{j=1}^p b_j^{(l)} - (\bm{\theta}^{(l+1)}) \tr \mathbf{Z} \tr \mathbf{1} \]^{-1} p,
  \end{align*}
  where 
  \begin{align*}
  e_d^{(l)} & = \sqrt{\frac{\delta_d^{(l)}}{\lambda^{(l)}(\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda^{(l)} (\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2 \delta_d^{(l)}} \)}{K_{\frac{p + 1}{2}} \( \sqrt{\lambda^{(l)} (\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2 \delta_d^{(l)}} \)} \text{ and} \\
  f_j^{(l)} & = \sqrt{\frac{\eta_j^{(l)}}{\nu^{(l)}(\mathbf{z}_j \tr \bm{\theta}^{(l)})^2}} \frac{K_{\frac{D - 1}{2}} \( \sqrt{\nu^{(l)} (\mathbf{z}_j \tr \bm{\theta}^{(l)})^2 \eta_j^{(l)}} \)}{K_{\frac{D + 1}{2}} \( \sqrt{\nu^{(l)} (\mathbf{z}_j \tr \bm{\theta}^{(l)})^2 \eta_j^{(l)}} \)}.
  \end{align*}

  \bibliographystyle{author_short3}  
	\bibliography{refs}

\end{document}