### 2019-08-13 docs/supplement2.Rnw

% load packages
<<packages, include=FALSE, echo=FALSE>>=
library(cambridge)
@

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("code/figures.R")
@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip,xr-hyper,hyperref}
\externaldocument[md-]{manuscript2}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\bsigma}{\bm{\sigma}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bepsilon}{\bm{\epsilon}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bPsi}{\bm{\Psi}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bgamma}{\bm{\gamma}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\Em}{\mathbf{E}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\Lm}{\mathbf{L}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Cov}{\mathbb{C}\text{ov}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

% settings
\pgfplotsset{compat=1.13}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Supplementary material to: Drug efficacy prediction in cell lines}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

\begin{document}

	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge, United Kingdom \\

	\section{Content overview}
	This document contains the Supplementary Material (SM) to the document `Drug efficacy prediction in cell lines'. In the following, this document is referred to as Main Document (MD).

	\section{Model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	We have continuous efficacy measures $y_{id}$ for cell lines $i=1,\dots, n$, 
	from tissues $t=1, \dots, T$, on drugs $d=1,\dots,D$. Throughout this paper 
	we assume the $y_{id}$ to be centred per drug and let $\y_d = \begin{bmatrix} 
	y_{1d} & \dots & y_{nd} \end{bmatrix} \tr$. We predict efficacy with molecular
	features $x_{ij}$, $j=1,\dots, p$, collected in 
	$\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} \end{bmatrix} \tr$. We have an
	\textit{a priori} partitioning of the omics features into $g=1, \dots, G$ 
	non-overlapping groups, which might be informative for the problem at hand For convenience, we let $g(j)$ denote the group index of molecular feature $j$. Furthermore, we have characteristics of the drugs available in the form of `co-data' $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & \mathbf{c}_D \end{bmatrix} \tr$. We code the tissue from which the cell lines were taken as a $T$ binary dummy variables 
	$$
	z_{it} = \begin{cases}
	1 & \text{if } \text{tissue}_i = t, \\
	0 & \text{otherwise},
	\end{cases}
	$$
	collected in $\mathbf{z}_i = \begin{bmatrix} z_{i1} & \dots & z_{iT} \end{bmatrix} \tr$.

	We assume that cell lines from the same tissue are related and model this relation through random effects $u_{td}$. We let the molecular feature effects $\beta_{jd}$ vary over the drugs and model drug efficacy as a linear combination of these two effects:
	\begin{subequations}\label{eq:linearmodel}
		\begin{align}
		y_{id} & = \beta_{0d} + \sum_{j=1}^p x_{ij} \beta_{jd} + \sum_{t=1}^T z_{it} u_{td} + \epsilon_{id} \\
		& = \beta_{0d} + \x_i \tr \bbeta_d + \mathbf{z}_i \tr \mathbf{u}_d + \epsilon_{id},\\
		\text{with } & \epsilon_{id} \sim \mathcal{N}(0, \sigma_d^2),
		\end{align}
	\end{subequations}
	where $p$-dimensional $\bbeta_d$ and $T$-dimensional $\bm{u}_d$ are the drug-specific omics feature and tissue effect vectors, respectively. Note that (\ref{eq:linearmodel}) gives rise to a system of $d=1, \dots, D$ linear simultaneous equations.

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate them out. The remaining parameters are endowed with the following priors:
	\begin{subequations}\label{eq:prior}
		\begin{align}
		\beta_{jd} & \sim \mathcal{N}_p (0, \phi^2_{g(j)} \gamma_d^2 \sigma_d^2), \label{eq:betaprior}\\
		\mathbf{u}_{d} & \sim \mathcal{N}_T (\mathbf{0}, \bm{\Xi}_d \sigma_d^2),
		\end{align}
	\end{subequations}
	and hyper-priors:
	\begin{subequations}\label{eq:hyperprior}
		\begin{align}
			\sigma_d^{2} & \sim 1/\sigma_d^{3} \\
			\gamma_d^{2} & \sim \mathcal{IG}(\theta_d, \lambda), \\
			\bm{\Xi}_d & \sim \mathcal{W}^{-1}_T(\bm{\Omega}, \nu),
		\end{align}
	\end{subequations}
	where $\mathcal{IG}(\theta, \lambda)$ is an inverse Gaussian distribution with mean $\theta$ and shape $\lambda$, and $\mathcal{W}^{-1}_T(\bm{\Omega}, \nu)$ is an inverse Wishart distribution with $T \times T$-dimensional positive definite (PD) scale matrix $\bm{\Omega}$ and degrees of freedom $\nu$. Note that for $\nu > T + 1$, this inverse Wishart prior implies an expected prior covariance matrix $(\nu - T - 1)^{-1} \bm{\Omega}$ for the tissue effects, common to all drugs. Likewise, the molecular feature effects are related through the common hyper-prior for the $\gamma_d^2$.

	- Jeffrey's prior for variance $1/\sigma^{3/2}$ (a priori independent data mean and variance) or $1/\sigma^3$ (joint prior, but only mean and variance, not variance and betas).

	A few remarks on the choice of priors are justified here: many authors endow error variance components with vague gamma priors. \cite{gelman_prior_2006}, among others, advises against this practice. The degree of `vagueness' has a large influence on the posterior, while degree of `vagueness' is a difficult parameter to set. This influence is especially pronounced if the likelihood is relatively flat, as may be reasonably expected in the large $p$, small $n$ setting we are in. We therefore model the error variance with Jeffrey's prior \cite[]{jeffreys_invariant_1946}.

	Furthermore, we choose to model the $\gamma^2_d$ by an inverse Gaussian distribution, as has been suggested in \cite{fabrizi_specification_2016} and \cite{caron_sparse_2008}, because it allows to model the mean $\theta_d$ as a function of the drug covariates $\mathbf{c}_d$, as explained in Section \ref{sec:empiricalbayes}.

	- IG hyperprior used in finance \cite[]{barndorff-nielsen_normal_1997}, where it is called the NIG prior.

	\section{Estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel}), (\ref{eq:prior}), and (\ref{eq:hyperprior}) is not available in closed form. We therefore approximate the posterior by variational Bayes, where we force the posterior density $Q$ to factorise as:
	\begin{align}\label{eq:variationalposterior}
	p(\B,\mathbf{U},\mathbf{\sigma}^2, \bgamma^2,\bm{\Xi}_1, \dots, \bm{\Xi}_D) \approx Q(\cdot) = q(\B) \cdot q(\mathbf{U}) \cdot q(\mathbf{\sigma}^2) \cdot q(\bgamma^2) \cdot q(\bm{\Xi}_1, \dots, \bm{\Xi}_D),
	\end{align}
	where $\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$ and $\U = \begin{bmatrix} \mathbf{u}_1 & \cdots & \mathbf{u}_D \end{bmatrix}$, are of dimensions $p \times D$ and $T \times D$, respectively. Under the factorisation in (\ref{eq:variationalposterior}), the marginal variational posteriors that minimise the Kullback-Leibler divergence of the true posterior to the variational Bayes approximation \cite[]{neal_view_1998}, are given by:
	\begin{subequations}\label{eq:variationalposterior2}
		\begin{align}
		q_{\B} (\B) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_p (\bmu_d, \bSigma_d), \\
		q_{\U}(\U) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_T (\mathbf{m}_d, \Sm_d), \\
		q_{\bgamma^{2}}(\bgamma^{2}) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG} \(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_{d} \), \\
		q_{\bsigma^{2}}(\bsigma^{2}) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + T + 1}{2}, \zeta_{d} \), \\
		q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \overset{D}{\propto} \prod_{d=1}^D \mathcal{W}_T^{-1} (\bm{\Psi}_d, \nu + 1).
		\end{align}
	\end{subequations}
	Here $\mathcal{GIG}(\cdot)$ and $\Gamma^{-1}(\cdot)$ denote the generalized inverse Gaussian and inverse gamma distributions, respectively.

	The parameters in (\ref{eq:variationalposterior2}) contain cyclic dependencies. We therefore iterate the following estimating equations until convergence to a local optimum:
	\begin{align*}
	\bSigma_d^{(h+1)} & = \frac{2\zeta_d^{(h)}}{n+p+T+1} \left\{\X \tr \X + a_d^{(h)} \cdot \diag (\phi_{g(j)}^{-2}) \right\}^{-1}, \\
	\bm{\mu}_d^{(h+1)} & = \left\{\X \tr \X + a_d^{(h)} \cdot \diag (\phi_{g(j)}^{-2}) \right\}^{-1} \X \tr ( \y_d - \Z \mathbf{m}_d^{(h)}), \\
	\Sm_d^{(h+1)} & = \frac{2\zeta_d^{(h)}}{n+p+T+1} \[ \Z \tr \Z + (\nu + 1) \cdot (\bm{\Psi}_d^{(h)})^{-1} \]^{-1}, \\
	\mathbf{m}_d^{(h+1)} & = \[ \Z \tr \Z + (\nu + 1) \cdot (\bm{\Psi}_d^{(h)})^{-1} \]^{-1} \Z \tr (\y_d - \X \bm{\mu}_d^{(h)}), \\
	\bm{\Psi}_d^{(h+1)} & = \frac{2\zeta_d^{(h)}}{n+p+T+1} \[ \Sm_d^{(h+1)} + \mathbf{m}_d^{(h+1)} (\mathbf{m}_d^{(h+1)}) \tr \] + \bm{\Omega}, \\
	\delta_d^{(h+1)} & =  \frac{n+p+T+1}{2\zeta_d^{(h)}} \sum_{j=1}^p \frac{(\mu_{jd}^{(h+1)})^2 + (\bSigma_d^{(h+1)})_{jj}}{\phi^2_{g(j)}} + \lambda, \\
	\zeta_d^{(h+1)} & = \frac{1}{2} \y_d \tr \y_d - \y_d \tr (\X \bmu_d^{(h+1)} + \Z \mathbf{m}_d^{(h+1)}) + (\bmu_d^{(h+1)}) \tr \X \tr \Z \mathbf{m}_d^{(h+1)} + \frac{1}{2} \trace (\X \tr \X \bSigma_d^{(h+1)}) \\
	& \,\,\,\,\,\,\,\,\,\, + \frac{1}{2} (\bmu_d^{(h+1)}) \tr \X \tr \X \bmu_d^{(h+1)} + \frac{1}{2} \trace (\Z \tr \Z \Sm_d^{(h+1)}) + \frac{1}{2} (\mathbf{m}_d^{(h+1)}) \tr \Z \tr \Z \mathbf{m}_d^{(h+1)} \\
	& \,\,\,\,\,\,\,\,\,\, + \frac{1}{2} a_d^{(h)} \sum_{j=1}^p \frac{(\mu_{jd}^{(h+1)})^2 + (\bSigma_d^{(h+1)})_{jj}}{\phi^2_{g(j)}} + \frac{1}{2} (\nu + 1) \cdot \trace \[ (\bm{\Psi}_d^{(h+1)})^{-1} \Sm_d^{(h+1)} \] \\
	& \,\,\,\,\,\,\,\,\,\, + \frac{1}{2} (\nu + 1) \cdot (\mathbf{m}_d^{(h+1)}) \tr (\bm{\Psi}_d^{(h+1)})^{-1} \mathbf{m}_d^{(h+1)},
	\end{align*}
	where
	$$
	a_d^{(h)} = \sqrt{\frac{\lambda}{\theta^2_d \delta_d^{(h)}}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda \delta_d^{(h)}}/\theta_d \)}{K_{\frac{p+1}{2}} \( \sqrt{\lambda \delta_d^{(h)}}/\theta_d \)} + \frac{p+1}{\delta_d^{(h)}},
	$$
	and $K_{\alpha}(\cdot)$ is the modified Bessel function of the second kind.

	\subsection{Empirical Bayes}\label{sec:empiricalbayes}
	A full Bayesian model requires the specification of hyper-parameters $\bm{\eta} = \begin{bmatrix} (\bm{\phi}^2) \tr & \bm{\theta} \tr & \lambda & \text{vec}(\bm{\Omega}) \tr & \nu \end{bmatrix} \tr$. These are abstract and hard to interpret parameters for which we generally lack expert knowledge. They do, however, have a significant influence on the shape of the posterior distribution. We therefore propose to estimate the unspecified hyper parameters $\bm{\eta}$ by empirical Bayes. Simply put, empirical Bayes fits the prior model to the data and is, as such, objective (up to model specification).

	The canonical method for empirical Bayes is maximisation of the marginal likelihood with respect to the hyper parameters. In \cite{casella_empirical_2001} the marginal likelihood is maximised by an EM algorithm:
	$$
	\bm{\eta}^{(l+1)} = \underset{\bm{\eta}}{\argmax}\E_{\cdot | \Y} [\log p(\Y, \B, \U, \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bgamma^2, \bsigma^2) | \bm{\eta}^{(l)}],
	$$
	where the expectation is with respect to the posterior. In our case, this posterior is not available in closed form, which renders the expectation difficult. While \cite{casella_empirical_2001} suggests to approximate the expectation by a Monte Carlo sample, we propose to use the variational Bayes approximation developed in Section \ref{sec:variationalbayes}:
	\begin{align}\label{eq:variationalbayes}
	\bm{\eta}^{(l+1)} & = \underset{\bm{\eta}}{\argmax}\E_{Q^{(l)}} [\log p(\Y, \B, \U, \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bgamma^2, \bsigma^2)] \nonumber \\
	& = \underset{\bm{\eta}}{\argmax} \E_{Q^{(l)}} [\log \pi (\B | \bgamma^2, \bsigma^2)] + \E_{Q^{(l)}} [\log \pi (\bgamma^2)] + \E_{Q^{(l)}} [\log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D)].
	\end{align}
	where now the expectation is with respect to the converged variational posterior $Q^{(l)}$. Inspection of (\ref{eq:variationalbayes}) learns us that the problem may be decomposed into three separate optimisation problems.

	\subsubsection{Group-specific variance components}
	The first sub-problem concerns the group-specific variance components of the molecular feature effects. In order to separate the overall variance of the molecular feature effects from the group-specific deviations, we impose $\prod_{g=1}^G (\phi_g^2)^{|\mathcal{G}_g|} = 1$. As a consequence, the overall level of shrinkage is determined by the $\gamma^2_d$, while the $\phi_g^2$ effectuate differential shrinkage of the groups. Because the $\phi_g^2$ are variance components, we additionally require $\phi^2_1, \dots, \phi^2_G> 0$, such that our first optimisation sub-problem becomes:
	\begin{align*}
	(\bm{\phi}^2)^{(l+1)} & = \underset{\bm{\phi}^2}{\argmax} \E_{Q^{(l)}} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& = \underset{\bm{\phi}^2}{\argmax} -\frac{D}{2} \sum_{g=1}^G |\G_g| \log \phi_g^2 - \frac{1}{2} \sum_{g=1}^G b_g^{(l)} \phi_g^{-2} \\
	& \text{subject to } \prod_{g=1}^G (\phi_g^2)^{|\G_g|} = 1 \text{ and } \phi^2_1, \dots, \phi^2_G> 0.
	\end{align*}
	where 
	$$
	b_g^{(l)}=\sum_{d=1}^D a_d^{(l)} \frac{n+p+T+1}{2\zeta_d^{(l)}} \sum_{j \in \mathcal{G}_g} \[ (\mu_{jd}^{(l)})^2 + (\bSigma_d^{(l)})_{jj} \].
	$$
	This is a convex optimisation problem and easily solved numerically.

	\subsubsection{Drug-specific variance components}
	The second sub-problem concerns the prior for the drug-specific variance components $\gamma_d^2$. As explained in Section \ref{sec:SEM}, we may have `covariates' $\mathbf{c}_d$ on the drugs available. We conjecture that these covariates may be informative for the omics effect size and therefore model the prior mean of the $\gamma_d^2$ as a function of these covariates: $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$. Then, the empirical Bayes estimation of $\bm{\alpha}$ effectively boils down to an inverse Gaussian regression of the $\gamma_d^{-2}$ on the $\mathbf{c}_d$ \cite[]{fries_optimal_1986,whitmore_regression_1983}. The problem may be formulated as:
	% \begin{align*}\label{eq:optproblem2}
	\begin{align*}
	\lambda^{(l+1)}, \bm{\alpha}^{(l+1)} & = \underset{\lambda,\bm{\alpha}}{\argmax} \E_{Q^{(l)}} [\log \pi (\bgamma^2)] \\
	& = \underset{\lambda,\bm{\alpha}}{\argmax} \frac{D}{2} \log \lambda - \frac{\lambda}{2} \bm{\alpha} \tr \mathbf{C} \tr \diag [\E_{Q^{(l)}}(\gamma_d^2)] \mathbf{C} \bm{\alpha} + \lambda \bm{\alpha} \tr \mathbf{C} \tr \mathbf{1} - \frac{\lambda}{2} \sum_{d=1}^D \E_{Q^{(l)}}(\gamma_d^{-2}),
	\end{align*}
	where $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & \mathbf{c}_D \end{bmatrix} \tr$. The solutions to this problem are:
	\begin{align*}
	\bm{\alpha}^{(l+1)} & = \left\{ \mathbf{C} \tr \Vm^{(l)} \mathbf{C} \right\}^{-1} \mathbf{C} \tr \mathbf{1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D a_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D,
	\end{align*}
	where 
	$$
	\Vm^{(l)} = \diag \[ \sqrt{\frac{\delta_d^{(l)}(\theta_d^{(l)})^2}{\lambda^{(l)}}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta^{(l)}_d \)}{K_{\frac{p + 1}{2}} \( \sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta^{(l)}_d \)} \].
	$$

	- Think about parametrisation of $\mathbf{C}$. Idea: use overall mean and deviations from it in alpha, such that the starting value can be the mean and 0 for alpha. Or: use intercept and code the rest as dummy variables.

	\subsubsection{Tissue effect covariance}
	The third sub-problem pertains to the tissue effect covariance prior. To avoid overfitting when freely estimating the $\frac{T(T+1)}{2}$ elements in $\bm{\Omega}$ and $\nu$, we parametrise them as $\bm{\Omega}=  \tau [(\tau - \rho) \I_T + \rho \mathbf{1}_{T \times T}]$ and $\nu = \tau + T + 1$, respectively. The justification for this parametrisation is three-fold: (a) it implies an easy to interpret prior covariance matrix with $\tau$ on its diagonal and $\rho$ on its off-diagonals, (b) it reduces the number of parameters to  estimate to just two, and (c) simplifies the positive definite constraint on  $\bm{\Omega}$. To ensure positive definiteness and a proper covariance matrix, we impose the constraints $\tau > \max\[\rho - T \rho,\rho, 0\]$ (see Appendix \ref{app:priortissuecovariance}). The resulting third sub-problem is a bit more involved than the first two:
	\begin{align*}
	\tau^{(l + 1)}, \rho^{(l+1)} & = \underset{\tau, \rho}{\argmax} \frac{D (T^2 + T)}{2} \log \tau + \frac{DT}{2} \tau \log \tau + \frac{D(T+1)}{2} \log \[ \tau - (T - 1) \rho\] \\
	& \,\,\,\,\,\,\,\,\,\, + \frac{D}{2} \tau \log \[ \tau - (T - 1) \rho\] + \frac{D(T^2 - 1)}{2} \log(\tau - \rho) +\frac{D(T - 1)}{2} \tau  \log(\tau - \rho) \\
	& \,\,\,\,\,\,\,\,\,\, - D \sum_{t=1}^T \log \Gamma \( \frac{\tau + t + 1}{2} \) - w_1^{(l)} \tau^2 + w_2^{(l)} \tau \rho + w_3^{(l)} \tau, \\
	& \text{subject to } \tau > \max\[\rho - T \rho,\rho, 0\],
	\end{align*}
	where
	\begin{align*}
	w_1^{(l)} & =  \frac{\tau^{(l)} + T + 2}{2} \sum_{d=1}^D \trace \[ (\bm{\Psi}_d^{(l)})^{-1}\] > 0,\\
	w_2^{(l)} & = \frac{\tau^{(l)} + T + 2}{2} \sum_{d=1}^D \left\{\trace \[ (\bm{\Psi}_d^{(l)})^{-1}\] - \trace \[\mathbf{1}_{T \times T} \cdot (\bm{\Psi}_d^{(l)})^{-1}\] \right\} < w_1^{(l)},\\
	w_3^{(l)} & = \frac{1}{2} \[D \sum_{t=1}^T \psi\(\frac{\tau^{(l)} + t + 2}{2}\) - \sum_{d=1}^D \log |\bm{\Psi}_d^{(l)}| \].
	\end{align*}

	- use mathematica to check convexity

	- we use the ELBO for VB convergence and parameters for EB convergence

	- we might be lenient for VB convergence. The extreme case would be 1 VB 
	iteration per EB iteration (corresponds to tolerance of infinity).

	\section{Prior tissue covariance constraints}\label{app:priortissuecovariance}
		We parametrise the prior covariance as a matrix with $\tau$ on the diagonal and $\rho$ on its off-diagonals. To that end we set $\nu = \tau + T + 1$ and the $T \times T$-dimensional matrix
		$$
		\bm{\Omega} = \tau \begin{bmatrix}
		\tau &  & \rho \\
		& \ddots &  \\
		\rho & & \tau
		\end{bmatrix},
		$$
		such that $\E_{\pi(\mathbf{\Xi}_d)} (\mathbf{\Xi}_d) = (\nu - T - 1)^{-1} \bm{\Omega}$ is of the desired form. To ensure a non-degenerate prior distribution we require $\tau > 0$ and $\bm{\Omega}$ PD. We use $\tau > 0$ to note that a PD $\bm{\Omega}$ implies positive roots $\lambda$ of the characteristic polynomial:
		\begin{align*}
		\left| \begin{bmatrix}
		\tau &  & \rho \\
		& \ddots &  \\
		\rho & & \tau
		\end{bmatrix} -\lambda \I_T \right| = (\tau - \lambda - \rho)^{T-1}[\tau - \lambda - (1-T)\rho] = 0.
		\end{align*}
		For $T > 1$, the solutions are:
		$$
		(\lambda = \tau - \rho) \lor (\lambda = T \rho - \rho + \tau),
		$$
		which are positive if the following condition holds: $\tau > \max(\rho,\rho - T \rho)$. We combine this with $\tau > 0$, to arrive at the full proper prior constraint: $\tau > \max(0,\rho,\rho - T \rho)$.

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	In the following all expectations are with respect to the variational posterior $Q$.
	\begin{align*}
	\log q_{\B} (\B) & \propto \E_{\mathbf{U}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\bgamma^2, \bsigma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\mathbf{u}_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E_{\gamma^2_d, \sigma^2_d} \( \frac{\beta_{jd}^2}{\gamma_d^2 \sigma_d^2} \), \\
	q_{\B} (\B) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	& \text{ with } \bSigma_d = \E(\sigma_d^{-2})^{-1} [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1}, \\
	& \text{ and } \bmu_d = [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1} \X \tr [\y_d - \Z \cdot \E(\mathbf{u}_d)].
	\end{align*}

	\begin{align*}
	\log q_{\U}(\U) & \propto \E_{\mathbf{B}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2} [\log \pi (\U | \mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2)] \\
	& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\bbeta_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \E_{\bm{\Xi}_d, \sigma_d^2}\( \frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \), \\
	q_{\U}(\U) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_T (\mathbf{m}_d, \Sm_d), \\
	& \text{ with } \Sm_d = \E(\sigma_d^{-2})^{-1} [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1}, \\
	& \text{ and } \mathbf{m}_d = [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1} \Z \tr [\y_d - \X \cdot \E(\bbeta_d)].
	\end{align*}

	\begin{align*}
	\log q_{\bgamma^{2}}(\bgamma^{2}) & \propto \E_{\B, \bsigma^2} [\log \pi (\B | \bgamma^{2}, \sigma^2)] + \log \pi (\bgamma^{2}) \\
	& \propto -\frac{p}{2} \sum_{d=1}^D \log \gamma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} - \frac{3}{2} \sum_{d=1}^D \log \gamma_d^{2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{\lambda}{2 \theta^2} \sum_{d=1}^D (\gamma_d^2 - \theta)^2 \gamma_d^{-2} \\
	& \propto \sum_{d=1}^D \left\{ \( -\frac{p + 1}{2} - 1\) \log \gamma_d^{2} - \frac{\lambda}{2\theta^2} \gamma_d^2 - \frac{1}{2}\[ \lambda + \E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^{2}}\] \gamma_d^{-2} \right\}, \\
	q_{\bgamma^{2}}(\bgamma^{2}) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG} \(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_{d} \), \\
	& \text{with } \delta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} + \lambda.
	\end{align*}

	\begin{align*}
	\log q_{\bsigma^{2}}(\bsigma^{2}) & \propto \E_{\B,\mathbf{U}} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\B, \bgamma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
	& \,\,\,\,\,\,\,\,\,\, + \E_{\mathbf{U}, \bm{\Xi}_1, \dots, \bm{\Xi}_D} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bsigma^{2}) \\
	& \propto -\frac{n}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \sigma_d^{2} - \frac{p}{2} \sum_{d=1}^D \log \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \E( \gamma_d^{-2}) \frac{\E ( \beta_{jd}^2)}{\phi_{g(j)}^{2}} \sigma_d^{-2} - \frac{T}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \E(\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}) \sigma_d^{-2} \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{3}{2} \sum_{d=1}^D \log \sigma^2_d \\
	& = -\sum_{d=1}^{D} \(\frac{n + p + T + 3}{2} \) \log \sigma_d^{2} - \sum_{d=1}^{D} \frac{1}{2}\Bigg\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \\
	& \,\,\,\,\,\,\,\,\,\, + \E(\gamma_d^{-2})\sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^2} + \E(\mathbf{u}_{d} \tr \bm{\Xi}_d^{-1} \mathbf{u}_d) \Bigg\} \sigma_d^{-2}, \\
	q_{\bsigma^{2}}(\bsigma^{2}) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + T + 1}{2}, \zeta_{d} \), \\
	& \text{ with } \zeta_{d} = \frac{\y_d \tr \y_d}{2} - \y_d \tr [\X \E(\bbeta_d) + \Z \E(\mathbf{u}_d)] + \E(\bbeta_d \tr) \X \tr \Z \E (\mathbf{u}_d) \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\trace [\X \tr \X \V(\bbeta_d)]}{2} + \frac{\E(\bbeta_d \tr) \X \tr \X \E(\bbeta_d)}{2} + \frac{\trace [\Z \tr \Z \V(\mathbf{u}_d)]}{2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\E(\mathbf{u}_d \tr) \Z \tr \Z \E(\mathbf{u}_d)}{2} + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} \\
	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+ \frac{\E(\mathbf{u}_d \tr) \E(\bm{\Xi}_d^{-1}) \E(\mathbf{u}_d)}{2} + \frac{\trace [\E(\bm{\Xi}_d^{-1}) \V(\mathbf{u}_d)]}{2},
	\end{align*}

	\begin{align*}
	\log q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \propto \E_{\mathbf{U}, \bm{\sigma}^2} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D) \\
	& \propto -\frac{1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \E \(\frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \) \\
	& \,\,\,\,\,\,\,\,\,\, - \frac{k + T + 1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace (\bm{\Omega} \bm{\Xi}_d^{-1}) \\
	& = -\frac{\tau + T + 2}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace \left\{ \[\E(\sigma_d^{-2}) \E( \mathbf{u}_d \mathbf{u}_d \tr) + \bm{\Omega}\] \bm{\Xi}_d^{-1} \right\}, \\
	q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \overset{D}{\propto} \prod_{d=1}^D \mathcal{W}_T^{-1} (\bm{\Psi}_d, \nu + 1),  \\ 
	& \text{ with } \bm{\Psi}_d =\E(\sigma^{-2}_d) \[\V(\mathbf{u}_d) + \E(\mathbf{u}_d) \E( \mathbf{u}_d \tr) \] + \bm{\Omega}.
	\end{align*}

	\section{Simulations}
	\subsection{Models}
	In order to investigate the inverse Gaussian prior model and its estimation, we consider a simpler model. The simpler model ignores tissue effects and fixes $\forall g: \phi_g^2=1$:
	\begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \gamma_d^2, \sigma_d^2 & \sim p(\beta_{jd}), \\
  \gamma_d^2 & \sim p(\gamma_d^2), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ and $\sigma_d^2$ should be independent \textit{a priori}. However, in our case that is debatable. The $D$ regressions need not be on the same scale, so some form of calibration of the $\bm{\beta}_d$ posteriors to this scale is required. A simple way of doing this is to include the scales $\sigma_d^2$ in the priors for the $\bm{\beta}_d$. We call the two versions of the model conjugate (with $\bm{\beta}_d$ and $\sigma_d^2$ dependent \textit{a priori}):
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2)
  $$
  and non-conjugate (with $\bm{\beta}_d$ and $\sigma_d^2$ independent \textit{a priori}):
  $$
  \beta_{jd} | \gamma_d^2, \sigma_d^2 \sim \beta_{jd} | \gamma_d^2 \sim \mathcal{N} (0, \gamma_d^2),
  $$
  and compare them in our simulations.

	In addition to the inverse Gaussian model for the prior variances $\gamma_d^2$ (introduced in the MD):
	$$
  \gamma_d^2 \sim \mathcal{IG}(\theta_d, \lambda_d),
  $$
	we investigate a generalisation of the models in \cite{kpogbezan_empirical_2017} and \cite{leday_gene_2017}, where we draw the $\gamma_d^2$ from an inverse Gamma distribution:
	$$
	\gamma_d^2 \sim \Gamma^{-1}(\eta_d/2, \lambda_d/2),
	$$
	with shape $\eta_d/2$ and scale $\lambda/2$. For the sake of comparability, we
	have rescaled the inverse Gamma parameters and reparametrised both 
	distributions as generalized inverse Gaussians:
	\begin{align*}
  \gamma_d^2 & \sim \mathcal{GIG}(-1/2, \lambda_d \theta_d^{-2}, \lambda_d), \\
  \gamma_d^2 & \sim \mathcal{GIG}(-\eta_d/2, 0, \lambda_d).
  \end{align*}
	The two models coincide if $\theta_d \to \infty$ and $\eta_d = 1$.

<<dens_igaussian_igamma, fig.cap="Several densities of the inverse Gaussian and inverse Gamma families", out.width="100%", fig.asp=2/3>>=
digauss <- function(x, theta, lambda, eta) {
  sqrt(lambda/(x^3*2*pi))*exp(-lambda*(x - theta)^2/(2*theta^2*x))
}

digamma <- function(x, theta, lambda, eta) {
  alpha <- eta/2
  beta <- lambda/2
  beta^alpha/gamma(alpha)*x^(-alpha - 1)*exp(-beta/x)
}

eta <- c(1, 2, 1, 1, 5)
lambda <- c(1, 1, 2, 1/2, 2)
theta <- c(1000, 1000, 1, 1, 1)

labels <- c("inv. Gaussian", "inv. Gamma")
col <- c(1, 2)

opar <- par(no.readonly=TRUE)
par(mar=opar$mar*c(1, 1.3, 1, 1))
layout(matrix(c(rep(c(1, 1, 2, 2, 3, 3), 2), rep(c(0, 4, 4, 5, 5, 0), 2)),
                 nrow=4, ncol=6, byrow=TRUE))
for(i in 1:length(eta)) {
  modes <- c(theta[i]*(sqrt(1 + 9*theta[i]^2/
                            (4*lambda[i]^2)) - 3*theta[i]/(2*lambda[i])),
             lambda[i]/(eta[i] + 2))
  ylim <- c(0, max(digauss(modes[1], theta[i], lambda[i], eta[i]),
                   digamma(modes[2], theta[i], lambda[i], eta[i])))
  curve(digauss(x, theta[i], lambda[i], eta[i]), 0.001, 5, n=1000,
        ylim=ylim, ylab="Density", "x", col=col[1], 
        main=bquote(eta==.(eta[i])*","~lambda==.(lambda[i])*","~theta==.(
          theta[i])))
  curve(digamma(x, theta[i], lambda[i], eta[i]), add=TRUE, 
        col=col[2], n=1000)
}
legend("topright", legend=labels, lty=1, col=col, seg.len=1)
par(opar)
@

<<dens_igaussian_marginalbeta, fig.cap="Marginal $p(\\beta_j)$ prior for several choices of $p(\\gamma^2_j)$ scaled to $\\V(\\beta_j)=1$.", out.width="60%", fig.asp=1>>=
library(GeneralizedHyperbolic)
# all scaled such that variance is one
sigma <- 1
theta <- 1/sigma^2
lambda <- c(0.1, 2)
lambda1 <- sqrt(2)
col <- c(1:4)
lty <- c(1:4)
labels <- as.expression(c(bquote("IG, "~lambda==.(lambda[1])~", "~
                                   theta==.(theta)),
                          bquote("IG, "~lambda==.(lambda[2])~", "~
                                   theta==.(theta)),
                          "Point mass (ridge)", "Exponential (lasso)"))
dprior <- function(x, lambda, theta, sigma) {
  dnig(x, 0, sigma*sqrt(lambda), sqrt(lambda/(theta*sigma)), 0)
}
dlasso <- function(x, lambda1) {
  0.5*lambda1*exp(-lambda1*abs(x))
}

curve(dprior(x, lambda[1], theta, sigma), -3, 3, ylab="Density", n=1000, 
      col=col[1], lty=lty[1])
curve(dprior(x, lambda[2], theta, sigma), add=TRUE, n=1000, 
      col=col[2], lty=lty[2])
curve(dnorm(x, 0, 1), add=TRUE, n=1000, col=col[3], lty=lty[3])
curve(dlasso(x, lambda1), add=TRUE, n=1000, col=col[4], lty=lty[4])
legend("topright", legend=labels, col=col, lty=lty)
@

	\subsection{Estimation}
	The variational distributions are as follows:
	\begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+\eta_d}{2}, \lambda_d \theta_d^{-2}, \delta_d\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{df + 1}{2}, \zeta_d\),
  \end{align*}
	where $\eta_d=1$ and $\theta_d \to \infty$ in the inverse Gaussian and inverse Gamma models, respectively. Furthermore, $df=n + p$ in the conjugate model and $df=n$ in the non-conjugate model. The variational parameters contain cyclic dependencies and are, in the conjugate setting, iteratively updated by:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} (\X \tr \X + c_d^{(h)} \I)^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = (\X \tr \X + c_d^{(h)} \I)^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = b_d^{(h)} (c_d^{(h)})^{-1} \bmu_d \tr \bmu_d + \lambda_d, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{b_d^{(h+1)} - c_d^{(h+1)} a_d^{(h+1)}}{1 - a_d^{(h+1)}} \trace ( \bm{\Sigma}_d^{(h+1)}) + \frac{b_d^{(h+1)} - c_d^{(h+1)} a_d^{(h+1)}}{1 - a_d^{(h+1)}} (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)}\bigg].
  \end{align*}
  Here, we set
  \begin{align}
  a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(df+1)/(2 \zeta_d^{(h)}), \nonumber \\
  b_d^{(h)} & = \E_{Q^{(h)}}(\gamma_d^{-2}) = \sqrt{\frac{\lambda_d}{\theta_d^2 \delta_d^{(h)}}} \frac{K_{\frac{p + \eta_d -2}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)}{K_{\frac{p+\eta_d}{2}} \( \sqrt{\lambda_d \delta_d^{(h)}}/\theta_d \)} + \frac{p+\eta_d^{(h)}}{\delta_d^{(h)}}, \label{eq:auxvar1} \\
  c_d^{(h)} & = \begin{cases*}
  b_d^{(h)} & in the conjugate setting \\
  b_d^{(h)}/a_d^{(h)} & in the non-conjugate setting .
  \end{cases*} \nonumber
  \end{align}
  Note that in the inverse Gamma model, the computation of (\ref{eq:auxvar1}) simplifies to $b_d^{(h)}=(p+\eta_d)/\delta_d^{(h)}$.

  \subsection{Empirical Bayes}
  To avoid subjectivity in our choice of prior, we estimate the hyper-parameters with empirical Bayes. The empirical Bayes procedure differs for the two prior models. In the inverse Gaussian case, we parametrise $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ to allow for the inclusion of drug covariates in the empirical Bayes estimation. The estimating equations are:
  \begin{align*}
  \bm{\alpha} & = \[ \mathbf{C} \tr \diag (\lambda_d e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \diag (\lambda_d) \mathbf{1}_{D \times 1}, \\
  \lambda_d & = \[ b_d^{(l)} + e_d^{(l)} (\mathbf{c}_d \tr \bm{\alpha})^2 - 2\mathbf{c}_d \tr \bm{\alpha} \]^{-1},
  \end{align*}
  where $e_d^{(l)} = \E_{Q^{(l)}}(\gamma_d^2) = [b_d^{(l)} - (p + \eta_d^{(h)})/\delta_d^{(l)})] \cdot \delta_d^{(l)} (\theta_d^{(h)})^2/\lambda_d^{(l)}$. Solving the equations is done by iteratively reweighted least squares of responses $(e_d^{(l)})^{-1}$ on predictors $\mathbf{c}_d$ with weights $\lambda_d e_d^{(l)}$. If we use one common $\lambda=\lambda_d$ then the solution is in closed form:
  \begin{align*}
	\bm{\alpha}^{(l+1)} & = \[ \mathbf{C} \tr \diag(e_d^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \mathbf{1}_{D \times 1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D b_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D.
	\end{align*}
	This corresponds to the model in the MD.

  In the inverse Gamma model we use the following empirical Bayes procedure: Suppose we have a partitioning of our drugs into drug classes and we let prior variance components $\gamma_d^2$ and $\gamma^2_{d'}$ be drawn from the same distribution if drug $d$ and drug $d'$ are from the same class. Or simply put: $\gamma^2_{d}, \gamma^2_{d'} \sim \Gamma^{-1}(\eta_c/2, \lambda_c/2)$ if $d, d' \in \text{class}_c$. Empirical bayes estimation of the $\eta_c$ and $\lambda_c$ by maximisation of the (approximated) marginal likelihood amounts to iteratively solving the estimating equations:
  \begin{subequations}
    \begin{align}
      \psi\(\frac{\eta_c}{2}\) & = \log \lambda_c - \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} - \log 2, \label{eq:estequation1}\\
      \lambda_c & = \eta_c \cdot |\text{class}_c| \cdot \(\sum_{d \in \text{class}_c} b_d^{(l)}\)^{-1}, \label{eq:estequation2}
    \end{align}
  \end{subequations}
  until convergence, where $e^{(l)}_d = \E_{Q^{(l)}}(\log \gamma_d^2) = \log \delta_d^{(l)} - \psi [ (p + \eta_d^{(l)})/2] - \log 2$. 
  We propose to do that by first solving (\ref{eq:estequation2}) for $\lambda_c$, plugging it into (\ref{eq:estequation1}) and bounding the result using \cite{alzer_inequalities_1997} to obtain an interval $( \eta_c^*, 2\eta_c^* )$ containing the solution to (\ref{eq:estequation1}), where:
  $$
  \eta_c^* = \left\{ \frac{\sum_{d \in \text{class}_c} e^{(l)}_d}{|\text{class}_c|} + \log \[ \frac{\sum_{d \in \text{class}_c} b_d^{(l)}}{|\text{class}_c|} \]  \right\}^{-1}.
  $$
  With this interval it is straightforward to find $\alpha^{(l+1)}_c$ by any root-finding algorithm and plugging the solution into (\ref{eq:estequation2}) to find $\lambda^{(l+1)}_c$. We may use $\eta_c^*$ as a starting value.

	The inverse Gamma model with independent drug classes allows for more flexible empirical Bayes estimation of the prior means for the different drug classes. As a consequence however, the risk of overfitting is increased. Additionally, it does not allow to include continuous covariates on the drugs.

	\subsection{Evidence lower bound}
	We monitor convergence through the evidence lower bound (ELBO). To compute the evidence lower bound after updating the EB parameters from $\eta_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}$ to $\eta_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$, we compute $\text{ELBO}^{(l + 1)} = \sum_{d=1}^D \text{ELBO}_d^{(l + 1)}$, with, in the inverse Gaussian case:
	\begin{align*}
	\text{ELBO}_d^{(l + 1)} & = \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	& + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)})(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)}) \trace (\bm{\Sigma}_d^{(l)}) \] \\
	& - \frac{df + 1}{2} \log \zeta_d^{(l)} + \frac{p + 1}{4} \log \lambda_d^{(l)} - \frac{p + 1}{2} \log \theta_d^{(l)} - \frac{p+1}{4} \log \delta_d^{(l)} \\
	& + \log K_{\frac{p+1}{2}}\(\sqrt{\lambda_d^{(l)} \delta_d^{(l)}}/\theta_d^{(l)}\) + \frac{\lambda_d^{(l)} e_d^{(l)}}{2(\theta_d^{(l)})^2} \\
	& + \frac{\lambda_d^{(l+1)}}{\theta_d^{(l+1)}} + \frac{1}{2}\log \lambda_d^{(l+1)} - \frac{\lambda_d^{(l+1)} e_d^{(l)}}{2 (\theta_d^{(l + 1)})^2} - \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2}.
	\end{align*}

	In the inverse Gamma setting, we have:
	\begin{align*}
	\text{ELBO}_d^{(l + 1)} & = \frac{1}{2} \log |\bm{\Sigma}_d^{(l)}| - \frac{a_d^{(l)}}{2} \[ \y_d \tr \y_d - 2\y_d \tr \X \bm{\mu}_d^{(l)} + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} + \trace ( \X \tr \X \bm{\Sigma}_d^{(l)})\] \\
	& + \frac{b_d^{(l)}}{2} \[ \delta_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)})(\bm{\mu}_d^{(l)}) \tr \bm{\mu}_d^{(l)} - (a_d^{(l)} + 1 - b_d^{(l)}/c_d^{(l)}) \trace (\bm{\Sigma}_d^{(l)}) \]  \\
	& - \frac{df + 1}{2} \log \zeta_d^{(l)} - \frac{p + \eta_d^{(l)}}{2} \log \delta_d^{(l)} + \frac{\eta_d^{(l)}}{2} e_d^{(l)} + \frac{\eta_d^{(l)}}{2} \log 2 + \log \Gamma \( \frac{p + \eta_d^{(l)}}{2} \) \\
	& - \log \Gamma \( \frac{\eta_d^{(l + 1)}}{2} \) - \frac{\eta_d^{(l+1)}}{2} e_d^{(l)} - \frac{\eta_d^{(l+1)}}{2} \log 2 + \frac{\eta_d^{(l + 1)}}{2} \log \lambda_d^{(l+1)} - \frac{\lambda_d^{(l + 1)} b_d^{(l)}}{2}.
	\end{align*}
  Note that the computation of $e_d^{(l)}$ differs between the inverse Gaussian and Gamma models. Furthermore, to compute the ELBO after a VB update we simply let $\eta_d^{(l)}, \theta_d^{(l)}, \lambda_d^{(l)}=\eta_d^{(l + 1)}, \theta_d^{(l + 1)}, \lambda_d^{(l + 1)}$.

  In practice, there are different options to assess convergence. We may monitor the parameters themselves, the ELBO, or a combination of the two. We may also iterate for an \textit{a priori} fixed number of times. We may also choose to use a different convergence criterium for the EB and VB iterations. Currently we fix the number of EB iterations to 20, while we do 2 VB iterations per EB iteration.

  \subsection{Efficient computation}
  The empirical Bayes updates require the following quantities: $\zeta_d$, $a_d$, $\delta_d$, $\trace (\bm{\Sigma}_d)$, $\trace (\X \tr \X \bm{\Sigma}_d)$, $\bm{\mu}_d \tr \bm{\mu}_d$, and $\bm{\mu}_d \tr \X \tr \X \bm{\mu}_d$. In addition, we need $\log|\bm{\Sigma}|$ and $\mathbf{y}_d \tr \X \bm{\mu}_d$ to monitor the ELBO. The first three quantities are obtained by scalar operations of $\mathcal{O}(n)$. Let $c_d^{(h)}=b_d^{(h)}$ in the conjugate setting and $c_d^{(h)}=b_d^{(h)}/a_d^{(h)}$ in the non-conjugate setting. Then, beforementioned quantities are easily calculated using the SVD $\X = \mathbf{U} \mathbf{D} \mathbf{V} \tr$:
  \begin{align*}
  \trace (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} \[\sum_{i=1}^n (v_i^2 + c_d^{(h)})^{-1} + \max (p - n, 0)\cdot (c_d^{(h)})^{-1}\], \\
  \trace (\X \tr \X \bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} \sum_{i=1}^n v_i^2(v_i^2 + c_d^{(h)})^{-1}, \\
  (\bm{\mu}_d^{(h+1)}) \tr \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)})^2, \\
  (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^4 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)})^2, \\
  \log|\bm{\Sigma}^{(h+1)}| & \propto -p \log a_d^{(h)} - \sum_{i=1} \log (v_i^2 + c_d^{(h)}) - \max(p - n, 0) \cdot \log c_d^{(h)}, \\
  \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} & = \sum_{i=1}^n v_i^2 [(\mathbf{U} \tr \mathbf{y}_d)_i]^2/(v_i^2 + c_d^{(h)}).
  \end{align*}
  where $v_i$ are the singular values of $\X$. The SVD and $\mathbf{U} \tr \mathbf{y}_d$ are $\mathcal{O}(pn^2)$ and $\mathcal{O}(n^2)$ operations, respectively, but have to be calculated only once at the start of the algorithm. The rest of the calculations involve just $n$ scalar multiplications and/or additions.

  \section{Simulation setting}
  - check model with Stan.

  - make one group of drugs and compare methods

  - use drug-specific lambda's

  - try smaller $p$ to see whether $\sigma_d^2$ become more identifiable.

  - scale-invariant instead of conjugate.

  We simulate from a correctly specified model given in \ref{sec:notissuemodel}. We assume random $y_{id}$, $\x_i$, and $\bm{\beta}_d$, while we fix $\mathbf{C}$, $\gamma_d^2$, $\sigma_d^2$ and $\bm{\alpha}$. We estimate the $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, $\bm{\alpha}$ and $\lambda$, while we fix the $\sigma_d^2$ to their true values.

  We set $n=100$, $p=200$, and $D=10$. We consider 5 evenly sized classes of drugs, such that $\mathbf{c}_{kd} = \mathbbm{1}(k = \text{class}_d)$ and $\begin{bmatrix} \mathbf{c}_1 & \cdots & \mathbf{c}_D \end{bmatrix} \tr$. From these the $\gamma_d^2 = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$ are created. Next, we simulate $\bm{\beta}_{jd} \sim \mathcal{N} (0, \sigma^2_d \gamma_d^2)$. We fix the mean signal-to noise ratio $\overline{\text{SNR}}_d$, such that the predictor data variance is $s^2 = \V(\mathbf{x}_i) = \overline{\text{SNR}}_d / (\overline{\gamma^2_d} p)$ (NOT CORRECT). The last step consists of simulating the data by $\mathbf{x}_{ij} \sim \mathcal{N}(0,s^2)$ and $y_{id} = \mathcal{N} (\mathbf{x}_i \tr \bm{\beta}_d, \sigma_d^2)$. The remaining parameters are set as follows: $\bm{\alpha} = \begin{bmatrix} 1 & \dots & 5 \end{bmatrix} \tr$, $\sigma_d^2 = 1$ for all $d$, and $\overline{\text{SNR}}_d = 10$. To account for the random data and parameter generation, we repeat the simulation 100 times. We compare the results to the models in \ref{sec:indvarmodel} and \ref{sec:invgammamodel}, and present the results in Figures \ref{fig:boxplot_igaussian_res1_prior_mean}-\ref{fig:boxplot_igaussian_res2_prior_mean}.

<<boxplot_igaussian_res1_prior_mean, fig.cap="Prior mean estimates", out.width="80%", fig.asp=2/3>>=
@

<<boxplot_igaussian_res2_prior_mean, fig.cap="Prior mean estimates", out.width="80%", fig.asp=2/3>>=
@

  \section{Starting values}
  \subsection{Method 1}
  Starting value for $\sigma_d^2$ as in \cite{chipman_bart:_2010}. They consider a scaled-Chi squared distribution for the error variance with degrees of freedom 3 and scale chosen such that the $q*100$th percentile matches the error variance in the data $s^2(\mathbf{y}_d)$. As starting value for $\sigma_d^2$ we take the mode of this distribution: $\hat{\sigma}_d^2 = 3a/5$, with $a$ the solution to:
  $$
  \Gamma\(\frac{3}{3}\) \[1 - q - F(3a/(2s^2(\mathbf{y}_d)); 3/2, 1) \] = 0,
  $$
  where $F(x; 3/2, 1)$ is the CDF of a gamma function with shape $3/2$ and scale $1$. In \cite{moran_variance_2018}, REFERENCE to Rockova and George (2018), and \cite{chipman_bart:_2010}, $q=0.9$ is suggested. We find that $q \in ( 0.9, 0.95 )$ gives good results.

  \subsection{Method 2}
  We generate starting values as follows. Consider the $\sigma_d^2$ and $\gamma_d^2$ as fixed and estimate them by MML of the regular ridge model:
  $$
  \hat{\sigma}_d^2, \hat{\gamma}_d^2= \underset{\sigma_d^2,\gamma_d^2}{\argmax} \int_{\bm{\beta}} p(\mathbf{y}_d | \bm{\beta}, \sigma_d^2) p(\bm{\beta} | \gamma_d^2, \sigma_d^2) d\bm{\beta}.
  $$
  The regular ridge model is conjugate, so MML is relatively simple. Next we estimate a common inverse Gaussian prior for the $\gamma_d^2$ by considering the estimates to be samples from the prior:
  \begin{align*}
  \hat{\theta} & = n^{-1} \sum_{d=1}^D \hat{\gamma}_d^2 / n, \\
  \hat{\lambda} & = n \left\{\sum_{d=1}^D \[ (\hat{\gamma}_d^2)^{-1} - \hat{\theta}^{-1}\]\right\}^{-1}.
  \end{align*}
  We estimate the mode of the $\hat{m}=\hat{\gamma}_d^2$ by kernel density estimation and equate it to the theoretical mode of the generalized inverse Gaussian distribution with the estimated $\hat{\theta}$ and $\hat{\lambda}$. We solve to find a common $\delta$:
  $$
  \hat{\delta} = \frac{\hat{m}^2 \hat{\lambda}}{\hat{\theta}^2} + (p + 3) \hat{m}.
  $$
  Lastly, we consider the $\hat{\sigma}^2_d$ fixed samples from the inverse Gamma distribution to estimate a common scale:
  $$
  \hat{\zeta} = \frac{D(n + p + 1)}{2 \sum_{d=1}^D[(\hat{\sigma}_d^2)^{-1}]}.
  $$

  \section{Ratios of modified Bessel functions}
  Ratios of modified Bessel functions of the second kind $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be large, this causes numerical issues in the calculation of various quantities. We alleviate the numerical issues through the following.

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
  \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
  \begin{cases}
  \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
  \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} + \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1}, & \text{for } p \text{ odd}.
  \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, such that the ratio may be computed without numerical issues.

  \section{General model}
  We have drug covariates $\mathbf{c}_d$ and omics feature covariates $\mathbf{z}_j$. The model is now:
  \begin{align*}
  y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} (\x_i \tr \bm{\beta}_d, \sigma_d^2), \\
  \beta_{jd} | \tau_j^2, \gamma_d^2, \sigma_d^2 & \sim \mathcal{N} (0, \sigma_d^2 \gamma_d^2 \tau_j^2), \\
  \gamma_d^2 & \sim \mathcal{IG}\((\mathbf{c}_d \tr \bm{\alpha})^{-1}, \lambda\), \\
  \tau_j^2 & \sim \mathcal{IG}\((\mathbf{z}_j \tr \bm{\theta})^{-1}, \nu\), \\
  \sigma_d^2 & \sim 1/\sigma_d^3.
  \end{align*}
  Variational posterior:
  \begin{align*}
  q(\bm{\beta}) & \overset{D}{=} \prod_{d=1}^D \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \\
  q(\bm{\gamma}^2) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG}\(-\frac{p+1}{2}, \lambda (\mathbf{c}_d \tr \bm{\alpha})^2, \delta_d\), \\
  q(\bm{\tau}^2) & \overset{D}{=} \prod_{j=1}^p \mathcal{GIG}\(-\frac{D+1}{2}, \nu (\mathbf{z}_j \tr \bm{\theta})^2, \eta_j\), \\
  q(\bm{\sigma}^2) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
  Variational parameters:
  \begin{align*}
  \bm{\Sigma}_d^{(h+1)} & = \frac{2 \zeta_d^{(h)}}{n+p+1} \[\X \tr \X + a_d^{(h)} \cdot \diag(b_j^{(h)})\]^{-1}, \\
  \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + a_d^{(h)} \cdot \diag(b_j^{(h)})\]^{-1} \X \tr \y_d, \\
  \delta_d^{(h+1)} & = \frac{n+p+1}{2 \zeta_d^{(h)}} \left\{\trace \[\diag(b_j^{(h)}) \bm{\Sigma}_d^{(h+1)} \] + (\bm{\mu}_d^{(h+1)}) \tr \diag(b_j^{(h)}) \bm{\mu}_d^{(h+1)}\right\} + \lambda, \\ 
  \eta_j^{(h+1)} & = \sum_{d=1}^D \frac{a_d^{(h)}(n+p+1)}{2 \zeta_d^{(h)}} \[ (\bmu_{jd}^{(h + 1)})^2 + (\bSigma_d^{(h + 1)})_{jj} \] + \nu, \\ 
  \zeta_d^{(h+1)} & = \frac{1}{2} \Big( \mathbf{y}_d \tr \mathbf{y}_d -2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + a_d^{(h+1)} \trace \left\{ \[\X \tr \X + \diag(b_j^{(h + 1)})\] \bm{\Sigma}_d^{(h+1)}\right\} \\
  & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + a_d^{(h+1)} (\bm{\mu}_d^{(h+1)}) \tr \[\X \tr \X + \diag(b_j^{(h + 1)})\]\bm{\mu}_d^{(h+1)}\Big),
  \end{align*}
  with
  \begin{align*}
  a_d^{(h)} & = \sqrt{\frac{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2}{\delta_d^{(h)}}} \frac{K_{\frac{p-1}{2}} \( \sqrt{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2 \delta_d^{(h)}} \)}{K_{\frac{p+1}{2}} \( \sqrt{\lambda (\mathbf{c}_d \tr \bm{\alpha})^2 \delta_d^{(h)}} \)} + \frac{p+1}{\delta_d^{(h)}} \text{ and} \\
  b_j^{(h)} & = \sqrt{\frac{\nu (\mathbf{z}_j \tr \bm{\theta})^2}{\eta_j^{(h)}}} \frac{K_{\frac{D-1}{2}} \( \sqrt{\nu (\mathbf{z}_j \tr \bm{\theta})^2 \eta_j^{(h)}} \)}{K_{\frac{D+1}{2}} \( \sqrt{\nu (\mathbf{z}_j \tr \bm{\theta})^2 \eta_j^{(h)}} \)} + \frac{D+1}{\eta_j^{(h)}}.
  \end{align*}
  Hyper-parameter updates:
  \begin{align*}
  \bm{\alpha}^{(l+1)} & = \left[ \mathbf{C} \tr \diag ( e_d^{(l)} ) \mathbf{C} \right]^{-1} \mathbf{C} \tr \mathbf{1}, \\
  \lambda^{(l+1)} & = \[ \sum_{d=1}^D a_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D, \\
  \bm{\theta}^{(l+1)} & = \left[ \mathbf{Z} \tr \diag ( f_j^{(l)} ) \mathbf{Z} \right]^{-1} \mathbf{Z} \tr \mathbf{1}, \\
  \nu^{(l+1)} & = \[ \sum_{j=1}^p b_j^{(l)} - (\bm{\theta}^{(l+1)}) \tr \mathbf{Z} \tr \mathbf{1} \]^{-1} p,
  \end{align*}
  where 
  \begin{align*}
  e_d^{(l)} & = \sqrt{\frac{\delta_d^{(l)}}{\lambda^{(l)}(\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda^{(l)} (\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2 \delta_d^{(l)}} \)}{K_{\frac{p + 1}{2}} \( \sqrt{\lambda^{(l)} (\mathbf{c}_d \tr \bm{\alpha}^{(l)})^2 \delta_d^{(l)}} \)} \text{ and} \\
  f_j^{(l)} & = \sqrt{\frac{\eta_j^{(l)}}{\nu^{(l)}(\mathbf{z}_j \tr \bm{\theta}^{(l)})^2}} \frac{K_{\frac{D - 1}{2}} \( \sqrt{\nu^{(l)} (\mathbf{z}_j \tr \bm{\theta}^{(l)})^2 \eta_j^{(l)}} \)}{K_{\frac{D + 1}{2}} \( \sqrt{\nu^{(l)} (\mathbf{z}_j \tr \bm{\theta}^{(l)})^2 \eta_j^{(l)}} \)}.
  \end{align*}

  \bibliographystyle{author_short3}  
	\bibliography{refs}

\end{document}