### 2019-11-19 docs/manuscript2.Rnw

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("code/figures.R")
@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip,xr}
\externaldocument[sm-]{supplement2}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\bsigma}{\bm{\sigma}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bepsilon}{\bm{\epsilon}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bPsi}{\bm{\Psi}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bgamma}{\bm{\gamma}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\Em}{\mathbf{E}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\Lm}{\mathbf{L}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Cov}{\mathbb{C}\text{ov}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Empirical Bayes estimation of the normal inverse Gaussian prior in 
simultaneous-equation models}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Mark A. van de Wiel$^{1,3}$, 
Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

\begin{document}

	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health
	research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands \\*
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge, 
	United Kingdom

	\begin{abstract}
		{...}
	\end{abstract}

	\noindent\textbf{Keywords}: ...

	\section{Introduction}
	Recently, the dimensions of data sets have increased in three directions:
	(i) samples, (ii) variables, and (iii) outcomes. Here, we will 
	concern ourselves with directions (ii) and (iii). More spefically, we will 
	focus on the regression setting
	where we have multiple outcomes and a medium to high dimensional feature 
	space, i.e., $p \gtrsim n$. Standard methods
	underperform or even break down in such medium to high dimensional settings, 
	so new methodology is required.

	As a motivating example, we present an expression quantitative loci (eQTLs)
	detection problem (see Section \ref{sec:application}). In eQTLs detection, 
	the aim is to find a set of single nucleotide loci (SNPs) that explains a
	(part of a) gene's expression. The resulting heritability of the gene 
	expression traits is thought to be indicative of the relationship between 
	genotype and phenotype. eQTLs detection is often done for a set of genes 
	simultaneously and as such comprises a simultaneous-equations model (SEM) with
	multiple outcomes and a high dimensional feature set (the SNPs). Apart from
	eQTLs, SEM models are also applied in network reconstruction 
	\cite[]{kpogbezan_incorporating_2019, leday_gene_2017} 
	and drug target discovery \cite[]{noh_inferring_2016}, [! ref to second 
	manuscript].

	In many multi-outcome regression settings, extra information on both the 
	outcomes and 
	features is available. In our motivating eQTLs detection example, extra 
	information may come in the form of (i) gene annotation (e.g., length of the
	gene), (ii) SNP annotation (e.g., whether the SNP is in the gene or not), 
	(iii) SNP-gene annotation (e.g., distance of SNP to gene). The presented 
	method allows for the incorporation of such information to aid in the 
	estimation, by pooling information both across equations (genes) and 
	features (SNPs).

	We are not the first to propose to learn from external data 
	across equations and features. \cite{kpogbezan_incorporating_2019} uses 
	the presence or absence of edges in a normal tissue apoptosis pathway, to aid 
	in the estimation of a cancer tissue apoptosis pathway. Furhtermore, in an 
	eQTLs detection problem they include the location of the SNPs (inside
	the gene or not) to aid in the estimation. \cite{oman_minimax_2002} uses 
	empirical Bayes to learn from external data concerning both the outcomes and 
	the features in a low dimensional chemometrics dataset. Our proposed method,
	termed \texttt{semnig},
	extends \cite{kpogbezan_incorporating_2019,kpogbezan_empirical_2017} by 
	allowing for the inclusion of continuous external information and several 
	sources of information.

	The rest of the paper is structured as follows. In Section 
	\ref{sec:model} we introduce our model, the estimation of which is detailed
	in Section \ref{sec:estimation}. In Sections \ref{sec:simulations} and
	\ref{sec:application} we demonstrate the method in simulations and an 
  eQTLs detection problem, respectively. We end with a discussion in Section 
  \ref{sec:discussion}.

	\section{Model}\label{sec:model}
	\subsection{Simultaneous equations model}\label{sec:SEM}
	We have continuous measurements $y_{id}$ for observation $i=1,\dots, n$ on 
	outcome $d=1,\dots,D$. Throughout this paper we assume the $y_{id}$ to be 
	centred per outcome and let $\y_d = \begin{bmatrix} y_{1d} & \dots & y_{nd} 
	\end{bmatrix} \tr$. We predict the $y_{id}$ with features $x_{ij}$, 
	$j=1,\dots, p$, collected in $\x_i = \begin{bmatrix} x_{i1} & \dots & x_{ip} 
	\end{bmatrix} \tr$ and centered per feature. We have $|\bm{\alpha}|$ 
	external variables available that describes the outcome $d$, feature $j$, or
	relationship between feature $j$ and outcome $d$, collected in 
  $\mathbf{C} = \begin{bmatrix} \mathbf{c}_{1}^1 & \cdots &
	\mathbf{c}_{1}^p & \cdots & \mathbf{c}_{D}^1 & \cdots & \mathbf{c}_{D}^p
  \end{bmatrix}_{|\bm{\alpha}| \times pD} \tr$. An example of such an external
  variable from eQTL detection is the distance between SNP and gene.

	We model the outcomes $y_{id}$ as a linear function of the features $x_{ij}$:
	\begin{subequations}\label{eq:linearmodel}
		\begin{align}
		y_{id} & = \beta_{0d} + \x_i \tr \bbeta_d + \epsilon_{id},\\
		\text{with } & \epsilon_{id} \sim \mathcal{N}(0, \sigma_d^2),
		\end{align}
	\end{subequations}
	where the $p$-dimensional $\bbeta_d$ are the outcome-specific feature effects.
	Note that (\ref{eq:linearmodel}) gives rise to a system of 
	$D$ linear simultaneous equations.

	\subsection{Bayesian prior model}
	We capture the uncertainty in the parameters through a Bayesian prior model. 
	We assign a flat prior to the intercept terms $\beta_{0d}$ and integrate 
	them out. The remaining parameters are endowed with priors such that we end
	up with the following model:
	\begin{subequations}\label{eq:priormodel}
    \begin{align}
      y_{id} | \bm{\beta}_d, \sigma_d^2 & \sim \mathcal{N} 
      \(\x_i \tr \bm{\beta}_d, \sigma_d^2\), \\
      \beta_{jd} | \gamma_{jd}^2, \sigma_d^2 & \sim \mathcal{N} 
      \(0, \sigma_d^2 \gamma_{jd}^2 \), \label{eq:betaprior} \\
      \gamma_{jd}^2 & \sim \mathcal{IG}
      \( [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}, \lambda_{d} \), \\
      \sigma_d^2 & \sim 1/\sigma_d^3,
    \end{align}
  \end{subequations}
  where $\mathcal{IG} ([(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}, \lambda_{d})$ 
  denotes the inverse Gaussian distribution with mean 
  $[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}$ and shape $\lambda_d$.

  The normal inverse Gaussian (NIG) prior model for $\beta_{jd}$ and 
  $\gamma^2_{jd}$ in (\ref{eq:priormodel}) was introduced in 
  \cite{barndorff-nielsen_hyperbolic_1978} and is, since 
  \cite{barndorff-nielsen_normal_1997}, routinely applied in mathematical 
  finance (see, e.g., \cite{kalemanova_normal_2007}). Supplementary Material
  (SM) Section \ref{sm-sec:nigprior} contains more details on this prior.
  In Figure \ref{fig:dens_igaussian_marginalbeta}, the NIG 
  prior distribution, marginalised over $\gamma_{jd}^2$, is depicted together 
  with three other common choices,
  the Student's $t$ (see Section \ref{sec:inversegammamodel} for more details on 
  this prior), normal (ridge) and Laplace (lasso) priors, 
  all scaled to variance one. 
  From this Figure we see that, depending on the choice
  of hyperparameters, the NIG prior can vary from heavier tails than the lasso, 
  to more Gaussian tails like the ridge, but behaves similarly to the 
  Student's $t$ prior. Our argumentation to model the 
  $\gamma^2_{jd}$ by an inverse Gaussian 
  distribution, as has been suggested in \cite{fabrizi_specification_2016} and 
  \cite{caron_sparse_2008}, is two-fold: (i) the NIG model is more flexible than
  the standard ridge and lasso priors (as seen from Figure 
  \ref{fig:dens_igaussian_marginalbeta}), and (ii) the NIG prior allows to model
  the mean as a 
  function of the external variables $\mathbf{c}_{d}^j$ more conveniently than 
  the Student's $t$ prior, as explained in Section 
  \ref{sec:empiricalbayes}. 
<<dens_igaussian_marginalbeta, fig.cap="Several choices of marginal $\\beta_j$ prior, scaled to $\\V(\\beta_j)=1$.", out.width="60%", fig.asp=3/4>>=
@

  \cite{moran_variance_2018} argue that in regression models, $\bm{\beta}_d$ 
  and $\sigma_d^2$ should be independent \textit{a priori}, i.e.
  $\beta_{jd} | \gamma_{jd}^2, \sigma_d^2 \sim \beta_{jd} | 
  \gamma_{jd}^2$. However, in our case
  that is debatable. The $D$ regressions need not be on the same scale, so in
  order to jointly consider the $\bm{\beta}_d$ some 
  form of calibration is required. A straightforward way 
  of doing this is to scale the $\bm{\beta}_d$ prior with 
  $\sigma_d^2$ as in (\ref{eq:betaprior}), resulting in a 
  scale-free prior with respect to the response. Apart from the scaling,
  in this setting, the conjugate prior model is preferred
  over the non-conjugate model, because the variational Bayes approximation 
  (see Section \ref{sec:estimation}) assumes independent $\sigma_d^2$ and 
  $\bm{\beta}_d$ in the posterior, which is actually true for 
  the conjugate model. Therefore, the approximate expectation used in the 
  empirical Bayes step (see Section \ref{sec:estimation}) is more accurate.

  A few remarks on the choice of error variance prior are justified here: 
  many authors endow error variance components with vague gamma priors. 
  \cite{gelman_prior_2006}, 
  among others, advises against this practice. The degree of `vagueness' has a 
  large influence on the posterior, while degree of `vagueness' is a difficult 
  parameter to set. This influence is especially pronounced if the likelihood is
  relatively flat, as may be reasonably expected in the large $p$, small $n$ 
  setting we are in. We therefore model the error variance with Jeffrey's 
  objective prior 
  \cite[]{jeffreys_invariant_1946}. In the derivation of our Jeffrey's prior
  for the error variance, we jointly consider an unknown mean and variance. This
  joint consideration results in the somewhat unorthodox $1/\sigma^3$ Jeffrey's
  prior.

  Several extensions of the model are described in [! ref to second paper here].

	\section{Estimation}\label{sec:estimation}
	\subsection{Variational Bayes}\label{sec:variationalbayes}
	The posterior corresponding to the model described in (\ref{eq:linearmodel})
	and (\ref{eq:priormodel}) is not available in closed form. We therefore 
	approximate the posterior by variational Bayes, where we force the approximate
	posterior 
	density to factorise as: $p(\bbeta_d, \sigma_d^2, \gamma_{1d}^2, \dots, 
	\gamma_{pd}^2) \approx 
	Q_d(\cdot) = q(\bbeta_d) \cdot q(\sigma_d^2) \cdot \prod_{j=1}^p 
	q(\gamma_{jd}^2)$. For
	notational convenience, we slightly abuse notation and let $q(\cdot)$ denote
	different densities for different inputs.
	Under such a factorisation, the marginal variational posteriors that 
	minimise the Kullback-Leibler divergence of the true posterior to the 
	variational Bayes approximation \cite[]{neal_view_1998} are given by:
	\begin{align*}
    q(\bm{\beta}_d) & \overset{D}{=} \mathcal{N}_p 
    (\bm{\mu}_d, \bm{\Sigma}_d), \\
    q(\gamma_{jd}^2) & \overset{D}{=} \mathcal{GIG}
    \(-1, \lambda_{d}[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2, \delta_{jd}\), \\
    q(\sigma^2_d) & \overset{D}{=} \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_d\).
  \end{align*}
	See SM Section \ref{sm-sec:vbderivations} for the derivations. The variational
	parameters $\bm{\mu}_d$, $\bm{\Sigma}_d$, $\delta_d$, and
	$\zeta_d$ contain cyclic dependencies and are iteratively 
	updated by:
  \begin{align*}
    \bm{\Sigma}_d^{(h+1)} & = (a_d^{(h)})^{-1} 
    \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1}, \\
    \bm{\mu}_d^{(h+1)} & = \[\X \tr \X + \diag(b_{jd}^{(h)})\]^{-1} 
    \X \tr \y_d, \\
    \delta_{jd}^{(h+1)} & = a_d^{(h)} \[(\bmu^{(h+1)}_{jd})^2 + 
    (\bSigma_d)_{jj}\] + \lambda_{d}, \\ 
    \zeta_d^{(h+1)} & = \frac{1}{2} \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(h+1)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(h+1)}) + 
    (\bm{\mu}_d^{(h+1)}) \tr \X \tr \X \bm{\mu}_d^{(h+1)} \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \trace \[ \diag (b_{jd}^{(h+1)}) \bm{\Sigma}_d^{(h+1)}\] + 
    (\bm{\mu}_d^{(h+1)}) \tr \diag (b_{jd}^{(h+1)}) 
    \bm{\mu}_d^{(h+1)}\bigg],
  \end{align*}
  until convergence. Here, we set
  \begin{align}
    a_d^{(h)} & =\E_{Q^{(h)}}(\sigma_d^{-2})=(n + p + 1)/(2 \zeta_d^{(h)}), 
    \nonumber\\
    b_{jd}^{(h)} & = \E_{Q^{(h)}}(\gamma_{jd}^{-2}) = 
    \sqrt{\frac{\lambda_{d}[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2}
    {\delta_{jd}^{(h)}}}
    \frac{K_0(\sqrt{\delta_{jd}^{(h)}\lambda_{d}}
    [(\mathbf{c}_{d}^j) \tr \bm{\alpha}])}
    {K_1(\sqrt{\delta_{jd}^{(h)}\lambda_{d}}
    [(\mathbf{c}_{d}^j) \tr \bm{\alpha}])} + 
    \frac{2}{\delta_{jd}^{(h)}}, \label{eq:ratiomodifiedbessel}
  \end{align}
  where $K_{\nu}(x)$ denotes the modified Bessel function of the second kind.
  A method for fast and numerically stable calculation of ratios of
  modified Bessel functions of the second kind, as in 
  (\ref{eq:ratiomodifiedbessel}), is given in SM Section 
  \ref{sm-sec:ratiosmodifiedbessels}.

  \subsection{Empirical Bayes}\label{sec:empiricalbayes}
  To make use of the external data, we parametrise the prior mean of the
  $\gamma_{jd}^2$ as $[(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^{-1}$. 
  This allows us to
  include continuous external variables $\mathbf{c}_{d}^j$ into the model. 
  Additionally, it reduces the number of hyperparameters from $pD$ to 
  $|\bm{\alpha}| + D$. A full Bayesian model
  then requires the specification of the hyperparameters 
  $\bm{\alpha}$ and $\blambda = \begin{bmatrix} \lambda_1 & \cdots & \lambda_D 
  \end{bmatrix} \tr$. These are abstract and hard to interpret parameters 
  for which we generally lack expert knowledge. They do, however, have a 
  significant influence on the shape of the posterior distribution. We 
  therefore propose to estimate these hyperparameters by empirical Bayes. 
  Simply put, empirical Bayes fits the prior model to the data and is, as such, 
  objective Bayes (up to model specification). In our case, this results in an   
  objective and data-driven inclusion of the external data.

	The canonical method for empirical Bayes is maximisation of the marginal 
	likelihood with respect to the hyper parameters. In 
	\cite{casella_empirical_2001} the marginal likelihood is maximised by an EM 
	algorithm:
	\begin{align*}
	  \bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} & = \underset{\bm{\alpha},
	  \bm{\lambda}}{\argmax}\E_{\cdot | \Y} 
	  [\log p(\Y, \B, \bm{\Gamma}^2, \bsigma^2) | \bm{\alpha}^{(l)},
	  \bm{\lambda}^{(l)}] \\
	  & = \underset{\bm{\alpha},\bm{\lambda}}{\argmax} 
	  \E_{\cdot | \Y} [\log \pi (\bm{\Gamma}^2) | \bm{\alpha}^{(l)},
	  \bm{\lambda}^{(l)}],
	\end{align*}
	where $\Y = \begin{bmatrix} \y_1 & \cdots & \y_D \end{bmatrix}$,
	$\B = \begin{bmatrix} \bbeta_1 & \cdots & \bbeta_D \end{bmatrix}$, 
	$\bsigma^2 = \begin{bmatrix} \sigma^2_1 & \cdots & 
	\sigma^2_D \end{bmatrix} \tr$,
	and $\bm{\Gamma}^2 = \begin{bmatrix} \bgamma^2_1 & \cdots & \bgamma^2_D 
	\end{bmatrix}\tr$, with 
	$\bgamma^2_d = \begin{bmatrix} \gamma^2_{1d} & \cdots & \gamma^2_{pd} 
	\end{bmatrix}\tr$,
	and the expectation is with respect to the joint posterior. In our case, this 
	posterior is not available in closed form, which renders the expectation 
	difficult. While \cite{casella_empirical_2001} suggests to approximate the 
	expectation by a Monte Carlo sample, we propose to use the variational Bayes 
	approximation developed in Section \ref{sec:variationalbayes}:
	$$
	\bm{\alpha}^{(l+1)},\bm{\lambda}^{(l+1)} = 
	\underset{\bm{\alpha},\bm{\lambda}}{\argmax}\E_{Q^{(l)}} 
	[\log \pi(\bm{\Gamma}^2)],
	$$
	where now the expectation is with respect to the converged variational 
	posterior $Q^{(l)}=\prod_{d=1}^D Q_d^{(l)}$. The resulting estimating 
	equations are:
  \begin{align*}
    \bm{\alpha} & = \[ \mathbf{C} \tr \bLambda \diag 
    (e_{jd}^{(l)}) \mathbf{C} \]^{-1} \mathbf{C} \tr \bLambda
    \mathbf{1}_{pD \times 1}, \\
    \lambda_{d} & = p \left\{ \sum_{j=1}^p b_{jd}^{(l)} + 
    \sum_{j=1}^p e_{jd}^{(l)} [(\mathbf{c}_{d}^j) \tr \bm{\alpha}]^2 - 
    2 \sum_{j=1}^p (\mathbf{c}_{d}^j) \tr \bm{\alpha} \right\}^{-1},
  \end{align*}
  where $\bLambda=\diag(\lambda_d) \otimes \I_p$
  and $e_{jd}^{(l)} = \E_{Q^{(l)}}(\gamma_{jd}^2) = (b_{jd}^{(l)} - 
  2/\delta_{jd}^{(l)}) \cdot \delta_{jd}^{(l)}/[\lambda_{d}^{(l)}
  \left\{(\mathbf{c}_{d}^j) \tr \bm{\alpha}^{(l)}]^2\right\}$. 
  Solving the equations is done by 
  iteratively reweighted least squares.

  Variaional Bayes approximations are known to underestimate posterior 
  variances \cite[]{rue_approximate_2009,consonni_mean-field_2007,
  bishop_pattern_2006,wang_inadequacy_2005}. If posterior variances are 
  required, 
  we suggest generating samples from the posterior after the empirical Bayes 
  iterations have converged. A Gibbs sampler to do so is described in 
  the SM Section \ref{sm-sec:gibbssampler}. Alternatively, the \texttt{R} 
  package \texttt{rstan} \cite[]{guo_rstan:_2018} allows the user to sample
  from a generic posterior. An implementation of model
  (\ref{eq:priormodel}) in \texttt{stan} code may be found on 
  \url{https://github.com/magnusmunch/cambridge}.

  Note that so far, the design matrix $\X$ is assumed to be shared across the
  equations. This is not a requirement. That is, we may have D different design 
  matrices $\X^d$ that need not have the same number of 
  columns, i.e., we have $p_d$ instead of $p$. This becomes useful in, e.g. 
  eQTLs detection, where we regress genes on a gene-specific selection
  of SNPs, often based on SNP-gene distance.

  \section{Simulations}\label{sec:simulations}
  We conducted a simulation study consisting of three parts investigating (i) 
  the NIG prior performance, (ii) the variational Bayes approximation, and (iii)
  the empirical Bayes estimation.

  \subsection{NIG prior performance}
  \label{sec:inversegammamodel}
  The aim of this part is to investigate the 
  performance of the NIG prior model. Since we may do so in the univariate 
  setting, so we drop all $d$ indices in the following. 
  A common default prior for latent variance components is the inverse Gamma 
  prior:
  $$
  \gamma_j^2 \sim \Gamma^{-1}(\eta/2,\xi/2),
  $$
  with shape $\eta/2$ and scale $\xi/2$, both rescaled to match our 
  inverse Gaussian prior formulation. The resulting marginal prior for the model
  parameters
  (conditional on the error variance $\sigma^2$) is then a scaled Student's 
  $t$-distribution:
  $$
  \beta_j | \sigma^2 \sim \text{Student's } 
  t\(\eta,\sigma \sqrt{\xi/\eta}\),
  $$
  with degrees of freedom $\eta$ and scale $\sigma \sqrt{\xi/\eta}$.
  See SM Section \ref{sm-sec:inversegamma} for more details on this prior. 
  We compare the 
  posteriors of the proposed NIG to the scaled Student's 
  $t$-distribution prior in a simulated setting.

  We compare the posterior distributions to the true posterior for a 
  number of measures. Divergence between the distributions
  is quantified through a multivariate extension of the Cram{\'e}r-von Mises
  criterion \cite[]{baringhaus_new_2004}, where smaller values of the criterion
  indicate a better fit. We also assess 95\% coverage of the 
  posterior credible intervals. The accuracy of the 
  point estimates of $\beta_j$ in the form of posterior means is assessed by
  the mean squared error (MSE) and correlation between the point estimates 
  and the true $\beta_j$.

  We studied two settings: one dense and one sparse. In the dense setting, 
  we simulated $n=100$ features $\x_i$ from a $p=200$ dimensional Gaussian 
  distribution with zero mean vector, variances equal to one, and all 
  covariances set to $\rho=0.5$. The outcomes were generated through $y_i = 
  \x_i\tr \bbeta + \epsilon_i$, with the $\epsilon_i$ from the standard normal 
  distribution. 
  Likewise, we generated the model parameters $\beta_j$ from a standard normal
  distribution.

  We choose the hyperparameters such that we get all combinations of marginal
  prior variances and kurtoses $\V(\beta_{j} | \sigma^2) \in 
  \{ 1/2, 1, 2 \}$ and $\mathcal{K}(\beta_{j} | \sigma^2) \in 
  \{ 4, 10 \}$. This leads to the settings in Table \ref{tab:hyper_set1}. For 
  all priors, we generated 1000 draws (after 1000 burnin) from the 
  posteriors using an MCMC sampler implemented in \texttt{rstan}. We repeated
  the simulation 100 times and report 
  results in Figures \ref{fig:boxplots_igaussian_res4.1_post1} and 
  \ref{fig:boxplots_igaussian_res4.1_post2}. 
<<hyper_set1>>= 
set <- read.table("results/simulations_igaussian_set4.csv")
kableExtra::add_header_above(kableExtra::kable_styling(
  knitr::kable(set[, c("ctalphainv", "lambda", "eta", "xi", "vars", "kurts")],
               col.names=c("$(\\mathbf{c} \\tr \\bm{\\alpha})^{-1}$",
                           "$\\lambda$", "$\\eta$", "$\\xi$",
                           "$\\V(\\beta_{j} | \\sigma^2)$",
                           "$\\mathcal{K}(\\beta_{j} | \\sigma^2)$"),
               caption="Hyperparameter settings",
               "latex", booktabs=TRUE, escape=FALSE, digits=2),
  latex_options=c("HOLD_position")),
  header=c(" "=1, "NIG"=2, "Student's t"=2, "Prior"=2))
@

<<boxplots_igaussian_res4.1_post1, fig.cap="Comparison of the NIG and Student's t priors in the dense setting in (a)-(d) simulation setting 1, (e)-(h) setting 2, and (i)-(l) setting 3, through Cramér-von Mises criterion between the model and true posteriors, correlation and MSD between the posterior means and true parameters, and 95\\% credible interval coverage of the true values.", out.width="100%", fig.asp=1>>=
@

<<boxplots_igaussian_res4.1_post2, fig.cap="Comparison of the NIG and Student's t priors in the dense setting in (a)-(d) simulation setting 4, (e)-(h) setting 5, and (i)-(l) setting 6, through Cramér-von Mises criterion between the model and true posteriors, correlation and MSD between the posterior means and true parameters, and 95\\% credible interval coverage of the true values.", out.width="100%", fig.asp=1>>=
@

  [! describe results here]

  In the sparse setting, the observed data generating process is the same as in
  the dense setting. The
  model parameters are simulated from an (absolutely continuous) spike-and-slab 
  prior with both Gaussian spike and slab:
  \begin{align*}
    w & \sim \text{Beta}(\phi, \chi), \\
    z_j | w & \sim \text{Bernoulli}(w), \\
    \beta_j | z_j & \sim z_j \mathcal{N} \( 0, 1 + 0.99 \frac{\chi}{\phi}\) +
    (1 - z_j)\mathcal{N} \( 0, 0.01\).
  \end{align*}
  The slab is chosen
  such that the we have marginal prior variance $\V(\beta_j)=1$. The sparsity 
  level is determined by the proportion of zero $\beta_j$: 
  $s=1 - \phi/(\phi + \chi)$. Note that the resulting marginal prior for
  $\beta_j$ is:
  $$
  \beta_j \sim (1 - s)\mathcal{N} \( 0, \frac{1 - 0.01 s}{1 - s}\) + 
  s \mathcal{N} \( 0, 0.01\).
  $$
  We use two settings, $s=0.5$ and $s=0.9$. The 
  hyperparameters for estimation are set the same as in the dense setting, 
  resulting in a total of 12 different simulation settings 
  (6 per sparsity setting).
  To mitigate the effect of random variation, we
  repeated the simulations 100 times and report the results in Figures 
  \ref{fig:boxplots_igaussian_res5_post1} and 
  \ref{fig:boxplots_igaussian_res5_post2}.

<<boxplots_igaussian_res5_post1, fig.cap="Comparison of the NIG and Student's t priors in the sparse setting in (a)-(d) simulation setting 1, (e)-(h) setting 2, and (i)-(l) setting 3, through Cramér-von Mises criterion between the model and true posteriors, correlation and MSD between the posterior means and true parameters, and 95\\% credible interval coverage of the true values.", out.width="100%", fig.asp=1>>=
@

<<boxplots_igaussian_res5_post2, fig.cap="Comparison of the NIG and Student's t priors in the sparse setting in (a)-(d) simulation setting 4, (e)-(h) setting 5, and (i)-(l) setting 6, through Cramér-von Mises criterion between the model and true posteriors, correlation and MSD between the posterior means and true parameters, and 95\\% credible interval coverage of the true values.", out.width="100%", fig.asp=1>>=
@

  [! describe results here]

  \subsection{Variational Bayes approximation}
  To investigate the accuracy of the variational Bayes approximation we compare 
  the variational Bayes posterior to 1000 (after 1000 burn-in) MCMC samples from
  the posterior obtained with \texttt{rstan}. Since the MCMC
  samples represent the exact posterior (up to Monte Carlo error), we use it as 
  a proxy for the true posterior. We include the comparison of our variational
  Bayes approximation to the approximate posterior from the
  automatic differentiation variational inference (ADVI)
  algorithm in \texttt{rstan} \cite[]{kucukelbir_automatic_2015}. ADVI is a
  variational Bayes algorithm that assumes a full factorization of the posterior
  and is fit through gradient based optimization of the variational objective.
  \texttt{rstan} offers an ADVI algorithm that assumes a fullrank posterior
  covariance, but this algorithm often failed to converge or resulted in errors, 
  so we do not include it in the comparisons. We measure divergence between the 
  MCMC and variational posteriors with the Cram{\'e}r-von Mises criterion.

  In addition to ADVI, \texttt{rstan} allows the user
  to maximise the posterior to obtain the maximum \textit{a posteriori} (MAP)
  estimates of the model parameters. We compare these MAP estimates, and the 
  ADVI posterior means,
  to the MAP estimates obtained from our model through the correlation with
  and the mean squared distance (MSD) from the MCMC posterior means.

  The interest of this simulation lies in the variational Bayes approximation, 
  so we restrict ourselves to the univariate setting again (i.e., $D=1$). The
  simulation setup is the same as in the dense setting 
  in Section \ref{sec:inversegammamodel}. Similarly, we use the same 
  hyperparameters settings (Table \ref{tab:hyper_set1}. For \texttt{glmnet}
  we either cross-validate the penalty parameter $\lambda$ or match it to the
  true prior variance. We
  repeated the simulations 100 times and report the results in Figures 
  \ref{fig:boxplots_igaussian_res4.2_post1}-
  \ref{fig:boxplots_igaussian_res4.2_post2}.

  In addition to the performance measures, we have plotted posterior samples 
  histograms of 8 randomly selected model parameters in setting 1, 
  along with the esimated
  marginal variational Bayes posteriors, and the MAP estimates in Figure 
  \ref{fig:hist_igaussian_fit4}.

<<boxplots_igaussian_res4.2_post1, fig.cap="Comparison of different approximations in (a)-(c) simulation setting 1, (d)-(f) setting 2, and (g)-(i) setting 3, through Cramér-von Mises criterion between the approximate and MCMC posteriors, correlation and MSD between the approximate and MCMC posterior means.", out.width="80%", fig.asp=1>>=
@

<<boxplots_igaussian_res4.2_post2, fig.cap="Comparison of different approximations in (a)-(c) simulation setting 4, (d)-(f) setting 5, and (g)-(i) setting 6, through Cramér-von Mises criterion between the approximate and MCMC posteriors, and correlation and MSD between the approximate and MCMC posterior means.", out.width="80%", fig.asp=1>>=
@

<<hist_igaussian_fit4, fig.cap="Histogram of posterior MCMC samples of 8 randomly selected model parameters in setting 1, along with the estimated variational posteriors and MAP estimates.", out.width="80%", fig.asp=1>>=
@

  [! describe results here]

  \subsection{Empirical Bayes estimation}
  To assess the empirical Bayes estimation of the inverse Gaussian prior, 
  we conduct a small simulation study where we simulate from
  model (\ref{eq:priormodel}), while keeping $\forall d: \sigma^2_d=1$ fixed. 
  More specifically, we simulate 
  $\gamma_{jd}^2 \sim \mathcal{IG}([(\mathbf{c}_d^j) \tr \bm{\alpha}]^{-1},
  \lambda_d)$, for $j=1,\dots,200$, $d=1,\dots,10$, where we set 
  $\bm{\alpha}= \begin{bmatrix} 1 & 1 & 2 & 3 & 4 \end{bmatrix}$. We use 
  balanced dummy coding for the $\mathbf{c}_d^j$, such that we end up with
  $\E(\gamma_{jd}) \in \{ 1, 0.5, 0.33, 0.25, 0.2 \}$, and fix 
  $\forall d: \lambda_d=1$.  
  Next, we simulate $\beta_{jd} \sim \mathcal{N} 
  (0, \gamma_{jd}^2)$.
  The observed data is generated from $x_{ij} \sim \mathcal{N} (0, 1)$ and 
  $y_{id} \sim \mathcal{N} (\mathbf{x}_i \tr \bm{\beta}_d, 1)$, for $i=100$. 
  We repeat the simulation 100 times and, for each simulation, estimate the 
  hyperparameters by the method described in Section \ref{sec:empiricalbayes} 
  and present the results in Figure \ref{fig:boxplots_igaussian_res6_eb}.

<<boxplots_igaussian_res6_eb, fig.cap="Empirical Bayes $\\hat{\\bm{\\alpha}}$ estimates", out.width="80%", fig.asp=2/3>>=
@

  [! describe results here]

  - seems to be an overall scaling issue, relative to each other the estimates
  are not so bad.

  - investigate convergence: (i) look at both error and prior variance 
  convergence and (ii) look at actual prior distribution convergence 
  (plot true distribution and estimated prior every 5 iterations or so)

  \section{Application in eQTLs mapping}\label{sec:application}
  This application concerns eQTLs detection from the GEUVADIS data
  \cite[]{lappalainen_transcriptome_2013}. After the removal of 9 genes due to
  missing SNP information, we have 99 genes (outcomes) from the p38MAPK pathway. 
  After removal of the subjects without any expression or SNP data, 373 subjects
  remained. For each gene, we selected SNPs with minor alles frequency (MAF) 
  larger than 0.05 and within 105 bases up- or downstream from the gene. This 
  resulted in 99 sets of SNPs (one for each gene) from the 373 subjects, where 
  the number of SNPs per gene ranges from 56 to 1169.

  We used Ensembl \cite[]{hunt_ensembl_2018} to create several external 
  variables: (i) SNP location 1 (binary, in- or outside of the gene),
  (ii) SNP location 2 (binary, in an exon or an intron), (iii) 
  genelength (continuous), (iv) minor allele frequency of the SNP
  (continuous), and (v) squared MAF. We expect that SNPs that are in the gene 
  and/or in an exon are more important in eQTLs detection. Since a larger gene
  might depend on more SNPs, we expect that longer genes yield lower 
  regression parameter estimates. We predict a positive correlation between
  the regression parameters and the MAF.

  We fit five different models to a subset of the subjects and 
  estimated performance measures on the rest of the subjects. The five models we
  fit differed only in the included external variables. The intercept only model
  is considered a reference model. The other four models all included the 
  genelengths and MAFs (variables (iii) and (iv)). 
  Models 1 and 2 included variable 
  (i), while models 1, 3, and 4 included variable (ii). In addition, model 4
  included the squared MAFs (v).

  For each model, we ran the algorithm for 100 EB iterations and obtained 
  the $\hat{\bm{\alpha}}$ hyperparameter estimates in Table 
  \ref{tab:hyperparameters_eQTL}. In addition, we calculated posterior means
  for the $\bm{\beta}_d$ from MCMC samples from the posterior with the estimated
  hyperparameters. We estimated mean squared error (MSE) for these posterior 
  means. A common measure 
  of Bayesian model fit is the log pseudo marginal likelihood
  (LPML) \cite[]{gelfand_model_1992}, the Bayesian version of the leave-one-out 
  cross-validated log likelihood (see Section \ref{sm-sec:lpml} for more 
  details). We estimated LPML for the different models and added them,
  together with the MSEs to Table \ref{tab:hyperparameters_eQTL}.
<<hyperparameters_eQTL>>= 
load("results/eQTL_fit1.Rdata")
seq.mse <- c(2.0699, 2.0629, 2.0649, 2.0628, 2.0638)
seq.lpml <- round(c(mean(lpml.ienig$lpml), mean(lpml1.enig$lpml),
                    mean(lpml2.enig$lpml), mean(lpml3.enig$lpml), 
                    mean(lpml4.enig$lpml)), 5)
tab <- cbind(round(rbind(c(fit.ienig$eb$alpha, rep(NA, 5)),
                         c(fit1.enig$eb$alpha, NA),
                         c(fit2.enig$eb$alpha[c(1:2)], NA, 
                           fit2.enig$eb$alpha[-c(1:2)], NA),
                         c(fit3.enig$eb$alpha[1], NA, 
                           fit3.enig$eb$alpha[-1], NA),
                         c(fit4.enig$eb$alpha[1], NA, 
                           fit4.enig$eb$alpha[-1])), 2),
             seq.mse, seq.lpml)
tab[which.min(tab[, 7]), 7] <- kableExtra::cell_spec(
  tab[which.min(tab[, 7]), 7], format="latex", bold=TRUE)
tab[which.max(tab[, 8]), 8] <- kableExtra::cell_spec(
  tab[which.min(tab[, 8]), 8], format="latex", bold=TRUE)
rownames(tab) <- c("intercept only", paste("model", 1:4))
options(knitr.kable.NA = '')
kableExtra::kable_styling(kableExtra::column_spec(
  knitr::kable(tab, align=rep("r", ncol(tab)),
               col.names=c("intercept", "in gene", "in exon", "length", 
                           "MAF", "MAF$^2$", "MSE", "LPML"),
               caption="Estimated $\\hat{\\bm{\\alpha}}$ hyperparameters, 
               MSE (lowest in bold), and LPML (highest in bold).", 
               "latex", booktabs=TRUE, escape=FALSE, 
               digits=c(rep(2, 6), 4, 4)), 8, border_left=TRUE),
  latex_options=c("HOLD_position"))
@

  - show influence of co-data with standardized $\bm{\alpha}$'s.

  A quantity that is often of interest in eQTLs detection studies, is the 
  gene expression heritability. It is the proportion of a gene's expression that
  is explained with the SNPs and a measure of the relationship between trait 
  phenotypic variation and genetic variation. Heritability (per gene) is 
  given by $h_d^2 = (\sigma^2_d \sum_{j=1}^{p_d} \gamma_{jd}^2)/
  (\sigma^2_d \sum_{j=1}^{p_d} \gamma_{jd}^2 + \sigma_d^2)$. We assess 
  mean heritability of the five models for a rank-based selection of SNPs.
  We present MCMC-based posterior means and one standard deviation error bars in
  Figure \ref{fig:lines_eqtl_hsq}.
<<lines_eqtl_hsq, fig.cap="Mean heritability in the eQTL detection data.", out.width="80%", fig.asp=2/3>>=
@

  [! describe results here]

  - maybe use explained heritability (fraction of full model heritability)

  - AUC of curve as overal measure

  % Annotation is from GRCh37 reference genome: 
  % \url{https://www.ncbi.nlm.nih.gov/assembly/GCF_000001405.13/}. 
  % Annotation file that Gino used: 
  % \url{https://www.gencodegenes.org/human/release_12.html}.
  % Ensembl version of this genome:
  % \url{https://grch37.ensembl.org/Homo_sapiens/Info/Index}.
  % Possibly use GeuvadisTranscriptExpr R package.

  \section{Discussion}\label{sec:discussion}
  In this paper we have introduced a normal inverse Gaussian prior for 
  multivariate high dimensional linear regression models that is estimated by
  empirical-variational Bayes. The estimation is adaptive to external data
  that describes the relation between the outcomes and features in a data-driven
  manner. In simulations we have shown that the inverse Gaussian model 
  does not perform worse, and often bettern, than a common alternative, 
  the Student's $t$ prior. in addition, the Student's $t$ prior does not admit 
  relatively straightforward estimation, compared to the inverse Gaussian 
  prior, as explained in Section \ref{sec:estimation}. The 
  simulations also show that the model is reasonably well approximated and 
  estimated using empirical-variational Bayes. An eQTL detection application 
  shows that external data may indeed benefit prior estimation and prediction
  performance.

  Apart from eQTL detection, the model may be used for network reconstruction as 
  in \cite{leday_gene_2017}, drug response prediction (manuscript in 
  preparation by the current authors), and other applications with multiple
  outcomes, large feature space, and available external information.

  A drawback of the method is the fact that the external data is required to
  be positive. That may not always be the case in practice. We note, however,
  that the $\mathbf{c}_d^j$ may be translated to be positive by adding a 
  constant. In the usual predictive feature sense this doesn't work, since one
  might still encounter negative future values. Here, the external data is
  only used for estimation and not for prediction, so this issue is 
  non-existent.

  \section*{Software}
	A (developmental) \texttt{R} package is available from 
	\url{https://github.com/magnusmunch/cambridge}.

	\section*{Supplementary Material}
	Supplementary Material is available online from 
	\url{https://github.com/magnusmunch/cambridge}.

	\section*{Reproducible Research}
	All results and documents may be recreated from 
	\url{https://github.com/magnusmunch/cambridge}.

	\section*{Acknowledgements}
	We thank Gino Kpogbezan for providing the pre-processed data in Section 
	\ref{sec:application}.
	\textit{Conflict of Interest}: None declared.

	\bibliographystyle{author_short3.bst}
	\bibliography{refs}

	\section*{Session info}

<<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=
devtools::session_info()
@

\end{document}