### 2020-03-30 docs/supplement1.Rnw

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("code/figures.R")
@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip,xr}
\externaldocument[md-]{manuscript1}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\bsigma}{\bm{\sigma}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bepsilon}{\bm{\epsilon}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bPsi}{\bm{\Psi}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bgamma}{\bm{\gamma}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\Em}{\mathbf{E}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\Lm}{\mathbf{L}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Cov}{\mathbb{C}\text{ov}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Drug sensitivity prediction with normal inverse Gaussian shrinkage 
informed by external data}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@amsterdamumc.nl}{m.munch@amsterdamumc.nl}}, Mark A. van de 
Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

\begin{document}

	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam UMC, VU University, 
	PO Box 7057, 1007 MB Amsterdam, The Netherlands \\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge,
	United Kingdom

	\section{Content overview}
	This document contains the Supplementary Material (SM) to the document 
	`Drug sensitivity prediction with the normal inverse Gaussian prior 
	informed by external data'. In the 
	following, this document is referred to as Main Document (MD).

	Section \ref{sec:prior} describes the moments of the normal
	inverse Gaussian (NIG) prior, Section \ref{sec:vbderivations} gives
	the variational Bayes (BV) derivations that lead to the estimating equations 
	in the MD, Section \ref{sec:elbo} gives the evidence lower bound (ELBO) that 
	is maximised during the VB iterations. In Section \ref{sec:computation} 
	efficient VB calculations are given, while Section 
	\ref{sec:ratiosmodifiedbessels} outlines a computationally stable method
	to calculate ratios of modified Bessel functions. Section \ref{sec:eb} shows
	that the empirical Bayes (EB) scale estimates $\bm{\lambda}$ are non-negative.
	The last Section \ref{sec:gibbssampler} introduces a Gibbs sampling scheme
	to sample from the posterior corresponding to the NIG model.

	\section{Normal inverse Gaussian prior moments}\label{sec:prior}
	Let $\beta \sim \mathcal{N}(0,\sigma^2 \tau^2 \gamma^2)$, where 
	$\sigma^2$ and $\tau^2$ are fixed and $\gamma^2 \sim 
	\mathcal{IG}(\phi,\lambda_{\text{feat}})$, then
	\begin{align*}
	  \E(\beta | \sigma^2, \tau^2) & = 0, \\
	  \V(\beta | \sigma^2, \tau^2) & = \sigma^2 \tau^2 \phi, \\
	  \mathcal{K}(\beta | \sigma^2, \tau^2) & = 
	  \frac{3\phi}{\lambda_{\text{feat}}} + 3,
	\end{align*}
	where $\mathcal{K}(\cdot)$ denotes the kurtosis. Now letting $\tau^2
	\sim \mathcal{IG}(\chi,\lambda_{\text{drug}})$ and integrating it out we have:
	\begin{align*}
	  \E(\beta | \sigma^2) & = 0, \\
	  \V(\beta | \sigma^2) & = \sigma^2 \chi \phi, \\
	  \mathcal{K}(\beta | \sigma^2) & = 
	  3\(\frac{\phi}{\lambda_{\text{feat}}} \frac{\chi}{\lambda_{\text{drug}}} +
	  \frac{\phi}{\lambda_{\text{feat}}} + \frac{\chi}{\lambda_{\text{drug}}}\) + 
	  3.
	\end{align*}

	\section{Hyperparameter settings}\label{sec:hyperparameters}
	For the NIG shrinkage weights prior in Figure \ref{md-fig:dens_kappa}a the 
	following hyperparameter settings (in terms of the inverse Gaussian prior) 
	were used: $\lambda=1$, $\phi=10$ (solid),
	$\lambda=0.1$, $\phi=10$ (dashed), and $\lambda=1$, $\phi=1$ (dotted).
	The Student`s $t$ prior on the regression parameters $\beta_{jd}$ arises if
	$\gamma_{jd}^2 \sim \Gamma^{-1}(a,b)$, with $\Gamma^{-1}(a,b)$ an inverse
	Gamma distribution with shape $a$ and scale $b$. Figure 
	\ref{md-fig:dens_kappa}b is then obtained by setting: $a=0.9$, $b=1.3$ 
	(solid), $a=1$, $b=4$ (dashed), and $a=1$, $b=0.2$ (dotted).
	The lasso prior results from an exponential prior on $\gamma_{jd}^2$. To
	create Figure \ref{md-fig:dens_kappa}c the following settings were
	used for rate $\lambda$: $\lambda=1$ (solid), $\lambda=0.2$ (dashed), and 
	$\lambda=0.1$ (dotted).

	\section{Jeffrey's prior}\label{sec:jeffreysprior}
	We parametrise the data with a Gaussian distribution with unknown mean and 
	variance, i.e., $y_i \sim \mathcal{N} (\mu, \sigma^2)$. This gives us the
	following Hessian (matrix of second derivatives) of the log density:
	$$
	\mathbf{H}(\log f(y_i | \mu, \sigma^2)) = \begin{bmatrix}
	-\frac{1}{\sigma^2} & y_i - \mu \\
	y_i - \mu & \frac{1}{2\sigma^4} - \frac{(y_i - \mu)^2}{\sigma^6}.
	\end{bmatrix}
	$$
	Jeffrey's prior is now found as:
	$$
	\pi(\mu, \sigma^2) \propto |\mathcal{I}(\mu, \sigma^2)|^{1/2} = 
	|-\E (\mathbf{H})|^{1/2} = \begin{vmatrix}
	\frac{1}{\sigma^2} & 0 \\
	0 & \frac{1}{2\sigma^4}
	\end{vmatrix}^{1/2} \propto 1/\sigma^3,
	$$
	where $\mathcal{I}(\theta)$ denotes the Fisher information of $\theta$.

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	For clarity we have indexed the variational 
	density functions with their respective parameters, which we omitted in the 
	MD. In the following all expectations are with respect to the 
	variational posterior $Q_d$. The variational posterior for $\bbeta_d$ is 
	found as follows:
	\begin{align*}
	  \log q_{\bbeta_d} (\bbeta_d) & \propto \E [\log \mathcal{L}
	  (\y_d | \bbeta_d, \sigma_d^2 )] + \E [\log \pi (\bbeta_d | \bgamma_{d}^2, 
	  \sigma_d^2, \tau_d^2)] \\
	  & \propto - \frac{1}{2} \sum_{i=1}^n 
	  \E \[ \frac{(y_{id} - \x_i \tr \bbeta_d)^2}{\sigma_d^2} \] - 
	  \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^2 \sigma_d^2 \tau_d^2} \), \\
	  q_{\bbeta_d} (\bbeta_d) & \overset{D}{=} 
	  \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	  & \text{with } \bSigma_d = \E(\sigma_d^{-2})^{-1} 
	  \{\X \tr \X + \E(\tau_d^{-2})\diag [\E(\gamma_{jd}^{-2})] \}^{-1}, \\
	  & \text{and } \bmu_d = \{\X \tr \X + \E(\tau_d^{-2})
	  \diag [\E(\gamma_{jd}^{-2})] \}^{-1} 
	  \X \tr \y_d.
	\end{align*}
	The variational posterior for $\gamma_{jd}^2$ is given by:
	\begin{align*}
	  \log q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \propto \E 
	  [\log \pi (\beta_{jd} | \gamma_{jd}^{2}, \sigma_d^2, \tau_d^2)] + 
	  \log \pi(\gamma_{jd}^{2}) \\
	  & \propto -\frac{1}{2} \log \gamma_{jd}^{2} - \frac{1}{2} 
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d \tau_d^2}\) \gamma_{jd}^{-2} - 
	  \frac{3}{2} \log \gamma_{jd}^{2} - 
	  \frac{\lambda_{\text{feat}} \{\gamma_{jd}^2 - 
	  \phi_{jd}\}^2}
	  {2\phi_{jd}^{2}} \gamma_{jd}^{-2} \\
	  & \propto \( -1 - 1\) \log \gamma_{jd}^{2} - 
	  \frac{\lambda_{\text{feat}}}{2\phi_{jd}^2} \gamma_{jd}^2 - 
	  \frac{1}{2}\[ \lambda_{\text{feat}} + \E(\tau_d^{-2})\E(\sigma_d^{-2}) 
	  \E(\beta_{jd}^2)\] \gamma_{jd}^{-2} , \\
	  q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-1, \lambda_{\text{feat}}/\phi_{jd}^{2}, \delta_{jd} \), \\
	  & \text{with } \delta_{jd} =\E(\tau_d^{-2})\E(\sigma_d^{-2})
	  \[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + \lambda_{\text{feat}}.
	\end{align*}
	Similarly, we derive the variational posterior for $\tau_d^2$ as:
	\begin{align*}
	  \log q_{\tau_{d}^{2}}(\tau_{d}^{2}) & \propto \E 
	  [\log \pi (\bbeta_{d} | \bgamma_{d}^{2}, \sigma_d^2, \tau_d^2)] + 
	  \log \pi(\tau_{d}^{2}) \\
	  & \propto -\frac{p}{2} \log \tau_{d}^{2} - \frac{1}{2} \sum_{j=1}^p
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d \gamma_{jd}^{2}}\)\tau_d^{-2}  - 
	  \frac{3}{2} \log \tau_{d}^{2} - 
	  \frac{\lambda_{\text{drug}} \{\tau_{d}^2 - \chi_{d}\}^2}
	  {2\chi_{d}^{2}} \tau_{d}^{-2} \\
	  & \propto \( -\frac{p + 1}{2} - 1\) \log \tau_{d}^{2} - 
	  \frac{\lambda_{\text{drug}}}{2\chi_{d}^2} \tau_{d}^2 - 
	  \frac{1}{2}\[ \lambda_{\text{drug}} + \E(\sigma_d^{-2}) \sum_{j=1}^p 
	  \E(\gamma_{jd}^{-2}) \E(\beta_{jd}^2)\] \tau_{d}^{-2} , \\
	  q_{\tau_{d}^{2}}(\tau_{d}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-\frac{p + 1}{2}, \lambda_{\text{drug}}/\chi_{d}^{2}, \eta_{d} \), \\
	  & \text{with } \eta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p
	  \E(\gamma_{jd}^{-2})\[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + 
	  \lambda_{\text{drug}}.
	\end{align*}
	Lastly, we have the variational posterior for $\sigma_d^2$:
	\begin{align*}
	  \log q_{\sigma_d^{2}}(\sigma_d^{2}) & \propto \E 
	  [\log \mathcal{L}(\y_d | \bbeta_d, \sigma_d^2 )] + 
	  \E [\log \pi (\bbeta_d | \bgamma_d^2, \sigma_d^2,\tau_d^2)] + 
	  \log \pi (\sigma_d^{2}) \\
	  & \propto -\frac{n}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{i=1}^n 
	  \frac{\E [(y_{id} - \x_i \tr \bbeta_d)^2]}{\sigma_d^{2}} - 
	  \frac{p}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^{2}\tau_d^2}\) \sigma_d^{-2} - 
	  \frac{3}{2} \log \sigma^2_d \\
	  & = \(-\frac{n + p + 1}{2} - 1 \) \log \sigma_d^{2} \\
	  & \,\,\,\,\,\,\,\,\,\, - 
	  \frac{1}{2}\left\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d)^2] + 
	  \E(\tau_d^{-2}) \sum_{j=1}^p \E(\gamma_{jd}^{-2})
	  \E(\beta_{jd}^2) \right\} \sigma_d^{-2}, \\
	  q_{\sigma_d^{2}}(\sigma_d^{2}) & \overset{D}{=} 
	  \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_{d} \), \\
	  & \text{with } \zeta_{d} = \frac{1}{2} \bigg\{\mathbf{y}_d \tr 
	  \mathbf{y}_d - 2 \mathbf{y}_d \tr \X \E(\bbeta_d) + 
    \trace [ \X \tr \X \V(\bbeta_d)] + 
    \E(\bbeta_d \tr)  \X \tr \X \E(\bbeta_d ) \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \E(\tau_d^{-2})\trace \{ \diag[\E(\gamma_{jd}^{-2})] \V(\bbeta_d)\} + 
    \E(\tau_d^{-2})\E(\bbeta_d \tr) \diag[\E(\gamma_{jd}^{-2})] \E(\bbeta_d )
    \bigg\}.
  \end{align*}
  If we fill in the expectations and variances, we arrive at the estimating
  equations in MD equations (\ref{md-eq:vbequations}).

  \section{Evidence lower bound}\label{sec:elbo}
  The evidence lower bound that is maximised during the iterations is: 
  $\text{ELBO}^{(l)}=\text{ELBO}_d^{(l)}$, with:
  \begin{align*}
    \text{ELBO}&^{(l)}_d \\
    & = -\frac{n+p+1}{2} \log \pi + \frac{p+1-n}{2}\log 2 + \frac{n+1}{2} + p
    + \log \Gamma \(\frac{n + p + 1}{2} \) \\
    & - \frac{n + p + 1}{2} \log \zeta_d^{(l)} + 
    \frac{1}{2} \log |\bSigma_d^{(l)}| - 
    \frac{n+p+1}{4\zeta_d^{(l)}}  \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(l)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(l)}) \\ 
    & + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} +
    g_{d}^{(l)} \trace \[ \diag (b_{jd}^{(l)}) \bm{\Sigma}_d^{(l)}\] + 
    g_{d}^{(l)} (\bm{\mu}_d^{(l)}) \tr \diag (b_{jd}^{(l)}) 
    \bm{\mu}_d^{(l)}\bigg] + p \log \lambda_{\text{feat}}^{(l)} \\
    & + \frac{p + 3}{4} \log \lambda_{\text{drug}}^{(l)} + 
    \frac{\eta_d^{(l)} - \lambda_{\text{drug}}^{(l)}}{2} g_d^{(l)} + 
    \sum_{j=1}^p \frac{\delta_{jd}^{(l)} - \lambda_{\text{feat}}^{(l)}}{2} 
    b_{jd}^{(l)} + \frac{\lambda_{\text{drug}}^{(l)}}{\chi_d^{(l)}} +
    \sum_{j=1}^p \frac{\lambda_{\text{feat}}^{(l)}}{\phi_{jd}^{(l)}} + \\
    & - \frac{p + 1}{2} \log \chi_d^{(l)} - \sum_{j=1}^p \log \phi_{jd}^{(l)}
    -\frac{p+1}{4} \log \eta_d^{(l)} - 
    \frac{1}{2} \sum_{j=1}^p \log \delta_{jd}^{(l)} \\
    & + \log K_{\frac{p+1}{2}}\(\sqrt{\eta_{d}^{(l)}
    \lambda_{\text{drug}}^{(l)}}/
    \chi_{d}^{(l)}\) + \sum_{j=1}^p \log K_{1}\(\sqrt{\delta_{jd}^{(l)}
    \lambda_{\text{feat}}^{(l)}}/\phi_{jd}^{(l)}\).
  \end{align*}

  \section{Efficient computation}\label{sec:computation}
  The empirical Bayes updates require the quantities
  $e_{jd}$, $b_{jd}$, $g_d$, and $f_d$, which in turn depend on the quantities 
  $\diag(\bSigma_d)$, $\bmu_{d}$, $\y_d \tr \X \bmu_d$, 
  $\trace(\X \tr \X \bSigma_d)$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$.
  Let $\mathbf{H}_d=g_d \diag[(b_{jd}^{(h)})^{-1}]$. Then, the first two
  quantities are efficiently calculated as:
  \begin{subequations}\label{eq:complexity1}
    \begin{align}
      \diag (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} 
      \H_d^{(h)} - (a_d^{(h)})^{-1} \{ [ \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} ] \circ
      (\H_d^{(h)} \X \tr ) \} \mathbf{1}_{n \times 1}, \\
      \bmu_d^{(h+1)}) & = \H_d^{(h)} \X \tr \y_d - \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr \y_d.
    \end{align}
  \end{subequations}
  The quantities $\y_d \tr \X \bmu_d$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$
  are easily calculated from the first two. For the remaining one we have:
  \begin{align}\label{eq:complexity2}
    \trace(\X \tr \X \bSigma_d^{(h+1)})  = 
    & (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \H_d^{(h)} \X \tr 
    \mathbf{1}_{n \times 1} \nonumber \\  
    & - (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \{ [  \H_d^{(h)} \X \tr 
    ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr ] \circ 
    \X \tr \} \mathbf{1}_{n \times 1}.
  \end{align}
  Both (\ref{eq:complexity1}) and (\ref{eq:complexity2}) are operations of
  complexity $\mathcal{O}(pn^2)$.

  \section{Ratios of modified Bessel functions}\label{sec:ratiosmodifiedbessels}
  Ratios of modified Bessel functions of the second kind 
  $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large
  $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be
  large, this may cause numerical issues in the calculation of various 
  quantities. We alleviate the numerical issues through the following.

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
    \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
    \begin{cases}
      \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + 
      \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + 
      \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
      \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} +
      \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1},
      & \text{for } p \text{ odd}.
    \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, so that it may be computed
  without numerical issues.

  \section{Empirical Bayes}\label{sec:eb}
  For the feature scale parameter $\lambda_{\text{feat}}$ updates to be 
  non-negative, we have to show that:
  \begin{align}\label{eq:inequality}
    &\sum_{d=1}^D \sum_{j=1}^{p} 
    b_{jd} + \bm{\alpha}_{\text{feat}} \tr \mathbf{C} \tr 
    \diag (e_{jd}) \mathbf{C} \bm{\alpha}_{\text{feat}} - 
    2 \bm{\alpha}_{\text{feat}} \tr \mathbf{C} \tr 
    \mathbf{1}_{pD \times 1}  = \sum_{d=1}^D \sum_{j=1}^{p} \(b_{jd} + 
    \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} \) \geq 0,
  \end{align}
  which holds if all summands in the left-hand side of (\ref{eq:inequality})
  are non-negative.

  Next, we note that through Jensen's inequality 
  $e_{jd}^{-1}=\E(\gamma_{jd}^{2})^{-1} \leq \E(\gamma_{jd}^{-2})=b_{jd}$,
  we may lowerbound the summands in left-hand side of (\ref{eq:inequality}):
  $$
  b_{jd} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} \geq
  e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1}.
  $$
  which we use to show that:
  \begin{align*}
    e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} & \geq 0 \\
    1 + \phi_{jd}^{-2} e_{jd}^2 - 2 \phi_{jd}^{-1} e_{jd} & \geq 0 \\
    \( 1 - \phi_{jd}^{-1} e_{jd} \)^2 & \geq 0.
  \end{align*}
  Similar reasoning holds for the drug scale parameter $\lambda_{\text{drug}}$.

  \section{Estimation of alternative prior}\label{sec:altprior}
  The hyperparameters $\bm{\alpha} = \begin{bmatrix} \bm{\alpha}_{\text{feat}}
  \tr & \bm{\alpha}_{\text{drug}} \tr \end{bmatrix} \tr$ in model 
  (\ref{md-eq:altprior}) are estimated by empirical Bayes maximisation of
  the marginal log likelihood:
  \begin{subequations}\label{eq:altmll}
    \begin{align}
      \log p (\bm{\alpha} | \mathbf{Y}) & = \sum_{d=1}^D
      \log \[ \int_{\bm{\beta}_d} \int_{\sigma^2_d} \mathcal{L} 
      (\mathbf{y}_d | \bm{\beta}_d, \sigma^2_d ) \pi(\bm{\beta}_d | \sigma^2_d) 
      \pi(\sigma^2_d) d\bm{\beta}_d d\sigma_d^2 \] \\
      & \propto - \sum_{d=1}^D \bigg[ \log \left|\I + \mathbf{X} 
      \diag \[ \exp\( \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr 
      \right| \\
      & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + (n + 1) \log \( \mathbf{y}_d \tr 
      \left\{ \I + \mathbf{X} 
      \diag \[ \exp\( \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} 
      \tr\right\}^{-1} \mathbf{y}_d \) \bigg],
    \end{align}
  \end{subequations}
  where $\tilde{\mathbf{C}}_d = \begin{bmatrix} \mathbf{C}_d & 
  \mathbf{1}_{p \times 1} \cdot \mathbf{z}_d \tr \end{bmatrix}$, and
  $\mathbf{C}_d = \begin{bmatrix} \mathbf{c}_{1d} & \cdots & 
  \mathbf{c}_{pd} \end{bmatrix} \tr$.
  The first term in the sum of (\ref{eq:altmll}) is concave in 
  $\bm{\alpha}$, while the second term is convex in $\bm{\alpha}$. This renders
  the marginal log likelihood a sum of non-concave (as well as non-convex) 
  functions. General purpose optimisers, such as \texttt{optim} in \texttt{R} 
  are generally not well suited to maximise non-concave functions. Maximisation
  is therefore done through a minorise-maximisation (MM) algorithm, 
  similar to \cite{zeng_incorporating_2020}.

  The MM algorithm seeks to minorise the marginal log likelihood,
  $\log p(\bm{\alpha} | \mathbf{Y})$, at current
  parameter estimate $\bm{\alpha}^{(t)}$ with function 
  $g(\bm{\alpha} | \bm{\alpha}^{(t)})$, such that
  $\forall \bm{\alpha}: g(\bm{\alpha} | \bm{\alpha}^{(t)}) \leq 
  \log p(\bm{\alpha} | \mathbf{Y})$ 
  and $g(\bm{\alpha}^{(t)} | \bm{\alpha}^{(t)}) = 
  \log p(\bm{\alpha} | \mathbf{Y})$. Then, the sequence
  $$
  \bm{\alpha}^{(t + 1)} = \underset{\bm{\alpha}}{\argmax} 
  g(\bm{\alpha} | \bm{\alpha}^{(t+1)})
  $$
  converges to a local maximum of $f(\bm{\alpha})$. Generally, the minorisation
  function is chosen such that it is concave and easy to maximise. In our case,
  a suitable minorisation is:
  \begin{subequations}\label{eq:minor}
    \begin{align}
      g(\bm{\alpha} | \bm{\alpha}^{(t+1)}) = -\sum_{d=1}^D \bigg[
      & \diag \( \X \tr \mathbf{A}_d^{(t)} \X \) \tr \exp (\tilde{\mathbf{C}}_d
      \bm{\alpha}) \\
      & + \frac{n + 1}{\mathbf{y}_d \tr \mathbf{A}_d^{(t)} \mathbf{y}_d} 
      \mathbf{y}_d \tr \left\{ \I + \mathbf{X} \diag \[ \exp\( 
      \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr \right\}^{-1} 
      \mathbf{y}_d\bigg],
    \end{align}
  \end{subequations}
  where $\mathbf{A}_d^{(t)} = \left\{ \I + \mathbf{X} \diag \[ \exp\( 
  \tilde{\mathbf{C}}_d \bm{\alpha}^{(t)} \) \] \mathbf{X} \tr \right\}^{-1}$.
  Note that (\ref{eq:minor}) is concave. However, it is slow to optimise. 
  In \cite{zeng_incorporating_2020}, another minorization is proposed for
  the second term:
  $$
  \mathbf{y}_d \tr \left\{ \I + \mathbf{X} \diag \[ \exp\( 
  \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr \right\}^{-1} 
  \mathbf{y}_d = \underset{\delta}{\max} 
  - ||\mathbf{y}_d - \mathbf{X} \delta ||_2^2 - 
  \delta \tr \diag \[ \exp\(
  -\tilde{\mathbf{C}}_d \bm{\alpha} \) \] \delta.
  $$
  The extra parameter $\delta$ is estimated by any standard ridge regression
  package, such as \texttt{glmnet} in \texttt{R}.

  \section{Gibbs sampler}\label{sec:gibbssampler}
  MCMC samples from the posterior corresponding to model 
  (\ref{md-eq:prior}) may be generated for each equation independently.
  MCMC samples may be generated by iteratively sampling the following 
  conditional distributions:
  \begin{subequations}\label{eq:gibbssampler}
    \begin{align}
      \bm{\beta}_d | \bgamma_d^2, \sigma_d^2, \tau_d^2, \y_d & \sim 
      \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \label{eq:betasample} \\
      \text{with } & \bm{\Sigma}_d = \sigma_d^2 [\X \tr \X + 
      \tau_d^{-2} \diag (\gamma_{jd}^{-2})]^{-1}, \label{eq:sigmasample} \\
      \text{and } & \bm{\mu}_d = [\X \tr \X + 
      \tau_d^{-2} \diag (\gamma_{jd}^{-2})]^{-1} \X \tr 
      \y_d \label{eq:musample}, \\
      \gamma_{jd}^2 | \beta_{jd}, \sigma_d^2, \tau_d^2, \y &
      \sim \mathcal{GIG}
      \(-1, \lambda_{\text{feat}}/ \phi_{jd}^{2}, \delta_{jd}\), \nonumber \\
      \text{with } & \delta_{jd} = \sigma_d^{-2} \tau_d^{-2} \beta_{jd}^2 + 
      \lambda_{\text{feat}}, \nonumber \\
      \tau_d^2 | \bm{\beta}_d, \sigma_d^2, \bgamma_{d}^2, \y & \sim
      \mathcal{GIG}
      \(-\frac{p+1}{2}, \lambda_{\text{drug}}/ \chi_{d}^{2}, \eta_d\), 
      \nonumber \\
      \text{with } & \eta_d = \sigma_d^{-2} \sum_{j=1}^p \gamma_{jd}^{-2}
      \beta_{jd}^2 + \lambda_{\text{drug}}, \nonumber \\
      \sigma_d^2 | \bm{\beta}_d, \bgamma_d^2, \tau_d^2, \y & \sim \Gamma^{-1} 
      \(\frac{n + p + 1}{2}, \zeta_d\), \nonumber \\
      \text{with } & \zeta_d = \frac{1}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -
      2 \mathbf{y}_d \tr \X \bm{\beta}_d + \bm{\beta}_d \X \tr \X \bm{\beta}_d + 
      \tau^{-2}_d \bm{\beta}_d \tr \diag(\gamma_{jd}^{-2}) \bm{\beta}_d) \]. 
      \nonumber
    \end{align}
  \end{subequations}
  In high dimensional space, the $p \times p$ matrix inversions in 
  (\ref{eq:sigmasample}) and (\ref{eq:musample}) are   
  significant computational bottlenecks. \cite{bhattacharya_fast_2016} describe 
  a method to sample from (\ref{eq:betasample}) without explicit calculation
  of this inverse, thereby offering a siginifcant speed-up compared to naive
  sampling.

  Sampling from (\ref{eq:gibbssampler}) either requires specifying 
  hyperparameters
  $\bm{\alpha}_{\text{feat}}$, $\bm{\alpha}_{\text{drug}}$, 
  $\lambda_{\text{feat}}$, and $\lambda_{\text{drug}}$, or to endow them with
  and extra layer of hyperpriors. In general, specifying them requires 
  rarely available, detailed subject knowledge, so we resort to endowing them
  with possibly flat, improper (as in \cite{upadhyay_bayesian_1996})
  hyperpriors:
  \begin{align*}
    \bm{\alpha}_{\text{feat},g} | \lambda_{\text{feat}} & \sim 
    \mathcal{N}(0, \nu_{\text{feat}}^2/\lambda_{\text{feat}}), \\
    \bm{\alpha}_{\text{drug},h} | \lambda_{\text{drug}} & \sim 
    \mathcal{N}(0, \nu_{\text{drug}}^2/\lambda_{\text{drug}}), \\
    \lambda_{\text{feat}} & \sim \Gamma(\kappa_{\text{feat}}, 
    \xi_{\text{feat}}), \\
    \lambda_{\text{drug}} & \sim \Gamma(\kappa_{\text{drug}}, 
    \xi_{\text{drug}}),
  \end{align*}
  with hyperparameters $\nu_{\text{feat}}^2$, $\nu_{\text{drug}}^2$, 
  $\kappa_{\text{feat}}$, $\xi_{\text{feat}}$, $\kappa_{\text{drug}}$, and
  $\xi_{\text{drug}}$. Improper priors occur if we let 
  $\nu_{\text{feat}}, \nu_{\text{drug}} \to \infty$ and 
  $\kappa_{\text{feat}}, \kappa_{\text{drug}}, \xi_{\text{feat}}, 
  \xi_{\text{drug}} \to 0$. The corresponding full conditionals are
  given by:
  \begin{align*}
    \bm{\alpha}_{\text{feat}} | \lambda_{\text{feat}}, \bm{\Gamma} & \sim 
    \mathcal{N} \( \mathbf{m}_{\text{feat}}, \mathbf{S}_{\text{feat}} \) \\
    \text{with } & \mathbf{S}_{\text{feat}} = \lambda_{\text{feat}}^{-1} 
    [\mathbf{C} \tr \diag(\gamma_{jd}^2) \mathbf{C} +
    \nu_{\text{feat}}^{-2} \I_G]^{-1}, \\
    \text{and } & \mathbf{m}_{\text{feat}} = [\mathbf{C} \tr 
    \diag(\gamma_{jd}^2) \mathbf{C} + \nu_{\text{feat}}^{-2} \I_G]^{-1}
    \mathbf{C} \tr \mathbf{1}_{pD \times 1}, \\
    \bm{\alpha}_{\text{drug}} | \lambda_{\text{drug}}, \bm{\tau} & \sim 
    \mathcal{N} \(\textbf{m}_{\text{drug}}, \mathbf{S}_{\text{drug}} \), \\
    \text{with } & \mathbf{S}_{\text{drug}} = \lambda_{\text{drug}}^{-1} 
    [\mathbf{Z} \tr \diag(\tau_{d}^2) \mathbf{Z} +
    \nu_{\text{drug}}^{-2} \I_H]^{-1}, \\
    \text{and } & \mathbf{m}_{\text{drug}} = [\mathbf{Z} \tr 
    \diag(\tau_{d}^2) \mathbf{Z} + \nu_{\text{drug}}^{-2} \I_H]^{-1}
    \mathbf{Z} \tr \mathbf{1}_{D \times 1}, \\
    \lambda_{\text{feat}} | \bm{\alpha}_{\text{feat}}, \bm{\Gamma} & 
    \sim \Gamma \(\frac{pD}{2} + \kappa_{\text{feat}}, 
    \upsilon_{\text{feat}}\), \\
    \text{with } & \upsilon_{\text{feat}} =\frac{1}{2} 
    \[\mathbf{C} \bm{\alpha}_{\text{feat}} - \diag(\gamma_{jd}^{-2}) 
    \mathbf{1}_{pD \times 1}\] \tr \diag(\gamma_{jd}^2) 
    \[\mathbf{C} \bm{\alpha}_{\text{feat}} - 
    \diag(\gamma_{jd}^{-2}) \mathbf{1}_{pD \times 1}\] + \xi_{\text{feat}}, \\
    \lambda_{\text{drug}} | \bm{\alpha}_{\text{drug}}, \bm{\tau} & 
    \sim \Gamma \(\frac{D}{2} + \kappa_{\text{drug}}, 
    \upsilon_{\text{drug}} \), \\
    \text{with } & \upsilon_{\text{drug}} =\frac{1}{2} 
    \[\mathbf{Z} \bm{\alpha}_{\text{drug}} - \diag(\tau_{d}^{-2}) 
    \mathbf{1}_{D \times 1}\] \tr \diag(\tau_{d}^2) 
    \[\mathbf{Z} \bm{\alpha}_{\text{drug}} - 
    \diag(\tau_{d}^{-2}) \mathbf{1}_{D \times 1}\] + \xi_{\text{drug}}.
  \end{align*}

  \bibliographystyle{author_short3}  
	\bibliography{refs}

	\section*{Session info}

<<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=
devtools::session_info()
@

\end{document}