### 2020-04-15 docs/supplement1.Rnw

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
options(knitr.kable.NA="", digits=3)
@

% read figure code
<<figures, include=FALSE>>=
read_chunk("code/figures.R")
@

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,
rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,
bbm,algorithm,algpseudocode,pgfplotstable,threeparttable,booktabs,mathtools,
dsfont,parskip,xr}
\externaldocument[md-]{manuscript1}

% vectors and matrices
\newcommand{\bbeta}{\bm{\beta}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\bsigma}{\bm{\sigma}}
\newcommand{\balpha}{\bm{\alpha}}
\newcommand{\bepsilon}{\bm{\epsilon}}
\newcommand{\bomega}{\bm{\omega}}
\newcommand{\bOmega}{\bm{\Omega}}
\newcommand{\bSigma}{\bm{\Sigma}}
\newcommand{\bkappa}{\bm{\kappa}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmu}{\bm{\mu}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\blambda}{\bm{\lambda}}
\newcommand{\bLambda}{\bm{\Lambda}}
\newcommand{\bPsi}{\bm{\Psi}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bgamma}{\bm{\gamma}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\W}{\mathbf{W}}
\renewcommand{\O}{\mathbf{O}}
\newcommand{\B}{\mathbf{B}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\Em}{\mathbf{E}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\Vm}{\mathbf{V}}
\newcommand{\rv}{\mathbf{r}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\Sm}{\mathbf{S}}
\newcommand{\0}{\bm{0}}
\newcommand{\tx}{\tilde{\mathbf{x}}}
\newcommand{\hy}{\hat{y}}
\newcommand{\tm}{\tilde{m}}
\newcommand{\tkappa}{\tilde{\kappa}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\C}{\mathbf{C}}
\renewcommand{\v}{\mathbf{v}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\s}{\mathbf{s}}
\renewcommand{\c}{\mathbf{c}}
\newcommand{\ones}{\mathbf{1}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\rvec}{\mathbf{r}}
\newcommand{\Lm}{\mathbf{L}}

% functions and operators
\renewcommand{\L}{\mathcal{L}}
\newcommand{\G}{\mathcal{G}}
\renewcommand{\P}{\mathcal{P}}
\newcommand{\argmax}{\text{argmax} \,}
\newcommand{\expit}{\text{expit}}
\newcommand{\erfc}{\text{erfc}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Cov}{\mathbb{C}\text{ov}}
\newcommand{\tr}{^{\text{T}}}
\newcommand{\diag}{\text{diag}}
\newcommand{\KL}{D_{\text{KL}}}
\newcommand{\trace}{\text{tr}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

% distributions
\newcommand{\N}{\text{N}}
\newcommand{\TG}{\text{TG}}
\newcommand{\Bi}{\text{Binom}}
\newcommand{\PG}{\text{PG}}
\newcommand{\Mu}{\text{Multi}}
\newcommand{\GIG}{\text{GIG}}
\newcommand{\IGauss}{\text{IGauss}}
\newcommand{\Un}{\text{U}}

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}

% settings
\graphicspath{{../figs/}}
\pgfplotsset{compat=1.16}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\title{Drug sensitivity prediction with normal inverse Gaussian shrinkage 
informed by external data}
\date{\today}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: 
\href{mailto:m.munch@amsterdamumc.nl}{m.munch@amsterdamumc.nl}}, Mark A. van de 
Wiel$^{1,3}$, Sylvia Richardson$^{3}$, \\ and Gwena{\"e}l G. R. Leday$^{3}$}

\begin{document}

	\maketitle

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam UMC, VU University, 
	PO Box 7057, 1007 MB Amsterdam, The Netherlands \\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge Institute of Public Health, Cambridge,
	United Kingdom

	\section{Content overview}
	This document contains the Supplementary Material (SM) to the document 
	`Drug sensitivity prediction with the normal inverse Gaussian prior 
	informed by external data'. In the 
	following, this document is referred to as Main Document (MD).

	Section \ref{sec:prior} describes the moments of the normal
	inverse Gaussian (NIG) prior, Section \ref{sec:vbderivations} gives
	the variational Bayes (BV) derivations that lead to the estimating equations 
	in the MD, Section \ref{sec:elbo} gives the evidence lower bound (ELBO) that 
	is maximised during the VB iterations. In Section \ref{sec:computation} 
	efficient VB calculations are given, while Section 
	\ref{sec:ratiosmodifiedbessels} outlines a computationally stable method
	to calculate ratios of modified Bessel functions. Section \ref{sec:eb} shows
	that the empirical Bayes (EB) scale estimates $\bm{\lambda}$ are non-negative.
	The last Section \ref{sec:gibbssampler} introduces a Gibbs sampling scheme
	to sample from the posterior corresponding to the NIG model.

	\section{Normal inverse Gaussian prior moments}\label{sec:prior}
	Let $\beta \sim \mathcal{N}(0,\sigma^2 \tau^2 \gamma^2)$, where 
	$\sigma^2$ and $\tau^2$ are fixed and $\gamma^2 \sim 
	\mathcal{IG}(\phi,\lambda_{\text{feat}})$, then
	\begin{align*}
	  \E(\beta | \sigma^2, \tau^2) & = 0, \\
	  \V(\beta | \sigma^2, \tau^2) & = \sigma^2 \tau^2 \phi, \\
	  \mathcal{K}(\beta | \sigma^2, \tau^2) & = 
	  \frac{3\phi}{\lambda_{\text{feat}}} + 3,
	\end{align*}
	where $\mathcal{K}(\cdot)$ denotes the kurtosis. Now letting $\tau^2
	\sim \mathcal{IG}(\chi,\lambda_{\text{drug}})$ and integrating it out we have:
	\begin{align*}
	  \E(\beta | \sigma^2) & = 0, \\
	  \V(\beta | \sigma^2) & = \sigma^2 \chi \phi, \\
	  \mathcal{K}(\beta | \sigma^2) & = 
	  3\(\frac{\phi}{\lambda_{\text{feat}}} \frac{\chi}{\lambda_{\text{drug}}} +
	  \frac{\phi}{\lambda_{\text{feat}}} + \frac{\chi}{\lambda_{\text{drug}}}\) + 
	  3.
	\end{align*}

	\section{Hyperparameter settings}\label{sec:hyperparameters}
	For the NIG shrinkage weights prior in Figure \ref{md-fig:dens_kappa}a the 
	following hyperparameter settings (in terms of the inverse Gaussian prior) 
	were used: $\lambda=1$, $\phi=10$ (solid),
	$\lambda=0.1$, $\phi=10$ (dashed), and $\lambda=1$, $\phi=1$ (dotted).
	The Student`s $t$ prior on the regression parameters $\beta_{jd}$ arises if
	$\gamma_{jd}^2 \sim \Gamma^{-1}(a,b)$, with $\Gamma^{-1}(a,b)$ an inverse
	Gamma distribution with shape $a$ and scale $b$. Figure 
	\ref{md-fig:dens_kappa}b is then obtained by setting: $a=0.9$, $b=1.3$ 
	(solid), $a=1$, $b=4$ (dashed), and $a=1$, $b=0.2$ (dotted).
	The lasso prior results from an exponential prior on $\gamma_{jd}^2$. To
	create Figure \ref{md-fig:dens_kappa}c the following settings were
	used for exponential rate $\lambda$: $\lambda=1$ (solid), $\lambda=0.2$ 
	(dashed), and $\lambda=0.1$ (dotted).

	\section{Jeffrey's prior}\label{sec:jeffreysprior}
	We parametrise the data with a Gaussian distribution with unknown mean and 
	variance, i.e., $y_i \sim \mathcal{N} (\mu, \sigma^2)$. This results in the
	following Hessian (matrix of second derivatives) of the log density:
	$$
	\mathbf{H}(\log f(y_i | \mu, \sigma^2)) = \begin{bmatrix}
	-\frac{1}{\sigma^2} & y_i - \mu \\
	y_i - \mu & \frac{1}{2\sigma^4} - \frac{(y_i - \mu)^2}{\sigma^6}.
	\end{bmatrix}
	$$
	Jeffreys prior \cite[]{jeffreys_invariant_1946} is now found as:
	$$
	\pi(\mu, \sigma^2) \propto |\mathcal{I}(\mu, \sigma^2)|^{1/2} = 
	|-\E (\mathbf{H})|^{1/2} = \begin{vmatrix}
	\frac{1}{\sigma^2} & 0 \\
	0 & \frac{1}{2\sigma^4}
	\end{vmatrix}^{1/2} \propto 1/\sigma^3,
	$$
	where $\mathcal{I}(\theta)$ denotes the Fisher information of $\theta$.

	\section{Variational Bayes derivations}\label{sec:vbderivations}
	For clarity we have indexed the variational 
	density functions with their respective parameters, which we omitted in the 
	MD. In the following all expectations are with respect to the 
	variational posterior $Q_d$. The variational posterior for $\bbeta_d$ is 
	found as follows:
	\begin{align*}
	  \log q_{\bbeta_d} (\bbeta_d) & \propto \E [\log \mathcal{L}
	  (\y_d | \bbeta_d, \sigma_d^2 )] + \E [\log \pi (\bbeta_d | \bgamma_{d}^2, 
	  \sigma_d^2, \tau_d^2)] \\
	  & \propto - \frac{1}{2} \sum_{i=1}^n 
	  \E \[ \frac{(y_{id} - \x_i \tr \bbeta_d)^2}{\sigma_d^2} \] - 
	  \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^2 \sigma_d^2 \tau_d^2} \), \\
	  q_{\bbeta_d} (\bbeta_d) & \overset{D}{=} 
	  \mathcal{N}_p (\bmu_d, \bSigma_d), \\
	  & \text{with } \bSigma_d = \E(\sigma_d^{-2})^{-1} 
	  \{\X \tr \X + \E(\tau_d^{-2})\diag [\E(\gamma_{jd}^{-2})] \}^{-1}, \\
	  & \text{and } \bmu_d = \{\X \tr \X + \E(\tau_d^{-2})
	  \diag [\E(\gamma_{jd}^{-2})] \}^{-1} 
	  \X \tr \y_d.
	\end{align*}
	The variational posterior for $\gamma_{jd}^2$ is given by:
	\begin{align*}
	  \log q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \propto \E 
	  [\log \pi (\beta_{jd} | \gamma_{jd}^{2}, \sigma_d^2, \tau_d^2)] + 
	  \log \pi(\gamma_{jd}^{2}) \\
	  & \propto -\frac{1}{2} \log \gamma_{jd}^{2} - \frac{1}{2} 
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d \tau_d^2}\) \gamma_{jd}^{-2} - 
	  \frac{3}{2} \log \gamma_{jd}^{2} - 
	  \frac{\lambda_{\text{feat}} \{\gamma_{jd}^2 - 
	  \phi_{jd}\}^2}
	  {2\phi_{jd}^{2}} \gamma_{jd}^{-2} \\
	  & \propto \( -1 - 1\) \log \gamma_{jd}^{2} - 
	  \frac{\lambda_{\text{feat}}}{2\phi_{jd}^2} \gamma_{jd}^2 - 
	  \frac{1}{2}\[ \lambda_{\text{feat}} + \E(\tau_d^{-2})\E(\sigma_d^{-2}) 
	  \E(\beta_{jd}^2)\] \gamma_{jd}^{-2} , \\
	  q_{\gamma_{jd}^{2}}(\gamma_{jd}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-1, \lambda_{\text{feat}}/\phi_{jd}^{2}, \delta_{jd} \), \\
	  & \text{with } \delta_{jd} =\E(\tau_d^{-2})\E(\sigma_d^{-2})
	  \[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + \lambda_{\text{feat}}.
	\end{align*}
	Similarly, we derive the variational posterior for $\tau_d^2$ as:
	\begin{align*}
	  \log q_{\tau_{d}^{2}}(\tau_{d}^{2}) & \propto \E 
	  [\log \pi (\bbeta_{d} | \bgamma_{d}^{2}, \sigma_d^2, \tau_d^2)] + 
	  \log \pi(\tau_{d}^{2}) \\
	  & \propto -\frac{p}{2} \log \tau_{d}^{2} - \frac{1}{2} \sum_{j=1}^p
	  \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d \gamma_{jd}^{2}}\)\tau_d^{-2}  - 
	  \frac{3}{2} \log \tau_{d}^{2} - 
	  \frac{\lambda_{\text{drug}} \{\tau_{d}^2 - \chi_{d}\}^2}
	  {2\chi_{d}^{2}} \tau_{d}^{-2} \\
	  & \propto \( -\frac{p + 1}{2} - 1\) \log \tau_{d}^{2} - 
	  \frac{\lambda_{\text{drug}}}{2\chi_{d}^2} \tau_{d}^2 - 
	  \frac{1}{2}\[ \lambda_{\text{drug}} + \E(\sigma_d^{-2}) \sum_{j=1}^p 
	  \E(\gamma_{jd}^{-2}) \E(\beta_{jd}^2)\] \tau_{d}^{-2} , \\
	  q_{\tau_{d}^{2}}(\tau_{d}^{2}) & \overset{D}{=} \mathcal{GIG} 
	  \(-\frac{p + 1}{2}, \lambda_{\text{drug}}/\chi_{d}^{2}, \eta_{d} \), \\
	  & \text{with } \eta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p
	  \E(\gamma_{jd}^{-2})\[\E(\beta_{jd})^2 + \V(\beta_{jd})\] + 
	  \lambda_{\text{drug}}.
	\end{align*}
	Lastly, we have the variational posterior for $\sigma_d^2$:
	\begin{align*}
	  \log q_{\sigma_d^{2}}(\sigma_d^{2}) & \propto \E 
	  [\log \mathcal{L}(\y_d | \bbeta_d, \sigma_d^2 )] + 
	  \E [\log \pi (\bbeta_d | \bgamma_d^2, \sigma_d^2,\tau_d^2)] + 
	  \log \pi (\sigma_d^{2}) \\
	  & \propto -\frac{n}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{i=1}^n 
	  \frac{\E [(y_{id} - \x_i \tr \bbeta_d)^2]}{\sigma_d^{2}} - 
	  \frac{p}{2} \log \sigma_d^{2} - \frac{1}{2} \sum_{j=1}^p 
	  \E \( \frac{\beta_{jd}^2}{\gamma_{jd}^{2}\tau_d^2}\) \sigma_d^{-2} - 
	  \frac{3}{2} \log \sigma^2_d \\
	  & = \(-\frac{n + p + 1}{2} - 1 \) \log \sigma_d^{2} \\
	  & \,\,\,\,\,\,\,\,\,\, - 
	  \frac{1}{2}\left\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d)^2] + 
	  \E(\tau_d^{-2}) \sum_{j=1}^p \E(\gamma_{jd}^{-2})
	  \E(\beta_{jd}^2) \right\} \sigma_d^{-2}, \\
	  q_{\sigma_d^{2}}(\sigma_d^{2}) & \overset{D}{=} 
	  \Gamma^{-1} \(\frac{n + p + 1}{2}, \zeta_{d} \), \\
	  & \text{with } \zeta_{d} = \frac{1}{2} \bigg\{\mathbf{y}_d \tr 
	  \mathbf{y}_d - 2 \mathbf{y}_d \tr \X \E(\bbeta_d) + 
    \trace [ \X \tr \X \V(\bbeta_d)] + 
    \E(\bbeta_d \tr)  \X \tr \X \E(\bbeta_d ) \\
    & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + 
    \E(\tau_d^{-2})\trace \{ \diag[\E(\gamma_{jd}^{-2})] \V(\bbeta_d)\} + 
    \E(\tau_d^{-2})\E(\bbeta_d \tr) \diag[\E(\gamma_{jd}^{-2})] \E(\bbeta_d )
    \bigg\}.
  \end{align*}
  If we fill in the expectations and variances, we arrive at the estimating
  equations in MD equations (\ref{md-eq:vbequations}).

  \section{Evidence lower bound}\label{sec:elbo}
  The evidence lower bound that is maximised during the iterations is: 
  $\text{ELBO}^{(l)}=\sum_{d=1}^D \text{ELBO}_d^{(l)}$, with:
  \begin{align*}
    \text{ELBO}&^{(l)}_d \\
    & = -\frac{n+p+1}{2} \log \pi + \frac{p+1-n}{2}\log 2 + \frac{n+1}{2} + p
    + \log \Gamma \(\frac{n + p + 1}{2} \) \\
    & - \frac{n + p + 1}{2} \log \zeta_d^{(l)} + 
    \frac{1}{2} \log |\bSigma_d^{(l)}| - 
    \frac{n+p+1}{4\zeta_d^{(l)}}  \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(l)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(l)}) \\ 
    & + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} +
    g_{d}^{(l)} \trace \[ \diag (b_{jd}^{(l)}) \bm{\Sigma}_d^{(l)}\] + 
    g_{d}^{(l)} (\bm{\mu}_d^{(l)}) \tr \diag (b_{jd}^{(l)}) 
    \bm{\mu}_d^{(l)}\bigg] + p \log \lambda_{\text{feat}}^{(l)} \\
    & + \frac{p + 3}{4} \log \lambda_{\text{drug}}^{(l)} + 
    \frac{\eta_d^{(l)} - \lambda_{\text{drug}}^{(l)}}{2} g_d^{(l)} + 
    \sum_{j=1}^p \frac{\delta_{jd}^{(l)} - \lambda_{\text{feat}}^{(l)}}{2} 
    b_{jd}^{(l)} + \frac{\lambda_{\text{drug}}^{(l)}}{\chi_d^{(l)}} +
    \sum_{j=1}^p \frac{\lambda_{\text{feat}}^{(l)}}{\phi_{jd}^{(l)}} + \\
    & - \frac{p + 1}{2} \log \chi_d^{(l)} - \sum_{j=1}^p \log \phi_{jd}^{(l)}
    -\frac{p+1}{4} \log \eta_d^{(l)} - 
    \frac{1}{2} \sum_{j=1}^p \log \delta_{jd}^{(l)} \\
    & + \log K_{\frac{p+1}{2}}\(\sqrt{\eta_{d}^{(l)}
    \lambda_{\text{drug}}^{(l)}}/
    \chi_{d}^{(l)}\) + \sum_{j=1}^p \log K_{1}\(\sqrt{\delta_{jd}^{(l)}
    \lambda_{\text{feat}}^{(l)}}/\phi_{jd}^{(l)}\).
  \end{align*}

  \section{Efficient computation}\label{sec:computation}
  The empirical Bayes updates require the quantities
  $e_{jd}$, $b_{jd}$, $g_d$, and $f_d$, which in turn depend on the quantities 
  $\diag(\bSigma_d)$, $\bmu_{d}$, $\y_d \tr \X \bmu_d$, 
  $\trace(\X \tr \X \bSigma_d)$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$.
  Let $\mathbf{H}_d=g_d \diag[(b_{jd}^{(h)})^{-1}]$. Then, the first two
  quantities are efficiently calculated as:
  \begin{subequations}\label{eq:complexity1}
    \begin{align}
      \diag (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} 
      \H_d^{(h)} - (a_d^{(h)})^{-1} \{ [ \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} ] \circ
      (\H_d^{(h)} \X \tr ) \} \mathbf{1}_{n \times 1}, \\
      \bmu_d^{(h+1)}) & = \H_d^{(h)} \X \tr \y_d - \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr \y_d.
    \end{align}
  \end{subequations}
  The quantities $\y_d \tr \X \bmu_d$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$
  are easily calculated from the first two. For the remaining one we have:
  \begin{align}\label{eq:complexity2}
    \trace(\X \tr \X \bSigma_d^{(h+1)})  = 
    & (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \H_d^{(h)} \X \tr 
    \mathbf{1}_{n \times 1} \nonumber \\  
    & - (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \{ [  \H_d^{(h)} \X \tr 
    ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr ] \circ 
    \X \tr \} \mathbf{1}_{n \times 1}.
  \end{align}
  Both (\ref{eq:complexity1}) and (\ref{eq:complexity2}) are operations of
  complexity $\mathcal{O}(pn^2)$ instead of $\mathcal{O}(p^3)$ for naive
  calculation.

  \section{Ratios of modified Bessel functions}\label{sec:ratiosmodifiedbessels}
  Ratios of modified Bessel functions of the second kind 
  $K_{\alpha - 1}(x)/K_{\alpha}(x)$ are prone to under- and overflow for large
  $\alpha$. In our case, $\alpha$ increases linearly with $p$. Since $p$ may be
  large, this may cause numerical issues in the calculation of various 
  quantities. We alleviate the numerical issues through the following.

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
    \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
    \begin{cases}
      \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + 
      \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + 
      \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
      \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} +
      \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1},
      & \text{for } p \text{ odd}.
    \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, so that an 
  arbitrary ratio $K_{\alpha - 1}(x)/K_{\alpha}(x)$ may be calculated as
  a sequence of numerically stable scalar sums, products and inverses.

  \section{Empirical Bayes}\label{sec:eb}
  For the feature scale parameter $\lambda_{\text{feat}}$ updates to be 
  non-negative, we have to show that:
  \begin{align}\label{eq:inequality}
    &\sum_{d=1}^D \sum_{j=1}^{p} 
    b_{jd} + \bm{\alpha}_{\text{feat}} \tr \mathbf{C} \tr 
    \diag (e_{jd}) \mathbf{C} \bm{\alpha}_{\text{feat}} - 
    2 \bm{\alpha}_{\text{feat}} \tr \mathbf{C} \tr 
    \mathbf{1}_{pD \times 1}  = \sum_{d=1}^D \sum_{j=1}^{p} \(b_{jd} + 
    \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} \) \geq 0,
  \end{align}
  which holds if all summands in the left-hand side of (\ref{eq:inequality})
  are non-negative.

  We note that through Jensen's inequality 
  $e_{jd}^{-1}=\E(\gamma_{jd}^{2})^{-1} \leq \E(\gamma_{jd}^{-2})=b_{jd}$,
  we may lowerbound the summands in left-hand side of (\ref{eq:inequality}):
  $$
  b_{jd} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} \geq
  e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1}.
  $$
  which we use to show that:
  \begin{align*}
    e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} & \geq 0 \\
    1 + \phi_{jd}^{-2} e_{jd}^2 - 2 \phi_{jd}^{-1} e_{jd} & \geq 0 \\
    \( 1 - \phi_{jd}^{-1} e_{jd} \)^2 & \geq 0.
  \end{align*}
  Similar reasoning holds for the drug scale parameter $\lambda_{\text{drug}}$.

  \section{Gibbs sampler}\label{sec:gibbssampler}
  MCMC samples from the posterior corresponding to model 
  (\ref{md-eq:prior}) may be generated for each equation independently.
  MCMC samples may be generated by iteratively sampling the following 
  conditional distributions:
  \begin{subequations}\label{eq:gibbssampler}
    \begin{align}
      \bm{\beta}_d | \bgamma_d^2, \sigma_d^2, \tau_d^2, \y_d & \sim 
      \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \label{eq:betasample} \\
      \text{with } & \bm{\Sigma}_d = \sigma_d^2 [\X \tr \X + 
      \tau_d^{-2} \diag (\gamma_{jd}^{-2})]^{-1}, \label{eq:sigmasample} \\
      \text{and } & \bm{\mu}_d = [\X \tr \X + 
      \tau_d^{-2} \diag (\gamma_{jd}^{-2})]^{-1} \X \tr 
      \y_d \label{eq:musample}, \\
      \gamma_{jd}^2 | \beta_{jd}, \sigma_d^2, \tau_d^2, \y &
      \sim \mathcal{GIG}
      \(-1, \lambda_{\text{feat}}/ \phi_{jd}^{2}, \delta_{jd}\), \\
      \text{with } & \delta_{jd} = \sigma_d^{-2} \tau_d^{-2} \beta_{jd}^2 + 
      \lambda_{\text{feat}}, \nonumber \\
      \tau_d^2 | \bm{\beta}_d, \sigma_d^2, \bgamma_{d}^2, \y & \sim
      \mathcal{GIG}
      \(-\frac{p+1}{2}, \lambda_{\text{drug}}/ \chi_{d}^{2}, \eta_d\), 
      \\
      \text{with } & \eta_d = \sigma_d^{-2} \sum_{j=1}^p \gamma_{jd}^{-2}
      \beta_{jd}^2 + \lambda_{\text{drug}}, \\
      \sigma_d^2 | \bm{\beta}_d, \bgamma_d^2, \tau_d^2, \y & \sim \Gamma^{-1} 
      \(\frac{n + p + 1}{2}, \zeta_d\), \\
      \text{with } & \zeta_d = \frac{1}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -
      2 \mathbf{y}_d \tr \X \bm{\beta}_d + \bm{\beta}_d \X \tr \X \bm{\beta}_d + 
      \tau^{-2}_d \bm{\beta}_d \tr \diag(\gamma_{jd}^{-2}) \bm{\beta}_d) \]. 
    \end{align}
  \end{subequations}
  In high dimensional space, the $p \times p$ matrix inversions in 
  (\ref{eq:sigmasample}) and (\ref{eq:musample}) are   
  significant computational bottlenecks. \cite{bhattacharya_fast_2016} describe 
  a method to sample from (\ref{eq:betasample}) without explicit calculation
  of this inverse, thereby offering a siginifcant speed-up compared to naive
  sampling.

  Sampling from (\ref{eq:gibbssampler}) either requires estimating 
  (as in \ref{md-sec:empiricalbayes}) or specifying 
  hyperparameters
  $\bm{\alpha}_{\text{feat}}$, $\bm{\alpha}_{\text{drug}}$, 
  $\lambda_{\text{feat}}$, and $\lambda_{\text{drug}}$, or to endow them with
  and extra layer of hyperpriors. In general, specifying them requires 
  rarely available, detailed subject knowledge, so we may resort to endowing 
  them with hyperpriors (as in \cite{upadhyay_bayesian_1996}):
  \begin{align*}
    \bm{\alpha}_{\text{feat},g} | \lambda_{\text{feat}} & \sim 
    \mathcal{N}(0, \nu_{\text{feat}}^2/\lambda_{\text{feat}}), \\
    \bm{\alpha}_{\text{drug},h} | \lambda_{\text{drug}} & \sim 
    \mathcal{N}(0, \nu_{\text{drug}}^2/\lambda_{\text{drug}}), \\
    \lambda_{\text{feat}} & \sim \Gamma(\kappa_{\text{feat}}, 
    \xi_{\text{feat}}), \\
    \lambda_{\text{drug}} & \sim \Gamma(\kappa_{\text{drug}}, 
    \xi_{\text{drug}}),
  \end{align*}
  with hyperparameters $\nu_{\text{feat}}^2$, $\nu_{\text{drug}}^2$, 
  $\kappa_{\text{feat}}$, $\xi_{\text{feat}}$, $\kappa_{\text{drug}}$, and
  $\xi_{\text{drug}}$. Improper, flat priors occur if we let 
  $\nu_{\text{feat}}, \nu_{\text{drug}} \to \infty$ and 
  $\kappa_{\text{feat}}, \kappa_{\text{drug}}, \xi_{\text{feat}}, 
  \xi_{\text{drug}} \to 0$. The corresponding full conditionals are
  given by:
  \begin{align*}
    \bm{\alpha}_{\text{feat}} | \lambda_{\text{feat}}, \bm{\Gamma} & \sim 
    \mathcal{N} \( \mathbf{m}_{\text{feat}}, \mathbf{S}_{\text{feat}} \) \\
    \text{with } & \mathbf{S}_{\text{feat}} = \lambda_{\text{feat}}^{-1} 
    [\mathbf{C} \tr \diag(\gamma_{jd}^2) \mathbf{C} +
    \nu_{\text{feat}}^{-2} \I_G]^{-1}, \\
    \text{and } & \mathbf{m}_{\text{feat}} = [\mathbf{C} \tr 
    \diag(\gamma_{jd}^2) \mathbf{C} + \nu_{\text{feat}}^{-2} \I_G]^{-1}
    \mathbf{C} \tr \mathbf{1}_{pD \times 1}, \\
    \bm{\alpha}_{\text{drug}} | \lambda_{\text{drug}}, \bm{\tau} & \sim 
    \mathcal{N} \(\textbf{m}_{\text{drug}}, \mathbf{S}_{\text{drug}} \), \\
    \text{with } & \mathbf{S}_{\text{drug}} = \lambda_{\text{drug}}^{-1} 
    [\mathbf{Z} \tr \diag(\tau_{d}^2) \mathbf{Z} +
    \nu_{\text{drug}}^{-2} \I_H]^{-1}, \\
    \text{and } & \mathbf{m}_{\text{drug}} = [\mathbf{Z} \tr 
    \diag(\tau_{d}^2) \mathbf{Z} + \nu_{\text{drug}}^{-2} \I_H]^{-1}
    \mathbf{Z} \tr \mathbf{1}_{D \times 1}, \\
    \lambda_{\text{feat}} | \bm{\alpha}_{\text{feat}}, \bm{\Gamma} & 
    \sim \Gamma \(\frac{pD}{2} + \kappa_{\text{feat}}, 
    \upsilon_{\text{feat}}\), \\
    \text{with } & \upsilon_{\text{feat}} =\frac{1}{2} 
    \[\mathbf{C} \bm{\alpha}_{\text{feat}} - \diag(\gamma_{jd}^{-2}) 
    \mathbf{1}_{pD \times 1}\] \tr \diag(\gamma_{jd}^2) 
    \[\mathbf{C} \bm{\alpha}_{\text{feat}} - 
    \diag(\gamma_{jd}^{-2}) \mathbf{1}_{pD \times 1}\] + \xi_{\text{feat}}, \\
    \lambda_{\text{drug}} | \bm{\alpha}_{\text{drug}}, \bm{\tau} & 
    \sim \Gamma \(\frac{D}{2} + \kappa_{\text{drug}}, 
    \upsilon_{\text{drug}} \), \\
    \text{with } & \upsilon_{\text{drug}} =\frac{1}{2} 
    \[\mathbf{Z} \bm{\alpha}_{\text{drug}} - \diag(\tau_{d}^{-2}) 
    \mathbf{1}_{D \times 1}\] \tr \diag(\tau_{d}^2) 
    \[\mathbf{Z} \bm{\alpha}_{\text{drug}} - 
    \diag(\tau_{d}^{-2}) \mathbf{1}_{D \times 1}\] + \xi_{\text{drug}},
  \end{align*}
  which may be sampled together with (\ref{eq:gibbssampler}).

  \section{Multivariate Bayesian extension of \texttt{xtune}}\label{sec:mxtune}
  Consider the following simpler alternative prior to 
  equation (\ref{md-eq:prior}) in the MD:
  \begin{subequations}\label{eq:altprior}
		\begin{align}
		  \beta_{jd} | \gamma^2_{jd}, \tau_d^2, \sigma^2_d & \sim \mathcal{N}_{p} 
		  (0, \gamma_{jd}^2 \tau_d^2 \sigma_d^2), \\
		  \log \gamma_{jd}^{2} & = \mathbf{c}_{jd} \tr \bm{\alpha}_{\text{feat}}, \\
		  \log \tau_d^{2} & = \mathbf{z}_{d} \tr \bm{\alpha}_{\text{drug}}\\
		  \sigma_d^{2} & \sim \Gamma^{-1} (\nu, \zeta).
		\end{align}
	\end{subequations}
	This model is a Bayesian, multivariate extension of 
	\cite{zeng_incorporating_2020}. The model does 
	not propagate uncertainty in the variance components $\gamma_{jd}^{2}$
	and $\tau_d^{2}$, but directly models the variance components as an
	exponentiated linear function of the external covariates. The linear functions
	are modelled on the log-scale to ensure non-negative variance components.
	Additionally, it allows for more flexibility in the error variance
	prior model. Note that in the limit $\zeta \to 0$ and $\nu=1/2$, the error
	variance follows the Jeffreys prior from (\ref{md-eq:prior}) in the MD. 
	Throughout the
	following, whenever $\zeta$ and $\nu$ are not specified, we use $\zeta \to 0$ 
	and $\nu=1/2$.

  The hyperparameters $\bm{\alpha} = \begin{bmatrix} \bm{\alpha}_{\text{feat}}
  \tr & \bm{\alpha}_{\text{drug}} \tr \end{bmatrix} \tr$ in model 
  (\ref{eq:altprior}) are estimated by empirical Bayes maximisation of
  the marginal log likelihood:
  \begin{subequations}\label{eq:altmll}
    \begin{align}
      \log p (\bm{\alpha} | \mathbf{Y}) & = \sum_{d=1}^D
      \log \[ \int_{\bm{\beta}_d} \int_{\sigma^2_d} \mathcal{L} 
      (\mathbf{y}_d | \bm{\beta}_d, \sigma^2_d ) \pi(\bm{\beta}_d | \sigma^2_d) 
      \pi(\sigma^2_d) d\bm{\beta}_d d\sigma_d^2 \] \\
      & \propto - \sum_{d=1}^D \bigg[ \log \left|\I + \mathbf{X} 
      \diag \[ \exp\( \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr 
      \right| \\
      & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + (n + 1) \log \( \mathbf{y}_d \tr 
      \left\{ \I + \mathbf{X} 
      \diag \[ \exp\( \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} 
      \tr\right\}^{-1} \mathbf{y}_d \) \bigg],
    \end{align}
  \end{subequations}
  where $\tilde{\mathbf{C}}_d = \begin{bmatrix} \mathbf{C}_d & 
  \mathbf{1}_{p \times 1} \cdot \mathbf{z}_d \tr \end{bmatrix}$, and
  $\mathbf{C}_d = \begin{bmatrix} \mathbf{c}_{1d} & \cdots & 
  \mathbf{c}_{pd} \end{bmatrix} \tr$.
  The first term in the sum of (\ref{eq:altmll}) is concave in 
  $\bm{\alpha}$, while the second term is convex in $\bm{\alpha}$. This renders
  the marginal log likelihood a sum of non-concave (as well as non-convex) 
  functions. General purpose optimisers, such as \texttt{optim} in \texttt{R} 
  are generally not well suited to maximise non-concave functions. Maximisation
  is therefore done through a minorise-maximisation (MM) algorithm, 
  similar to \cite{zeng_incorporating_2020}.

  The MM algorithm seeks to minorise the marginal log likelihood,
  $\log p(\bm{\alpha} | \mathbf{Y})$, at current
  parameter estimate $\bm{\alpha}^{(l)}$ with function 
  $g(\bm{\alpha} | \bm{\alpha}^{(l)})$, such that
  $\forall \bm{\alpha}: g(\bm{\alpha} | \bm{\alpha}^{(l)}) \leq 
  \log p(\bm{\alpha} | \mathbf{Y})$ 
  and $g(\bm{\alpha}^{(l)} | \bm{\alpha}^{(l)}) = 
  \log p(\bm{\alpha}^{(l)} | \mathbf{Y})$. Then, the sequence
  $$
  \bm{\alpha}^{(l + 1)} = \underset{\bm{\alpha}}{\argmax} 
  g(\bm{\alpha} | \bm{\alpha}^{(l)})
  $$
  converges to a local maximum of $\log p(\bm{\alpha} | \mathbf{Y})$. 
  Generally, the minorisation
  function is chosen such that it is concave and easy to maximise. In our case,
  a suitable minorisation is:
  \begin{subequations}\label{eq:minor}
    \begin{align}
      g(\bm{\alpha} | \bm{\alpha}^{(l)}) = -\sum_{d=1}^D \bigg[
      & \diag \( \X \tr \mathbf{A}_d^{(l)} \X \) \tr \exp (\tilde{\mathbf{C}}_d
      \bm{\alpha}) \\
      & + \frac{n + 1}{\mathbf{y}_d \tr \mathbf{A}_d^{(l)} \mathbf{y}_d} 
      \mathbf{y}_d \tr \left\{ \I + \mathbf{X} \diag \[ \exp\( 
      \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr \right\}^{-1} 
      \mathbf{y}_d\bigg],
    \end{align}
  \end{subequations}
  where $\mathbf{A}_d^{(l)} = \left\{ \I + \mathbf{X} \diag \[ \exp\( 
  \tilde{\mathbf{C}}_d \bm{\alpha}^{(l)} \) \] \mathbf{X} \tr \right\}^{-1}$.
  Note that (\ref{eq:minor}) is concave. However, it is slow to optimise. 
  In \cite{zeng_incorporating_2020}, another minorization is proposed for
  the second term:
  $$
  \mathbf{y}_d \tr \left\{ \I + \mathbf{X} \diag \[ \exp\( 
  \tilde{\mathbf{C}}_d \bm{\alpha}^{(l)} \) \] \mathbf{X} \tr \right\}^{-1} 
  \mathbf{y}_d = \underset{\bm{\delta}_d}{\max} 
  - ||\mathbf{y}_d - \mathbf{X} \bm{\delta}_d ||_2^2 - 
  \bm{\delta}_d \tr \diag \[ \exp\(
  -\tilde{\mathbf{C}}_d \bm{\alpha}^{(l)} \) \] \bm{\delta}_d.
  $$
  The extra parameters $\bm{\delta}_d$ are estimated per drug and by any 
  standard ridge regression package, such as \texttt{glmnet} in \texttt{R}.

  \section{Extra simulation results}\label{sec:simulations}
  The lasso and ridge models were fit separately per drug, using the \texttt{R}
	package \texttt{glmnet} \cite[]{friedman_regularization_2010} with cross
	validated penalty parameters. The \texttt{xtune} model is fit with the 
	corresponding \texttt{R} package \cite[]{ni_bayesian_2019}.

<<simulation_gdsc>>=
# simulation 1
res <- read.table("results/simulations_gdsc_res1.txt", row.names=NULL)
temp <- res[, 1]
res <- as.matrix(res[, -1])
rownames(res) <- temp

tabm <- aggregate(res, by=list(substr(rownames(res), 1, 5)), FUN="mean")[, -6]
tabm <- t(tabm[tabm$Group.1 %in% c("emse.", "emseh", "emsel", "pmse."), -1])

D <- nreps <- 100
tabsd <- sapply(c("emse.", "emseh", "emsel", "pmse."), function(s) {
  apply(aggregate(res[substr(rownames(res), 1, 5)==s, ], 
                  list(rep(1:nreps, each=D)), mean), 2, sd)})[-c(1, 6), ]

tab1 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab1) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab1) <- c("NIG$_{\\text{f}}^-$", 
                     "NIG$_{\\text{f}}$", "ridge", "lasso", "mxtune")

# simulation 2
res <- read.table("results/simulations_gdsc_res2.txt", row.names=NULL)
temp <- res[, 1]
res <- as.matrix(res[, -1])
rownames(res) <- temp

tabm <- aggregate(res, by=list(substr(rownames(res), 1, 5)), FUN="mean")[, -6]
tabm <- t(tabm[tabm$Group.1 %in% c("emse.", "emseh", "emsel", "pmse."), -1])

D <- nreps <- 100
tabsd <- sapply(c("emse.", "emseh", "emsel", "pmse."), function(s) {
  apply(aggregate(res[substr(rownames(res), 1, 5)==s, ], 
                  list(rep(1:nreps, each=D)), mean), 2, sd)})[-c(1, 6), ]

tab2 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab2) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab2) <- c("NIG$_{\\text{d}}^-$", 
                     "NIG$_{\\text{d}}$", "ridge", "lasso")

# simulation 3
res <- read.table("results/simulations_gdsc_res3.txt", row.names=NULL)
temp <- res[, 1]
res <- as.matrix(res[, -1])
rownames(res) <- temp

tabm <- tabm3 <- aggregate(res, by=list(substr(rownames(res), 1, 5)), 
                           FUN="mean")[, -6]
tabm <- t(tabm[tabm$Group.1 %in% c("emse.", "emseh", "emsel", "pmse."), -1])

D <- nreps <- 100
tabsd <- tabsd3 <- sapply(c("emse.", "emseh", "emsel", "pmse."), function(s) {
  apply(aggregate(res[substr(rownames(res), 1, 5)==s, ], 
                  list(rep(1:nreps, each=D)), mean), 2, sd)})[-c(1, 6), ]

tab3 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab3) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab3) <- c("NIG$_{\\text{d}}^-$", 
                     "NIG$_{\\text{d}}$", "ridge", "lasso", "mxtune")
@	
	Tables \ref{tab:simulations_gdsc_tab1}-\ref{tab:simulation_gdsc_tab3} 
	display averaged estimation mean squared error (EMSE), and prediction mean
	squared error (PMSE), calculated on independent test data, for simulation 
	Scenarios 1-4. PMSE is calculated as in Section \ref{md-sec:gdsc} of the MD, 
	while EMSE is calculated as: $\text{EMSE} = D^{-1} p^{-1} \sum_{d=1}^D 
  \sum_{j=1}^p (\beta_{jd} - \hat{\beta}_{jd})^2$, with
  $\hat{\beta}_{jd}$ the estimator for $\beta_{jd}$. In the NIG
  models, that provide full posteriors, the posterior mean 
  $\E(\beta_{jd} | \mathbf{y}_d)$ is used as point estimate.
	EMSE is further split into the contribution of the
	bottom 90\% of the $\beta_{jd}$ in terms of size (in absolute value) and the 
	top 10\% of the $\beta_{jd}$ in size.

%   
%   
%   $\lambda_{\text{feat}}$ is overestimated with median estimate 
% 	\Sexpr{round(est.lambdaf[2], 2)} (the
% 	true value is 1). The overestimated $\lambda_{\text{feat}}$ translates 
% 	into underestimated $\gamma_{jd}^2$ variances. 
% 	
% 	
% 	
<<simulations_gdsc_tab1>>=
kableExtra::kable_styling(knitr::kable(
  tab1, align="r",
  caption="Mean EMSE (split for bottom 10\\% and top 90\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 1, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_tab2>>=
kableExtra::kable_styling(knitr::kable(
  tab2, align="r",
  caption="Mean EMSE (split for bottom 10\\% and top 90\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 2, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_tab3>>=
kableExtra::kable_styling(knitr::kable(
  tab3, align="r",
  caption="Mean EMSE (split for bottom 10\\% and top 90\\% of $|\\hat{\\beta}_{jd}|$) and PMSE for simulation Scenario 2, estimated on test data (lowest in bold).", 
  format="latex", booktabs=TRUE, escape=FALSE), 
  latex_options=c("HOLD_position"))
@

<<simulations_gdsc_pmse4, fig.cap="Simulation results for Scenario 4: Average PMSE against the proportion of permuted rows in the external features.", out.width="100%", fig.asp=1>>= 
@

	Focussing on EMSE, we see that the NIG$_{\text{f}}$ model,
	that includes the external covariates, performs best, both in terms of 
	estimation and prediction. The lower MSEs confirm that the model that
	includes the external data (NIG$_{\text{f}}$) learns the underlying structure 
	in the data better than model NIG$_{\text{f}}^-$ that does not include the
	external data. Furthermore, we see that estimation of the larger $\beta_{jd}$ 
	is where ridge loses the most compared to the NIG and lasso models. 
	The heavier tails of the NIG and lasso priors allow for larger $\beta_{jd}$
	estimates and thus lower EMSE for these $\beta_{jd}$.

%   $\lambda_{\text{feat}}$ is overestimated with median estimate
% 	\Sexpr{round(est.lambdaf[1], 2)} (the true value is one) and thus 
% 	again underestimates the 
% 	prior $\gamma_{jd}^2$ variance.
% 	
% 	Median $\lambda_{\text{drug}}=\Sexpr{round(est.lambdad[2], 2)}$ is
% 	overestimated compared to true value one. 
% 	Scale $\lambda_{\text{drug}}$ is underestimated with median estimate
% 	\Sexpr{round(est.lambdad[1], 2)} compared to the true value one.
% 	
%   Median performance measures for simulation Scenario 2 are displayed in Table
% 	\ref{tab:simulations_gdsc_tab2}. Again, we see that NIG$_{\text{d}}$ performs
% 	best in terms of MSE. However, NIG$_{\text{d}}^-$ performance is the same 
% 	here, indicating that the external covariates do not add much here.
% 	Similarly as in Scenario 1, most of the estimation performance gain over ridge
% 	is through the top 10\% of the $\beta_{jd}$.
% 	
%   
%   Median $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are
% 	are over- and underestimated with \Sexpr{round(est.lambdaf[2], 2)}
% 	and \Sexpr{round(est.lambdad[2], 2)}, respectively, compared to their true
% 	value 1. 
% 	
%   NIG$_{\text{f}+\text{d}}^-$ also over- and
% 	underestimates 
% 	$\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ with median estimates
% 	\Sexpr{round(est.lambdaf[1], 2)}
% 	and \Sexpr{round(est.lambdad[1], 2)}, respectively. 
% 	
%   Table 
% 	\ref{tab:simulations_gdsc_tab3}, with median performance measures, shows
% 	that NIG$_{\text{f}+\text{d}}$ outperforms the other models in terms of
% 	prediction and estimation MSE, with again most of the edge over ridge in the
% 	top 10\% of the $\beta_{jd}$.

<<simulations_gdsc_post5, fig.cap="Simulation results for Scenario 3: Samples from posterior with variational posterior superimposed (red line) for estimated hyperparameters (top row) $\\hat{\\chi}_1$, (second row) $\\hat{\\chi}_2$, (third row) $\\hat{\\chi}_3$, (bottom row) $\\hat{\\chi}_4$, (left column) $\\hat{\\phi}_1$, (second column) $\\hat{\\phi}_2$, (third column) $\\hat{\\phi}_3$, and (last column) $\\hat{\\phi}_4$.", out.width="100%", fig.asp=1>>= 
@

  \section{Extra GDSC data results}\label{sec:gdsc}
  $$
  \text{CPO}_{id} = p(y_i | \bm{y}_{-i}) = 
  $$

<<analysis_gdsc_res2, fig.cap="Simulation results for Scenario 4: Median PMSE against the proportion of permuted rows in the external features.", out.width="50%", fig.asp=1>>= 
@

  \bibliographystyle{author_short3}  
	\bibliography{refs}

	\section*{Session info}

<<r session-info, eval=TRUE, echo=TRUE, tidy=TRUE>>=
devtools::session_info()
@

\end{document}